### 2014-08-08 Article.tex
Tex commit

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[french, english]{babel}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Bare Demo of IEEEtran.cls for Conferences}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678-2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
There is a growing interest for coalitional concepts in the smart grid, spanning multiple objectives such as power losses reduction, benefits maximization or market stability... In this paper, we study coalitions of prosumers (agents that produce but also consume depending on time) which aim at selling energy to the grid. We reckon that the grid is able to specify two requirements (stability and minimum production) that enable a coalition to enter the market. Based on probability distributions of units available power, coalitions which are able to achieve low standard deviations are considered as stable coalitions. As it seems intuitively understandable that maintaining stable coalitions would be less expensive in terms of communication load than highly versatible ones, our objective consists in forming coalitions that satisfy grid requirements both in terms of stability and minimum production. Our definition of stability being implicitely linked to correlation between agents timeseries (see section 2), we organize them over a correlation graph and use a variation of clique percolation to construct valid coalitions in reasonable time.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
One of the most basic ideas in the smart grid revolves around the introduction of communication means in the power grids, which could enable complex improvements in the energy management and lead progressively to a greener energetic system [1] [2]. Distributed energy resources (DER) such as wind turbines or photovoltaic panels are supposed to emerge and populate not only spread out farms, but also classical residential neighborhoods. Together with electric vehicles, and demand side management tools, they constitute the building blocks which will help turn the today pure energy consumers into true actors of the grid operation [3]. Such energy aware agents that consume, produce, and are ready to make concessions (appliances delays, V2G…) to ensure grid stability are commonly called prosumers [4] [5].

It is rather clear that there should be communication flows between prosumers and the grid as most of the demand side management concepts suppose such a link [6] [7]. Nevertheless as the expected number of active prosumers in the future seems quite large, the literature assume more and more communication between prosumers as this is a key ingredient to maintain more complex structures (such as coalitions), and thereby decrease the communication load on the grid’s side [8]. Furthermore, it has been shown that coalitions in the smart grid provide nice ways of improving stability while posing exciting challenges in terms of network infrastructure and management. Self sufficient microgrids that can disconnect from the main grid [9], virtual power plants (VPP) that map small DER in sizeable and adjustable plants [10], or coalitions of electric vehicles that back up the grid in emergency situations (V2G) [11] are just a few examples of how coalitional concepts can enhance grid reliability.

In this paper, our focus is on how coalitions of prosumers, aiming at selling energy to the grid, should be formed in order to support it efficiently. Namely, improving statistically production stability while decreasing communication needs for staying in stable state. The idea is that, by using prediction techniques, coalitions can propose contract values to the grid, enabling it to schedule production on its side accordingly. Actually, contract values are particularly relevant in power exchanges contexts where energy is traded based on predictions (day-ahead estimations for instance). It seems obvious that, in such systems, participants should try to minimize prediction errors in order to maintain a stable state of grid operation, and that some penalty rules should be applied to ensure that coalitions are willing to report contract values correctly. Furthermore, as the coalition productions are supposed to come from (uncertain) renewables, the notion of reliability comes into play with a visible tradeoff between contract values worth and reliability.

In this paper, our concern is on how prosumers should be aggregated in order to optimize this tradeoff, that is, how coalitions should be formed with the intention of announcing “high contract values with high reliability”. We thus aim at:

\begin{itemize}
\item Realistic representations (prosumers productions are non independent variables and come (indirectly) from real traces)
\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section 2)
\item Being (relatively) scalable with the number of prosumers
\end{itemize}

In more details, and as will be explained in section 2, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).  We used real meteorological traces [13] to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations. With these ingredients, we are able to run simulations and record the different output profiles as timeseries.

As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering [16] [17] (see section 3). We will see that in such graphs, cliques represent structures of interest. In our settings, it means that cliques are “good” (from the utility point of view) foundations for high valued coalitions. We thereby use a slightly modified clique percolation algorithm (see section 4) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements (see results in section 5).

\section{Related Work}
Related work text here.

\section{Model}
Smart grid perspectives assume that smart meters populate the network, recording production/consumption values and reporting these to data centers. Any entities (DER, prosumers, microgrids, VPP and so on) would thereby be characterizable through timeseries expressing how much these entities produce or consume.

Through our model, we wish thus to be able to generate, for any prosumer, its available production over time, which depend both on its pure production and consumption. Naturally, these are far from random in the sense that there are recurring patterns (consumption or expected solar irradience for instance) or statistical properties (see wind speed studies) that shape these profiles. We thus used meteorological traces as base inputs because they naturaly provide these key ingredients (sources). On the production side, as we only consider wind turbines and photovoltaic panels as generators, we concentrate primarily on wind speed and solar irradience traces. Variables such as temperature or humidity that correlates very well with consumption were also of great use.

Basically, the user only has to indicate the number of prosumers and the simulation dates. Prosumers are then positioned randomly on a squared lattice. Any point of this lattice is linked to a meteorological profile derived from the input traces. A few parameters individualizing the prosumers (number and kind of DER owned, consumption habits, and so on...) are also picked randomly and account for diversity.

We denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value of agent i at instant t (its instantaneous production minus consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t). During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval, we denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting timeserie of agent i. For any agent i, we note $ \mathcal{P}_{i} $ the probability distribution drawn from $ \mathcal{T}_{i} $ and by $ P_{i} $ a random variable following $ \mathcal{P}_{i} $.

We now extand these notations to any coalition $ S \subset \mathcal{A} $ : 
\begin{itemize}
\item $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $
\item $ \mathcal{T}_{S} = \{ P_{S}(t_{0}),...,P_{S}(t_{K}) \} $ with $ \mathcal{P}_{S} $ its probability distribution
\item A coalition S has the possibility to announce a contract value $ P_{S}^{CRCT} $ on the market.
\end{itemize}

In this paper, we consider that coalition announcements are constrained by the grid. Namely, the grid has the possibility of fixing two parameters :

\begin{itemize}
\item the reliability (denoted by $ \phi \in [0,1] $)
\item the minimum contract value under a given reliability ( $ P_{\phi}^{MIN} $). 
\end{itemize}

The reliability stipulates that for any coalition S willing to join the market, the probability of S value (at any instant t) being below its contract value should at most be $ \phi $. That is, $ \forall S,\ \forall t,\ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] \leq \phi $. Furthermore, for consistency, we restrict $ \phi $ to small values ( $ \phi << 0.5 $). The minimum contract value under $ \phi $ states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P_{\phi}^{MIN} $) will be accepted.

At this point, if a coalition S whishes to join the market, it has to choose a contract value that fulfill the two conditions. For simplification, we consider that coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value that obeys the reliability rule (a value we denote by $ P_{\phi}(S) $). If $ P_{\phi}(S) \geq P_{\phi}^{MIN} $, meaning that it also obeys the minimum contract value rule, then S annonces this value on the market : $ P_{S}^{CRCT} = P_{\phi}(S) $. Otherwise, coalition S is not able to enter.
Basically, a coalition S is valid if and only if :
\[ \left\{ \begin{array}{lll}
		\forall t,\ Pr[ P_{S}(t) \leq P_{\phi}(S)] \leq \phi\ \textit{{\scriptsize (reliability rule)}} \\
		and\ P_{\phi}(S) \geq P_{\phi}^{MIN}\ \textit{{\scriptsize (min value rule)}}

\end{array} \right. \]

We now choose a very simple utility function that derives directly from the above remarks. If a coalition cannot provide a valid contract value, it receives naturally a utility of zero. Furthermore, it seems obvious that the utility should increase with the contract value. The $ 1/|S| $ term indicates that we favorise small coalitions, mainly because they are easier to maintain in term of communication.

\[ \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} \]

Obviously, maximising this utility function amounts to maximizing the coalition contract value with the minimum possible number of agents.

In order to illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with gaussian value probability distributions $ \mathcal{P}_{i} = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{P}_{j} = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{P}_{ij} $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :

\[ \left\{ \begin{array}{lll}
		\mu_{ij} = \mu_{i} + \mu_{j} \\
		\sigma_{ij} = \sqrt{\sigma_{i}^{2} + \sigma_{j}^{2} + \rho_{ij}\sigma_{i}\sigma_{j}}
\end{array} \right. \]

with $ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the reliability condition $ Pr[P_{ij}(t) \leq P_{\phi}(ij) ] \leq \phi $ as :

\[ \dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P_{\phi}(ij) - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] \leq \phi \]

The strategy of $ \{ij\} $ consists in maximizing its contract values as long as it respects this inequality, e.g to annonce $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $. It thus appears (as it was intuitively understandable) that, for equivalent sizes, coalitions with low relative standard deviations ( $ \sigma_{ij} / \mu_{ij} $ ) are able to announce higher contract values.

What this paper investigates in the following is the developement of an algorithm that organizes prosumers such that the synergy term of the standard deviation ( $ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) is minimized. In such settings, and through the clique percolation procedure (section 4), we will show that coalitions with low relative standard deviation, e.g high utility coalitions, can be computed.

\section{Correlation Graphs}

When dealing with numerous timeseries, a common objectiv consists in finding clusters whose elements exhibit high similarities while elements in different clusters are very dissimilar. This problem is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering. In [] [], financial stocks are clustered according to their daily log return timeseries. Similarity measures based on Pearson correlation coefficient are established ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})}\ or\ d_{ij} = 1 - \rho_{ij}^{2} $ ) and a complete graph $ G = (V,E,\omega) $ is constructed such that a vertex stands for a stock and the weight of the edge connecting i and j is equal to $ d_{ij} $. In this complete state, the graph G is of little use, but several techniques for filtering information have been proposed such as minimum spanning tree [], k-nearest graphs [], or $\epsilon$-graphs [][].

The idea of $\epsilon$-graphs consists in filtering edges based on their weight, meaning only keeping edges such that their weight is less than $ \epsilon $. Despite the procedure simplicity, the choice of $ \epsilon $ (or k for k-nearest graphs) is not trivial and may influence strongly the results. There is indeed to our knowledge no optimal rule for chosing $ \epsilon $. In [], the authors studied the topological properties (mainly average clustering and connectivity) of the correlation graph depending on the threshold $ \epsilon $. One of the conclusions was that "strong links" (i.e links between strongly correlated timeseries) are responsible for clustering while "weak links" provide network connectivity (in such a network, weak links are indeed connecting densely connected communities).

Applying this technique to the production timeseries of our prosumers (after seasonal effects had been removed) leads to well defined clusters. Indeed, depending on where they are (work, home...), groups of people consume differently but similarly with others of the same group. In the same maner, groups of DER that are subject to similar meteorological conditions produce accordingly. Nevertheless, these clusters of strongly correlated prosumers are the exact opposite of what we are seeking. We thus opt for reversing the metric ($ d_{ij} = \rho_{ij}^{2} $) such that "strong links" become links between uncorrelated timeseries and "weak links", between (anti)correlated timeseries. As expected, independently of the $ \epsilon $ parameter, the graph exhibits henceforth much less clustering and communities are hardly visible. Therefore, using classical clustering or community detection algorithms do not provide very good results (little better than a random algorithm from the utility point of view). However, as it is intuitively understandable, cliques of this graph exhibit very good utility values. But, as some of them may be too small to reach the grid minimum value requirement or because some of them may benefit from additional agents even if they don't form a clique anymore, we used a slightly modified clique percolation algorithm in order to expand them while still controling their expansion. Next section provide some details on clique percolation.

\section{Clique Percolation}

Communities in networks are often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities [1]. The clique percolation algorithm uses directly this observation and the idea behind it is actually quite simple [2] [3]. It starts indeed by searching for cliques (a complete subgraph) in the given graph and considers them as potential seeds for the different communities.

The next ingredient the algorithm needs in order to make the seeds grow is a fitness function that gives some value to a group of nodes. In community detection, it is of common use to employ topological functions that compare internal and external degree, but other functions can also be well suited depending on the problem. Seeds will then alternatively look in their respective neighborhood, select the node that yields the best increase in fitness, and finally incorporate it within the seed. The algorithm stops when seeds cease to grow or when all nodes are affected to at least one seed.
At this point, some coalitions may be partially overlapping, which can be of great use especially in social networks where persons often belongs to several communities (family, friends, work colleagues...). For simplicity, in this paper, we wish to keep the coalitions separated and  leave the management of overlapping coalitions for future work. We thus implemented a simple heuristic that consider nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss. More formally, for a coalition S and a node $ i \in S $, we define :

\[ \tau_{i}(S) = \dfrac{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) - \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S-\{i\}) }{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S)} \]

If node i belongs to  multiple coalitions, the only coalition retaining node i is the one that maximizes $ \tau_{i} $.

We now test our implementation of the clique percolation against well-known benchmark networks in community detection, namely the Zachary karate club, and the UK faculty network. Results can be seen in figure 1.

Left graph is the Zachary karate club graph, shapes represents real communities and color the algorithm output. We can see that only one node was misclassified. Right graph is the UK faculty graph where the algorithm is able to split the graph in three distinct communities.

\section{Results}

We now test the algorithm on a 200 prosumers example. Simulations were run from february 2006 to december 2010 such that we are dealing with 200 hourly sampled timeseries of available power over this period. The first step consists in removing seasonal effects attributable to wind, sun, and temperature natural patterns. We then construct the complete (un)correlation graph as explained in section 3.

At this point, we have four degrees of freedom, namely the reliability ($ \phi $), the required power to enter the market ($P_{\phi}^{MIN}$), the prunning parameter ($\epsilon$) in order to obtain the $\epsilon$-graph, and the number of desired coalitions ($ N_{COAL} $). A first remark would be that $ \epsilon $ has a strong impact on the number of possible coalitions as it is directly responsible for the number and sizes of the cliques in the $ \epsilon$-graph. If $\epsilon $ is too low, the $\epsilon$-graph will not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information and cliques are able to percolate without topological constraints, leading to completely overlapping giant coalitions. Besides, high values on $\epsilon $ give rise to higher computational time as more complete graphs are considered. For $ N_{COAL}=n$, we thus choose $\epsilon $ as the smallest threshold such that the resulting graph contains at least n non overlaping cliques.

The first two parameters ($\phi$ and $P_{\phi}^{MIN}$) shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P_{\phi}^{MIN}$ could lead to valid coalitions (and positive utility values). Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P_{\phi}^{MIN}$ (see figure 1, left plot). In the following, we fix the reliability to a given empirical value and observe how the coalitions evolves for different values of $P_{\phi}^{MIN}$.

As is visible on the right plot of figure 1, when grid constraints are fixed and only the number of coalitions varies, we can see an increase in social welfare (sum of coalitions utilities) up to a maximum point before a decrease. The reason is that increasing the number of coalitions allow, with our algorithm, more coalitions to be stable and enter the market, but there is a point where nodes bringing stability are not sufficient in coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility. Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to graphs where information is flooded, meaning that the algorithm performances decrease.
Naturally, it is also visible that small values of $P_{\phi}^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.

Now that the behavior of the utility is more clear, we have a look at how the algorithm performs. As a comparison, we use a completely random algorithm that only ask for a number of coalitions and partition the agents in a random fashion. For consistency, we always average the results of this algorithm over 100 realization and the errorbars in the plots stands for the standard deviations of the results.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2014-08-19 Article.tex
update

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Bare Demo of IEEEtran.cls for Conferences}
+\title{Allowing Stable Groups of Prosumers to Enter the Market through Proper Aggregation}
 
@@
 % affiliations
-\author{\IEEEauthorblockN{Michael Shell}
-\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
+\author{\IEEEauthorblockN{Nicolas Gensollen}
+\IEEEauthorblockA{Institut Mines Telecom\\Computer Engineering\\
 Georgia Institute of Technology\\
@@
 \and
-\IEEEauthorblockN{Homer Simpson}
+\IEEEauthorblockN{Vincent Gauthier}
 \IEEEauthorblockA{Twentieth Century Fox\\
@@
 \and
-\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
+\IEEEauthorblockN{Michel Marot\\ and Monique Becker}
 \IEEEauthorblockA{Starfleet Academy\\
@@
 \item Realistic representations (prosumers productions are non independent variables and come (indirectly) from real traces)
-\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section 2)
+\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section 3)
 \item Being (relatively) scalable with the number of prosumers
@@
 
-In more details, and as will be explained in section 2, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
+In more details, and as will be explained in section 3, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
 We used real meteorological traces [13] to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
@@
 
-As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering [16] [17] (see section 3).
+As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering [16] [17] (see section 2).
 We will see that in such graphs, cliques represent structures of interest.
 In our settings, it means that cliques are “good” (from the utility point of view) foundations for high valued coalitions.
-We thereby use a slightly modified clique percolation algorithm (see section 4) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements (see results in section 5).
+We thereby use a slightly modified clique percolation algorithm (see section 4) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
+
+The rest of the paper is organized as follows, section 2 locates our work in the related litterature.
+Section 3 explains the model we used to obtain realistic prosumers available power timeseries as well as some notations.
+Section 4 relates the method for building coalitions based on the agent timeseries.
+Finally, section 5 provides some results.
 
 \section{Related Work}
-Related work text here.
+Traditionally, forming coalitions in a pool of agents can be done either in a centralized way where a single central unit is responsible for all the computations or in a distributed way where agents have only local knowledges and take actions accordingly.
+There is a very nice and active smart grid litterature that covers the formation of coalitions trying to optimize various quantities (energy losses, communication bandwidth, cascade of failure sizes...).
+A common point being that a given quantity, usually transcribed by a utility function, should be optimized.
+In this paper, we restrict agents such that they do not take any actions in the coalition formation process.
+The focus is indeed not on the mechanism that drives the agents to an optimum structure, but rather on finding stable groups able to enter the power market for different levels of requirement.
+We will see in section 3 and 4 that, during the formation of coalitions, the agents are completely interchangeable with the timeseries representing their available production.
+
+The setting is thus similar with some fiancial studies on stock exchanges, where researchers tried to find relevant clusters of stocks based on their daily prices variations.
+The problem of clustering is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering.
+Nevertheless, in [] [], stocks are represented by the timeseries of their daily log returns.
+Quite intuitive similarity measures based on Pearson correlation coefficient are established ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})}\ or\ d_{ij} = 1 - \rho_{ij}^{2} $ ) such that correleted stocks are close to each other.
+Such a situation can be represented by a complete graph $ G = (V,E,\omega) $ such that a vertex stands for a stock and the weight of the edge connecting two stocks i and j is equal to $ d_{ij} $.
+However, in this complete state, the graph G is of little use as precious information is flooded.
+Several filtering techniques have been proposed, among which, minimum spanning tree [], k-nearest graphs [], or $\epsilon$-graphs [][].
+
+The idea of $\epsilon$-graphs consists in filtering edges based on their weight, only keeping edges such that their weight is less than $ \epsilon $.
+Despite the procedure simplicity, the choice of $ \epsilon $ (or k for k-nearest graphs) is not trivial and may influence strongly the results.
+There is indeed to our knowledge no optimal rule for chosing $ \epsilon $.
+In [], the authors studied the topological properties (mainly average clustering and connectivity) of the correlation graph depending on the threshold $ \epsilon $.
+One of the conclusions was that "strong links" (i.e links between strongly correlated timeseries) are responsible for clustering while "weak links" provide network connectivity (in such a network, weak links are indeed connecting densely connected communities).
+
+Presented in this way, the timeseries clustering task is also very close to graph community detection.
+Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities [1].
+Although several techniques exists (based on different graph properties : spectrum of Laplacian, modularity, edge betweeness...), the clique percolation algorithm uses directly this observation and the idea behind it is actually quite simple [2] [3].
+It starts indeed by searching for cliques (a complete subgraph) in the given graph and considers them as potential seeds for the different communities.
+
+The next ingredient the algorithm needs in order to make the seeds grow is a fitness function that gives some value to a group of nodes.
+In community detection, it is of common use to employ topological functions that compare internal and external degree, but other functions can also be well suited depending on the problem.
+Seeds will then alternatively look in their respective neighborhood, select the node that yields the best increase in fitness, and finally incorporate it within the seed.
+The algorithm stops when seeds cease to grow or when all nodes are affected to at least one seed.
+At this point, some coalitions may be partially overlapping, which can be of great use especially in social networks where persons often belongs to several communities (family, friends, work colleagues...).
 
 \section{Model}
-Smart grid perspectives assume that smart meters populate the network, recording production/consumption values and reporting these to data centers.
+A major concern while designing our model was its ability to translate correctly realistic patterns (consumption and production) as well as realistic geographical correlations between patterns.
+Largely because of these reasons (but also because they are easily, and sometimes freely, available online) we chose to use meteorological traces as inputs.
+In this paper we used namely both french data from 2006 to 2012 sampled every 3 hours (see link), as well as Western USA data over 2010 sampled every hour (see link).
+The weather stations already offer a quite good sampling of a given territory, and, with the growth of personal weather stations constantly updating data bases, the mesh becomes finer and finer.
+As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering traces for these stations.
+In this paper, we consider that prosumers can only produce through wind turbines and photovoltaic panels (PV).
+On the other hand, we built their consumption patterns according to two major cycles :
+\begin{itemize}
+\item Daily cycles : Consumption is low during night, and higher during the day with two picks in the morning and evening. Some noise is added so that prosumers have similar but not identical cycles.
+\item Seasonal cycles : Consumption is higher in the winter because of heating and low in the summer (we did not consider air conditioning in this paper).
+Temperature traces were the principal ingredient for modeling these cycles.
+\end{itemize}
+
+\begin{center}
+\includegraphics[scale=0.45]{figure2.jpg}
+\end{center}
+
+We thus collect for all chosen weather stations three kind of traces (average wind speed, cloudiness, and average temperature), which we will call a climate vector in the following.
+We consider that cimate vectors are constant over their area, meaning that, if two agents are on the same area, they are exposed to the same climate vector.
+
+At this point, modeling agents consists in fixing a few parameters (most of the time drawn from random distributions) such as the geographical position, the number of wind turbines, PV, appliances, the temperature of confort, and so on...
+The objective is that, depending on the weather of her zone, the DER and appliances she owns, and the way she decides to heat her home, a given prosumer is able to compute her production/consumption at any time.
+We thus used simple and known power curve models for linking weather variables to output power, such that a wind turbine (or a PV for instance) takes a wind speed (a solar irradiance) as input and gives an instantaneous output power [] [].
+Any agent is thus able to compute at any time (within the traces bounds) how much she produces and consumes.
+
+%Smart grid perspectives assume that smart meters populate the network, recording production/consumption values and reporting these to data centers.
 Any entities (DER, prosumers, microgrids, VPP and so on) would thereby be characterizable through timeseries expressing how much these entities produce or consume.
 
-Through our model, we wish thus to be able to generate, for any prosumer, its available production over time, which depend both on its pure production and consumption.
+%Through our model, we wish thus to be able to generate, for any prosumer, its available production over time, which depend both on its pure production and consumption.
 Naturally, these are far from random in the sense that there are recurring patterns (consumption or expected solar irradience for instance) or statistical properties (see wind speed studies) that shape these profiles.
@@
 
-Basically, the user only has to indicate the number of prosumers and the simulation dates.
+%Basically, the user only has to indicate the number of prosumers and the simulation dates.
 Prosumers are then positioned randomly on a squared lattice.
@@
 
-We denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value of agent i at instant t (its instantaneous production minus consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t).
-During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval, we denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting timeserie of agent i.
+More formally, we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value of agent i at instant t (its instantaneous production minus consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t).
+During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
+As expected (because we introduced them), the timeseries exhibit high seasonal patterns, but completely different from agent to agent because they depend both on the energetic mix and habits of a given prosumer.
+Nevertheless, these macro variations can hide the interesting information in the correlation coefficients and twist the rest of our process.
+Thereby, we remove these seasonal effects for all agents (see second block of figure 1).
+We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting timeserie for  agent i.
 For any agent i, we note $ \mathcal{P}_{i} $ the probability distribution drawn from $ \mathcal{T}_{i} $ and by $ P_{i} $ a random variable following $ \mathcal{P}_{i} $.
@@
 \item the reliability (denoted by $ \phi \in [0,1] $)
-\item the minimum contract value under a given reliability ( $ P_{\phi}^{MIN} $). 
+\item the minimum contract value ( $ P_{\phi}^{MIN} $). 
 \end{itemize}
@@
 
-When dealing with numerous timeseries, a common objectiv consists in finding clusters whose elements exhibit high similarities while elements in different clusters are very dissimilar.
-This problem is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering.
-In [] [], financial stocks are clustered according to their daily log return timeseries.
-Similarity measures based on Pearson correlation coefficient are established ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})}\ or\ d_{ij} = 1 - \rho_{ij}^{2} $ ) and a complete graph $ G = (V,E,\omega) $ is constructed such that a vertex stands for a stock and the weight of the edge connecting i and j is equal to $ d_{ij} $.
-In this complete state, the graph G is of little use, but several techniques for filtering information have been proposed such as minimum spanning tree [], k-nearest graphs [], or $\epsilon$-graphs [][].
-
-The idea of $\epsilon$-graphs consists in filtering edges based on their weight, meaning only keeping edges such that their weight is less than $ \epsilon $.
-Despite the procedure simplicity, the choice of $ \epsilon $ (or k for k-nearest graphs) is not trivial and may influence strongly the results.
-There is indeed to our knowledge no optimal rule for chosing $ \epsilon $.
-In [], the authors studied the topological properties (mainly average clustering and connectivity) of the correlation graph depending on the threshold $ \epsilon $.
-One of the conclusions was that "strong links" (i.e links between strongly correlated timeseries) are responsible for clustering while "weak links" provide network connectivity (in such a network, weak links are indeed connecting densely connected communities).
-
 Applying this technique to the production timeseries of our prosumers (after seasonal effects had been removed) leads to well defined clusters.
@@
 
-Communities in networks are often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities [1].
-The clique percolation algorithm uses directly this observation and the idea behind it is actually quite simple [2] [3].
-It starts indeed by searching for cliques (a complete subgraph) in the given graph and considers them as potential seeds for the different communities.
-
-The next ingredient the algorithm needs in order to make the seeds grow is a fitness function that gives some value to a group of nodes.
-In community detection, it is of common use to employ topological functions that compare internal and external degree, but other functions can also be well suited depending on the problem.
-Seeds will then alternatively look in their respective neighborhood, select the node that yields the best increase in fitness, and finally incorporate it within the seed.
-The algorithm stops when seeds cease to grow or when all nodes are affected to at least one seed.
-At this point, some coalitions may be partially overlapping, which can be of great use especially in social networks where persons often belongs to several communities (family, friends, work colleagues...).
  For simplicity, in this paper, we wish to keep the coalitions separated and  leave the management of overlapping coalitions for future work.
@@
 
+\begin{center}
+\begin{tabular}{cc}
+   \includegraphics[scale=0.44]{./figure3/utility.eps} &
+   \includegraphics[scale=0.44]{./figure3/utility.eps} \\
+\end{tabular}
+\end{center}
+
 At this point, we have four degrees of freedom, namely the reliability ($ \phi $), the required power to enter the market ($P_{\phi}^{MIN}$), the prunning parameter ($\epsilon$) in order to obtain the $\epsilon$-graph, and the number of desired coalitions ($ N_{COAL} $).
@@
 Naturally, it is also visible that small values of $P_{\phi}^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.
+
+\begin{center}
+   \includegraphics[scale=0.7]{./figure4/survival.eps}
+\end{center}
 

### 2014-08-20 Article.tex
update

@@
 As it seems intuitively understandable that maintaining stable coalitions would be less expensive in terms of communication load than highly versatible ones, our objective consists in forming coalitions that satisfy grid requirements both in terms of stability and minimum production.
-Our definition of stability being implicitely linked to correlation between agents timeseries (see section 2), we organize them over a correlation graph and use a variation of clique percolation to construct valid coalitions in reasonable time.
+Our definition of stability being implicitely linked to correlation between agents timeseries (see section 3), we organize them over a correlation graph and use a variation of clique percolation to construct valid coalitions in reasonable time.
 \end{abstract}
@@
 In our settings, it means that cliques are “good” (from the utility point of view) foundations for high valued coalitions.
-We thereby use a slightly modified clique percolation algorithm (see section 4) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
+We thereby use a modified clique percolation algorithm (see section 4) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
 
 The rest of the paper is organized as follows, section 2 locates our work in the related litterature.
-Section 3 explains the model we used to obtain realistic prosumers available power timeseries as well as some notations.
+Section 3 explains the model we used to obtain realistic prosumers available power timeseries and introduces some notations.
 Section 4 relates the method for building coalitions based on the agent timeseries.
@@
 
-The idea of $\epsilon$-graphs consists in filtering edges based on their weight, only keeping edges such that their weight is less than $ \epsilon $.
+The idea of $\epsilon$-graphs consists in filtering edges based on their weight, only keeping edges whose weights are less than $ \epsilon $.
 Despite the procedure simplicity, the choice of $ \epsilon $ (or k for k-nearest graphs) is not trivial and may influence strongly the results.
@@
 \begin{itemize}
-\item Daily cycles : Consumption is low during night, and higher during the day with two picks in the morning and evening. Some noise is added so that prosumers have similar but not identical cycles.
-\item Seasonal cycles : Consumption is higher in the winter because of heating and low in the summer (we did not consider air conditioning in this paper).
+\item \textbf{Daily cycles} : Consumption is low during night, and higher during the day with two picks in the morning and evening.
+Some noise is added so that prosumers have similar but not identical cycles.
+\item \textbf{Seasonal cycles} : Consumption is higher in the winter because of heating and low in the summer (we did not consider air conditioning in this paper).
 Temperature traces were the principal ingredient for modeling these cycles.
@@
 \begin{center}
+\begin{figure}
 \includegraphics[scale=0.45]{figure2.jpg}
+\caption{Process diagram}
+\end{figure}
 \end{center}
@@
 We thus collect for all chosen weather stations three kind of traces (average wind speed, cloudiness, and average temperature), which we will call a climate vector in the following.
-We consider that cimate vectors are constant over their area, meaning that, if two agents are on the same area, they are exposed to the same climate vector.
+We consider that cimate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector.
 
@@
 \begin{itemize}
-\item the reliability (denoted by $ \phi \in [0,1] $)
-\item the minimum contract value ( $ P_{\phi}^{MIN} $). 
+\item the \textbf{reliability} (denoted by $ \phi \in [0,1] $)
+\item the \textbf{minimum contract value} ( $ P_{\phi}^{MIN} $). 
 \end{itemize}
@@
 
-\section{Correlation Graphs}
+\section{Forming coalitions}
 
-Applying this technique to the production timeseries of our prosumers (after seasonal effects had been removed) leads to well defined clusters.
-Indeed, depending on where they are (work, home...), groups of people consume differently but similarly with others of the same group.
-In the same maner, groups of DER that are subject to similar meteorological conditions produce accordingly.
-Nevertheless, these clusters of strongly correlated prosumers are the exact opposite of what we are seeking.
-We thus opt for reversing the metric ($ d_{ij} = \rho_{ij}^{2} $) such that "strong links" become links between uncorrelated timeseries and "weak links", between (anti)correlated timeseries.
-As expected, independently of the $ \epsilon $ parameter, the graph exhibits henceforth much less clustering and communities are hardly visible.
-Therefore, using classical clustering or community detection algorithms do not provide very good results (little better than a random algorithm from the utility point of view).
-However, as it is intuitively understandable, cliques of this graph exhibit very good utility values.
-But, as some of them may be too small to reach the grid minimum value requirement or because some of them may benefit from additional agents even if they don't form a clique anymore, we used a slightly modified clique percolation algorithm in order to expand them while still controling their expansion.
-Next section provide some details on clique percolation.
+This section explains the process with which we form the coalitions (see the third block of Fig.
+1).
+First, we need to simulate the timeseries of available power (first two blocks of Fig.
+1).
+We consider a pool $ \mathcal{A} $ of 200 agents, whose parameters were chosen randomly.
+The prosumers are positioned (also randomly) on a square lattice previously filled with climate vectors obtained from the french data sets (see section 3).
+Simulations were run from february 2006 to december 2010 such that we are dealing with 200 hourly sampled timeseries of available power over this period.
+Removing season trends finally leads us to the formation of coalitions.
 
-\section{Clique Percolation}
+The model we used in order to simulate timeseries of available production provide some diversity because of the energetic mix/climate vectors combo.
+Nevertheless, as the number of agents grows, the timeseries tend to exhibit similar patterns.
+This is particularly visible when creating a correlation graph $ G_{1}(\mathcal{A},E_{1},\omega_{1}) $ with the same kind of metric as [] or [] ($ d_{ij} = 1 - \rho_{ij}^{2} $).
+Very well defined clusters appear in the corresponding $ \epsilon $-graph for large range of $ \epsilon $.
 
+However, these clusters of strongly correlated timeseries are the exact opposite of what we are seeking.
+We can indeed consider them directly as coalitions and compute their utilities, and the results show, as expected, terrible values (far worse than a random split of the agents in the same number of coalitions).
+
+We thus opt for reversing the metric ($ d_{ij}^{'} = \rho_{ij}^{2} $) such that uncorrelated timeseries are close to each other and (anti)correlated timeseries are distant in $ G_{2}(\mathcal{A},E_{2},\omega_{2}) $.
+As expected [], independently of the $ \epsilon $ parameter, the resulting $ \epsilon $-graph exhibits henceforth much less clustering than in $ G_{1} $ and communities seem hardly visible.
+Therefore, using classical clustering or community detection algorithms do not provide very good results.
+
+However, as seems intuitively understandable, cliques of this graph tend to exhibit very good utility values.
+Such  structures contain ideed a link between every two nodes, meaning that the overall correlation is quite small.
+Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques, if it is too low, the $\epsilon$-graph of $ G_{2} $ will not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly.
+For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even becomes computationally repulsive.
+Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ seems too brutal.
+Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(S) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore.
+
+Cliques for small values of $ \epsilon $ appear thus as good seeds for stable coalitions, but we still have to find a way to bring them above the grid requirements as well as increasing the social welfare, both by incorporating more nodes.
+This is exactly where clique percolation proves useful for our concerns.
+The main idea of making the seeds grow as long as it increases a fitness function is exactly what we need.
+The only difference being that our fitness function is simply chosen as the utility function.
+
+As explained in section 2, clique percolation leads generally to overlapping communities.
 For simplicity, in this paper, we wish to keep the coalitions separated and  leave the management of overlapping coalitions for future work.
@@
 
-We now test our implementation of the clique percolation against well-known benchmark networks in community detection, namely the Zachary karate club, and the UK faculty network.
-Results can be seen in figure 1.
+At this point, we have three degrees of freedom, namely the reliability ($ \phi $), the required power to enter the market ($P_{\phi}^{MIN}$), and the prunning parameter ($\epsilon$).
+For conveniance, we introduce the number of desired coalitions ($ N_{COAL} $) and we fix $\epsilon $ as the smallest value such that the resulting $ \epsilon $-graph contains at least $ N_{COAL} $ non overlaping cliques.
+The overall utility depends now on the grid requirements and on the number of desired coalitions.
 
-Left graph is the Zachary karate club graph, shapes represents real communities and color the algorithm output.
-We can see that only one node was misclassified.
-Right graph is the UK faculty graph where the algorithm is able to split the graph in three distinct communities.
+The next section shows how the utility behave within the parameters space and provides the results of our 200 agents test.
 
@@
 
-We now test the algorithm on a 200 prosumers example.
-Simulations were run from february 2006 to december 2010 such that we are dealing with 200 hourly sampled timeseries of available power over this period.
-The first step consists in removing seasonal effects attributable to wind, sun, and temperature natural patterns.
-We then construct the complete (un)correlation graph as explained in section 3.
-
 \begin{center}
+\begin{figure}
 \begin{tabular}{cc}
@@
 \end{tabular}
+\caption{Utility function in parameters space}
+\end{figure}
 \end{center}
-
-At this point, we have four degrees of freedom, namely the reliability ($ \phi $), the required power to enter the market ($P_{\phi}^{MIN}$), the prunning parameter ($\epsilon$) in order to obtain the $\epsilon$-graph, and the number of desired coalitions ($ N_{COAL} $).
-A first remark would be that $ \epsilon $ has a strong impact on the number of possible coalitions as it is directly responsible for the number and sizes of the cliques in the $ \epsilon$-graph.
-If $\epsilon $ is too low, the $\epsilon$-graph will not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information and cliques are able to percolate without topological constraints, leading to completely overlapping giant coalitions.
-Besides, high values on $\epsilon $ give rise to higher computational time as more complete graphs are considered.
-For $ N_{COAL}=n$, we thus choose $\epsilon $ as the smallest threshold such that the resulting graph contains at least n non overlaping cliques.
 
@@
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P_{\phi}^{MIN}$ (see figure 1, left plot).
+
 In the following, we fix the reliability to a given empirical value and observe how the coalitions evolves for different values of $P_{\phi}^{MIN}$.
+Note that the opposite is also possible although a little less intuitive, but not shown here because it leads to similar conclusions.
 
 As is visible on the right plot of figure 1, when grid constraints are fixed and only the number of coalitions varies, we can see an increase in social welfare (sum of coalitions utilities) up to a maximum point before a decrease.
-The reason is that increasing the number of coalitions allow, with our algorithm, more coalitions to be stable and enter the market, but there is a point where nodes bringing stability are not sufficient in coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility.
+The reason is that increasing the number of coalitions allows more coalitions to reach stability and enter the market, but there is a point where nodes bringing stability are not sufficient inside the coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility.
 Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to graphs where information is flooded, meaning that the algorithm performances decrease.
@@
 \begin{center}
-   \includegraphics[scale=0.7]{./figure4/survival.eps}
+\begin{figure}
+  \includegraphics[scale=0.8]{./figure5/coals.pdf}
+  \caption{Evolution of the coalitions for different values of grid requirement $ P_{MIN} $. Blue dots represents clique percolation results while red dots stands for random process results.}
+\end{figure}
 \end{center}
 
-Now that the behavior of the utility is more clear, we have a look at how the algorithm performs.
-As a comparison, we use a completely random algorithm that only ask for a number of coalitions and partition the agents in a random fashion.
-For consistency, we always average the results of this algorithm over 100 realization and the errorbars in the plots stands for the standard deviations of the results.
+For our 200 prosumers, we suppose that the grid fixes a reliability $ \phi = 0.1 $ (meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time.
+We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P_{MIN} $ requirement.
+As a comparison, we use a completely random algorithm that only asks for a number of coalitions and partitions the agents in a random fashion.
+Figure 3 shows this evolution for our algorithm (blue dots) and for the random process (red dots).
+The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility.
+The $ P_{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.
+
+Looking at figure 3, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements.
+The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither to low or to high ($ P_{MIN} = 5 $ in figure 3).
+
+In order to explore a little bit more in this direction, we have a look, on figure 4, at the percentage of coalitions able to enter the market for different values of $ P_{MIN} $.
+For consistency, we always average the results of the random procedure over 100 realizations (5 realizations for our algorithm, but the results are relatively stable), and the errorbars in the plots stands for the standard deviations of the results.
+
+When the grid requirements are very low ( $ P_{MIN} $ close to zero), all coalitions for both algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
+But as $ P_{MIN} $ increases, we see the percentage of the random procedure quickly dropping while it stays constant for our algorithm.
+For $ P_{MIN} = 4 $, we see that only a little more than half of the coalitions for the random case are able to enter while all of them enters for the clique percolation algorithm.
+Finally, when the grid requirement becomes to high, the acceptance percentage of our algorithm tends slowly to zero (only partly shown in this plot for readability).
+
+\begin{center}
+\begin{figure}
+   \includegraphics[scale=0.7]{./figure4/survival.eps}
+   \caption{Percentage of coalitions able to enter the market for different values of grid requirement $ P_{MIN} $.
+Blue curve represents clique percolation results while red curve stands for random process results}
+\end{figure}
+\end{center}
 

### 2014-08-21 Article.tex
update

@@
 \author{\IEEEauthorblockN{Nicolas Gensollen}
-\IEEEauthorblockA{Institut Mines Telecom\\Computer Engineering\\
-Georgia Institute of Technology\\
-Atlanta, Georgia 30332--0250\\
+\IEEEauthorblockA{Institut Mines Telecom\\
 Email: http://www.michaelshell.org/contact.html}
@@
 Together with electric vehicles, and demand side management tools, they constitute the building blocks which will help turn the today pure energy consumers into true actors of the grid operation [3].
-Such energy aware agents that consume, produce, and are ready to make concessions (appliances delays, V2G…) to ensure grid stability are commonly called prosumers [4] [5].
+Such energy aware agents that consume, produce, and are ready to make concessions (appliances delays, V2G…) to ensure grid stability are commonly called prosumers \cite{Rathnayaka2012}.
 
@@
 \section{Related Work}
+
 Traditionally, forming coalitions in a pool of agents can be done either in a centralized way where a single central unit is responsible for all the computations or in a distributed way where agents have only local knowledges and take actions accordingly.
-There is a very nice and active smart grid litterature that covers the formation of coalitions trying to optimize various quantities (energy losses, communication bandwidth, cascade of failure sizes...).
-A common point being that a given quantity, usually transcribed by a utility function, should be optimized.
-In this paper, we restrict agents such that they do not take any actions in the coalition formation process.
-The focus is indeed not on the mechanism that drives the agents to an optimum structure, but rather on finding stable groups able to enter the power market for different levels of requirement.
-We will see in section 3 and 4 that, during the formation of coalitions, the agents are completely interchangeable with the timeseries representing their available production.
+It is of common use to represent the situation and assess the stability of the solution by using game theoretic tools.
+Some papers focus on finding an optimal coalition structure giving a pool of autonomous self interested agents.
+In this direction, \cite{Saad2009} and \cite{Luan2014} design a distributed merge and split algorithm based on modified Pareto order with utility functions aiming at optimizing a parameter of interest such as energy losses or communication costs.
+
+In \cite{Kota2011}, the authors study the formation of virtual power plants (VPP) composed of multiple self-interested DER.
+On the grid side, two requirements for the formation of virtual power plants are derived, namely the reliability of supply and the minimization of entities the grid has to deal with.
+From this, \cite{Kota2011} builds a pricing mechanism that encourages VPP to report true estimates of their aggregated production and penalizes prediction errors.
+A redistribution scheme of the VPP to the DER is also constructed such that the payoff allocation lies in the core of the game, meaning that no DER has an incentive to leave the coalition.
+
+In this paper and as will be explained in section 3, coalitions utilities depend on statistical properties of historical values.
+The goal being to form coalitions that, statistically spealing, are more likely to exhibit stable behaviors.
+We will see that, when it comes to the formation of the coalitions, a prosumer is completely interchangeable with the timeserie representing its available production.
 
@@
 The problem of clustering is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering.
-Nevertheless, in [] [], stocks are represented by the timeseries of their daily log returns.
+Nevertheless, in \cite{Mantegna1999} \cite{Garas2008}, stocks are represented by the timeseries of their daily log returns.
 Quite intuitive similarity measures based on Pearson correlation coefficient are established ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})}\ or\ d_{ij} = 1 - \rho_{ij}^{2} $ ) such that correleted stocks are close to each other.
@@
 However, in this complete state, the graph G is of little use as precious information is flooded.
-Several filtering techniques have been proposed, among which, minimum spanning tree [], k-nearest graphs [], or $\epsilon$-graphs [][].
+Several filtering techniques have been proposed, among which, minimum spanning tree \cite{Mantegna1999}, k-nearest graphs [], or $\epsilon$-graphs [][].
 
@@
 There is indeed to our knowledge no optimal rule for chosing $ \epsilon $.
-In [], the authors studied the topological properties (mainly average clustering and connectivity) of the correlation graph depending on the threshold $ \epsilon $.
+In \cite{Garas2008}, the authors studied the topological properties (mainly average clustering and connectivity) of the correlation graph depending on the threshold $ \epsilon $.
 One of the conclusions was that "strong links" (i.e links between strongly correlated timeseries) are responsible for clustering while "weak links" provide network connectivity (in such a network, weak links are indeed connecting densely connected communities).
@@
 Presented in this way, the timeseries clustering task is also very close to graph community detection.
-Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities [1].
-Although several techniques exists (based on different graph properties : spectrum of Laplacian, modularity, edge betweeness...), the clique percolation algorithm uses directly this observation and the idea behind it is actually quite simple [2] [3].
-It starts indeed by searching for cliques (a complete subgraph) in the given graph and considers them as potential seeds for the different communities.
+Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities \cite{Newmanb}.
+Although several techniques exists (based on different graph properties : modularity \cite{Newmanb}, edge betweeness \cite{Girvan2002}, spectrum of Laplacian \cite{Newman}...), the clique percolation algorithm \cite{Lancichinetti} uses directly this observation and the idea behind it is actually quite simple.
+It starts indeed by searching for cliques (a complete subgraph) in the given graph as potential community cores.
+Because other nodes can naturally belong to a community with a less strong connectivity requirement, the algorithm uses a fitness function that quantifies the quality (according to some criterions) of a community structure inside the graph.
+Expansed cliques alternatively look in their respective neighborhood, select the node that yields the best increase in fitness, and finally incorporate it.
+The percolation algorithm stops when seeds cease to grow or when all nodes are affected to at least one seed.
+At this point, some coalitions may be partially overlapping, meaning that a node can belong to multiple communities.
+Detection of overlapping communities is actually a very active field of research, especially in social networks where persons often belongs to several communities (family, friends, work colleagues...).
 
-The next ingredient the algorithm needs in order to make the seeds grow is a fitness function that gives some value to a group of nodes.
-In community detection, it is of common use to employ topological functions that compare internal and external degree, but other functions can also be well suited depending on the problem.
-Seeds will then alternatively look in their respective neighborhood, select the node that yields the best increase in fitness, and finally incorporate it within the seed.
-The algorithm stops when seeds cease to grow or when all nodes are affected to at least one seed.
-At this point, some coalitions may be partially overlapping, which can be of great use especially in social networks where persons often belongs to several communities (family, friends, work colleagues...).
+The following section of this paper explains the model we built in order obtain sufficient historical values for our prosumers, such that forming coalitions based on statistics is actually meaningfull.
 
@@
 Largely because of these reasons (but also because they are easily, and sometimes freely, available online) we chose to use meteorological traces as inputs.
-In this paper we used namely both french data from 2006 to 2012 sampled every 3 hours (see link), as well as Western USA data over 2010 sampled every hour (see link).
+In this paper we used namely both french data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data for the United States covering year 2010 can also be found at \cite{NCDC}).
 The weather stations already offer a quite good sampling of a given territory, and, with the growth of personal weather stations constantly updating data bases, the mesh becomes finer and finer.
@@
 \begin{tabular}{cc}
-   \includegraphics[scale=0.44]{./figure3/utility.eps} &
+   \includegraphics[scale=0.44]{./figure1/util.eps} &
    \includegraphics[scale=0.44]{./figure3/utility.eps} \\
@@
 \section{Conclusion}
-The conclusion goes here.
+In this paper we collected meteorological traces for weather stations sampling a given territory.
+After discetization of the space, each zone receives the climate vector corresponding to its mother weather station.
+Prosumers are then distributed randomly between zones and each prosumer receives a random number of DER and loads as well as diverse consumption habits.
+Simulations are then runned and allow us to obtain, for each prosumer, the timeserie describing the instantaneous power that he could potentially be wanted to sell on the market.
+Nevertheless, as entering on the energy market is constrained by the grid operator, such that only sufficiently reliable and sufficiently productive units are allowed to enter, prosumers have usually no other choice than forming coalitions.
+We then introduced a framework such that the grid requirements take a precise meaning and such that coalitions announce contract values accordingly.
+
+Through this paper, our interest was on how we could circumscribe groups of prosumers that can rise above the grid requirements.
+In this direction, we removed seasonal variations from the prosumers timeseries and proposed to organize them over a correlation graph with a reversed metric.
+A greedy clique percolation process on the $ \epsilon $-graph enables us to finally establish coalitions of agents.
+We showed in the last section that our algorithm was actually finding coalitions with high utility values but also that it was far more robust than a random search over coalition structures when the grid operator increases his requirements (Fig.
+4).
+
+An interresting lead for future work would be the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analysed.
+Besides, we believe that not restricting the algorithm to non overlapping coalitions and studying the strategies and weights of nodes with multiple options could lead to interesting works.
 
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
 
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+%\begin{thebibliography}{1}
 
-\end{thebibliography}
+%\bibitem{IEEEhowto:kopka}
+%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
+%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+
+\bibliographystyle{IEEEtran}  
+\bibliography{Article}
+
+%\end{thebibliography}
 

### 2014-08-22 Article.tex
update

@@
 As it seems intuitively understandable that maintaining stable coalitions would be less expensive in terms of communication load than highly versatible ones, our objective consists in forming coalitions that satisfy grid requirements both in terms of stability and minimum production.
-Our definition of stability being implicitely linked to correlation between agents timeseries (see section 3), we organize them over a correlation graph and use a variation of clique percolation to construct valid coalitions in reasonable time.
+Our definition of stability being implicitely linked to correlation between agents timeseries (see section \ref{sec:model}), we organize them over a correlation graph and use a variation of clique percolation to construct valid coalitions in reasonable time.
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:introduction}
 % no \IEEEPARstart
@@
 \item Realistic representations (prosumers productions are non independent variables and come (indirectly) from real traces)
-\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section 3)
+\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section \ref{sec:model})
 \item Being (relatively) scalable with the number of prosumers
@@
 
-In more details, and as will be explained in section 3, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
+In more details, and as will be explained in section \ref{sec:model}, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
 We used real meteorological traces [13] to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
@@
 
-As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering [16] [17] (see section 2).
+As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering [16] [17] (see section \ref{sec:related}).
 We will see that in such graphs, cliques represent structures of interest.
 In our settings, it means that cliques are “good” (from the utility point of view) foundations for high valued coalitions.
-We thereby use a modified clique percolation algorithm (see section 4) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
+We thereby use a modified clique percolation algorithm (see section \ref{sec:results}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
 
-The rest of the paper is organized as follows, section 2 locates our work in the related litterature.
-Section 3 explains the model we used to obtain realistic prosumers available power timeseries and introduces some notations.
-Section 4 relates the method for building coalitions based on the agent timeseries.
-Finally, section 5 provides some results.
+The rest of the paper is organized as follows, section \ref{sec:related} locates our work in the related litterature.
+Section \ref{sec:model} explains the model we used to obtain realistic prosumers available power timeseries and introduces some notations.
+Section \ref{sec:forming} relates the method for building coalitions based on the agent timeseries.
+Finally, section \ref{sec:results} provides some results.
 
 \section{Related Work}
+\label{sec:related}
 
@@
 
-In this paper and as will be explained in section 3, coalitions utilities depend on statistical properties of historical values.
+In this paper and as will be explained in section \ref{sec:model}, coalitions utilities depend on statistical properties of historical values.
 The goal being to form coalitions that, statistically spealing, are more likely to exhibit stable behaviors.
@@
 The problem of clustering is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering.
-Nevertheless, in \cite{Mantegna1999} \cite{Garas2008}, stocks are represented by the timeseries of their daily log returns.
-Quite intuitive similarity measures based on Pearson correlation coefficient are established ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})}\ or\ d_{ij} = 1 - \rho_{ij}^{2} $ ) such that correleted stocks are close to each other.
-Such a situation can be represented by a complete graph $ G = (V,E,\omega) $ such that a vertex stands for a stock and the weight of the edge connecting two stocks i and j is equal to $ d_{ij} $.
-However, in this complete state, the graph G is of little use as precious information is flooded.
-Several filtering techniques have been proposed, among which, minimum spanning tree \cite{Mantegna1999}, k-nearest graphs [], or $\epsilon$-graphs [][].
+Nevertheless, in \cite{Mantegna1999} Mantegna introduced another approach where stocks, represented by the timeseries of their daily log returns, are organized in a graph such that stocks (vertices) exhibiting similar price fluctuation patterns are close to each other.
+This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ but sometimes also $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
+Because there is a distance between every two nodes, \cite{Mantegna1999} obtains a complete graph where precious information is flooded.
+In order to discard unsignificant information, \cite{Mantegna1999} used a minimum spanning tree oriented technique and achieves a hierarchical clustering of the stocks.
+Since then, several other filtering techniques have been proposed, such as k-nearest graphs, or $ \epsilon $-graphs, that relax the tree assumption \cite{Onnela2004}.
 
@@
 There is indeed to our knowledge no optimal rule for chosing $ \epsilon $.
-In \cite{Garas2008}, the authors studied the topological properties (mainly average clustering and connectivity) of the correlation graph depending on the threshold $ \epsilon $.
-One of the conclusions was that "strong links" (i.e links between strongly correlated timeseries) are responsible for clustering while "weak links" provide network connectivity (in such a network, weak links are indeed connecting densely connected communities).
+In \cite{Garas2008} and \cite{Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques...) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
+One of the conclusions was that "strong links" (i.e links between strongly correlated timeseries) are responsible for clustering while "weak links" provide network connectivity.
 
-Presented in this way, the timeseries clustering task is also very close to graph community detection.
+Presented in this way, the timeseries clustering task seems very close to graph community detection.
 Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities \cite{Newmanb}.
@@
 \section{Model}
+\label{sec:model}
+
 A major concern while designing our model was its ability to translate correctly realistic patterns (consumption and production) as well as realistic geographical correlations between patterns.
@@
 \caption{Process diagram}
+\label{Fig1}
 \end{figure}
@@
 The objective is that, depending on the weather of her zone, the DER and appliances she owns, and the way she decides to heat her home, a given prosumer is able to compute her production/consumption at any time.
-We thus used simple and known power curve models for linking weather variables to output power, such that a wind turbine (or a PV for instance) takes a wind speed (a solar irradiance) as input and gives an instantaneous output power [] [].
+We thus used simple and known power curve models for linking weather variables to output power, such that a wind turbine (or a PV for instance) takes a wind speed (a solar irradiance) as input and gives an instantaneous output power \cite{Kota2011} \cite{windturbinemodel}.
 Any agent is thus able to compute at any time (within the traces bounds) how much she produces and consumes.
@@
 Nevertheless, these macro variations can hide the interesting information in the correlation coefficients and twist the rest of our process.
-Thereby, we remove these seasonal effects for all agents (see second block of figure 1).
+Thereby, we remove these seasonal effects for all agents (see second block of figure \ref{Fig1}).
 We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting timeserie for  agent i.
@@
 What this paper investigates in the following is the developement of an algorithm that organizes prosumers such that the synergy term of the standard deviation ( $ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) is minimized.
-In such settings, and through the clique percolation procedure (section 4), we will show that coalitions with low relative standard deviation, e.g high utility coalitions, can be computed.
+In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with low relative standard deviation, e.g high utility coalitions, can be computed.
 
 \section{Forming coalitions}
+\label{sec:forming}
 
-This section explains the process with which we form the coalitions (see the third block of Fig.
-1).
-First, we need to simulate the timeseries of available power (first two blocks of Fig.
-1).
+This section explains the process with which we form the coalitions (see the third block of Figure \ref{Fig1}).
+First, we need to simulate the timeseries of available power (first two blocks of Figure \ref{Fig1}).
 We consider a pool $ \mathcal{A} $ of 200 agents, whose parameters were chosen randomly.
-The prosumers are positioned (also randomly) on a square lattice previously filled with climate vectors obtained from the french data sets (see section 3).
+The prosumers are positioned (also randomly) on a square lattice previously filled with climate vectors obtained from the french data sets (see section \ref{sec:model}).
 Simulations were run from february 2006 to december 2010 such that we are dealing with 200 hourly sampled timeseries of available power over this period.
@@
 Nevertheless, as the number of agents grows, the timeseries tend to exhibit similar patterns.
-This is particularly visible when creating a correlation graph $ G_{1}(\mathcal{A},E_{1},\omega_{1}) $ with the same kind of metric as [] or [] ($ d_{ij} = 1 - \rho_{ij}^{2} $).
+This is particularly visible when creating a correlation graph $ G_{1}(\mathcal{A},E_{1},\omega_{1}) $ with the same kind of metric as \cite{Garas2008} or \cite{Onnela2004} ($ d_{ij} = 1 - \rho_{ij}^{2} $).
 Very well defined clusters appear in the corresponding $ \epsilon $-graph for large range of $ \epsilon $.
@@
 We thus opt for reversing the metric ($ d_{ij}^{'} = \rho_{ij}^{2} $) such that uncorrelated timeseries are close to each other and (anti)correlated timeseries are distant in $ G_{2}(\mathcal{A},E_{2},\omega_{2}) $.
-As expected [], independently of the $ \epsilon $ parameter, the resulting $ \epsilon $-graph exhibits henceforth much less clustering than in $ G_{1} $ and communities seem hardly visible.
+As expected \cite{Onnela2004}, independently of the $ \epsilon $ parameter, the resulting $ \epsilon $-graph exhibits henceforth much less clustering than in $ G_{1} $ and communities seem hardly visible.
 Therefore, using classical clustering or community detection algorithms do not provide very good results.
@@
 
-As explained in section 2, clique percolation leads generally to overlapping communities.
+As explained in section \ref{sec:related}, clique percolation leads generally to overlapping communities.
 For simplicity, in this paper, we wish to keep the coalitions separated and  leave the management of overlapping coalitions for future work.
@@
 \section{Results}
+\label{sec:results}
 
@@
 \begin{figure}
+\setlength{\tabcolsep}{-0.7em}
 \begin{tabular}{cc}
-   \includegraphics[scale=0.44]{./figure1/util.eps} &
-   \includegraphics[scale=0.44]{./figure3/utility.eps} \\
+   \includegraphics[scale=0.5]{./figure1/util.eps} &
+   \includegraphics[scale=0.5]{./figure6/util2.eps} \\
 \end{tabular}
-\caption{Utility function in parameters space}
+\caption{Utility function in parameters space. Left in $ (\phi, P_{\phi}^{MIN}) $ when $ N_{COAL} $ is fixed. Right in $ (N_{COAL}, P_{\phi}^{MIN}) $ for a fixed $ \phi $}
+\label{Fig2}
 \end{figure}
@@
 
-The first two parameters ($\phi$ and $P_{\phi}^{MIN}$) shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P_{\phi}^{MIN}$ could lead to valid coalitions (and positive utility values).
-Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P_{\phi}^{MIN}$ (see figure 1, left plot).
+As visible on the left plot of figure \ref{Fig2}, the parameters $\phi$ and $P_{\phi}^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P_{\phi}^{MIN}$ could potentially lead to valid coalitions (and positive utility values).
+Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P_{\phi}^{MIN}$.
+Obviously, the highest utility values are found for high $ \phi $, because coalitions are able then to announce higher contract values, yielding higer utilities.
 
@@
 
-As is visible on the right plot of figure 1, when grid constraints are fixed and only the number of coalitions varies, we can see an increase in social welfare (sum of coalitions utilities) up to a maximum point before a decrease.
+On the right plot of figure \ref{Fig2}, we can see that when the grid constraints are fixed ($ \phi = 0.08 $ and $ P_{\phi}^{MIN} = 2 $ for instance) and only the number of coalitions varies, there is an increase in social welfare (sum of coalitions utilities) up to a maximum point (for $ N_{COAL} = 3 $ here) before a decrease.
 The reason is that increasing the number of coalitions allows more coalitions to reach stability and enter the market, but there is a point where nodes bringing stability are not sufficient inside the coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility.
-Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to graphs where information is flooded, meaning that the algorithm performances decrease.
+Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs where information is flooded, meaning that the algorithm performances decrease.
 Naturally, it is also visible that small values of $P_{\phi}^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.
@@
 \begin{center}
-\begin{figure}
+\begin{figure}[b]
   \includegraphics[scale=0.8]{./figure5/coals.pdf}
   \caption{Evolution of the coalitions for different values of grid requirement $ P_{MIN} $. Blue dots represents clique percolation results while red dots stands for random process results.}
+  \label{Fig3}
 \end{figure}
@@
 
-For our 200 prosumers, we suppose that the grid fixes a reliability $ \phi = 0.1 $ (meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time.
+For our 200 prosumers, we suppose that the grid fixes a reliability $ \phi = 0.1 $ (meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time).
 We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P_{MIN} $ requirement.
 As a comparison, we use a completely random algorithm that only asks for a number of coalitions and partitions the agents in a random fashion.
-Figure 3 shows this evolution for our algorithm (blue dots) and for the random process (red dots).
+Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots).
 The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility.
@@
 
-Looking at figure 3, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements.
-The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither to low or to high ($ P_{MIN} = 5 $ in figure 3).
+Looking at figure \ref{Fig3}, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements.
+The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither to low or to high ($ P_{MIN} = 5 $ in figure \ref{Fig3}).
 
-In order to explore a little bit more in this direction, we have a look, on figure 4, at the percentage of coalitions able to enter the market for different values of $ P_{MIN} $.
+In order to explore a little bit more in this direction, we have a look, on figure \ref{Fig4}, at the percentage of coalitions able to enter the market for different values of $ P_{MIN} $.
 For consistency, we always average the results of the random procedure over 100 realizations (5 realizations for our algorithm, but the results are relatively stable), and the errorbars in the plots stands for the standard deviations of the results.
@@
 Blue curve represents clique percolation results while red curve stands for random process results}
+   \label{Fig4}
 \end{figure}
@@
 \section{Conclusion}
+\label{sec: conclusion}
+
 In this paper we collected meteorological traces for weather stations sampling a given territory.
@@
 A greedy clique percolation process on the $ \epsilon $-graph enables us to finally establish coalitions of agents.
-We showed in the last section that our algorithm was actually finding coalitions with high utility values but also that it was far more robust than a random search over coalition structures when the grid operator increases his requirements (Fig.
-4).
+We showed in section \ref{sec:results} that our algorithm was actually finding coalitions with high utility values but also that it was far more robust than a random search over coalition structures when the grid operator increases his requirements (Figure \ref{Fig4}).
 

### 2014-08-26 Article.tex
update

@@
 % no \IEEEPARstart
-One of the most basic ideas in the smart grid revolves around the introduction of communication means in the power grids, which could enable complex improvements in the energy management and lead progressively to a greener energetic system [1] [2].
+One of the most basic ideas in the smart grid revolves around the introduction of communication means in the power grids, which could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
 Distributed energy resources (DER) such as wind turbines or photovoltaic panels are supposed to emerge and populate not only spread out farms, but also classical residential neighborhoods.
-Together with electric vehicles, and demand side management tools, they constitute the building blocks which will help turn the today pure energy consumers into true actors of the grid operation [3].
-Such energy aware agents that consume, produce, and are ready to make concessions (appliances delays, V2G…) to ensure grid stability are commonly called prosumers \cite{Rathnayaka2012}.
+Together with electric vehicles, and demand side management tools, they constitute the building blocks which will help turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
+Such energy aware agents that consume, produce, and are ready to make concessions (appliances delays, V2G…) to ensure grid stability are commonly called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
 
-It is rather clear that there should be communication flows between prosumers and the grid as most of the demand side management concepts suppose such a link [6] [7].
-Nevertheless as the expected number of active prosumers in the future seems quite large, the literature assume more and more communication between prosumers as this is a key ingredient to maintain more complex structures (such as coalitions), and thereby decrease the communication load on the grid’s side [8].
+It is rather clear that there should be communication flows between prosumers and the grid as most of the demand side management concepts suppose such a link \cite{WuHamedHuangBook2011}.
+Nevertheless as the expected number of active prosumers in the future seems quite large, the literature assume more and more communication between prosumers as this is a key ingredient to maintain more complex structures (such as coalitions), and thereby decrease the communication load on the grid’s side.
 Furthermore, it has been shown that coalitions in the smart grid provide nice ways of improving stability while posing exciting challenges in terms of network infrastructure and management.
-Self sufficient microgrids that can disconnect from the main grid [9], virtual power plants (VPP) that map small DER in sizeable and adjustable plants [10], or coalitions of electric vehicles that back up the grid in emergency situations (V2G) [11] are just a few examples of how coalitional concepts can enhance grid reliability.
+Self sufficient microgrids that can disconnect from the main grid \cite{Pahwa}, virtual power plants (VPP) that map small DER in sizeable and adjustable plants \cite{Braun} \cite{Ramchurn}, or coalitions of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn} are just a few examples of how coalitional concepts can enhance grid reliability.
 
 In this paper, our focus is on how coalitions of prosumers, aiming at selling energy to the grid, should be formed in order to support it efficiently.
-Namely, improving statistically production stability while decreasing communication needs for staying in stable state.
+Namely, improving statistically the production stability while decreasing the communication needs for staying in a stable state.
 The idea is that, by using prediction techniques, coalitions can propose contract values to the grid, enabling it to schedule production on its side accordingly.
@@
 In more details, and as will be explained in section \ref{sec:model}, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
-We used real meteorological traces [13] to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
+We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
 With these ingredients, we are able to run simulations and record the different output profiles as timeseries.
 
-As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering [16] [17] (see section \ref{sec:related}).
+As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999} (see section \ref{sec:related}).
 We will see that in such graphs, cliques represent structures of interest.
@@
 
-In \cite{Kota2011}, the authors study the formation of virtual power plants (VPP) composed of multiple self-interested DER.
-On the grid side, two requirements for the formation of virtual power plants are derived, namely the reliability of supply and the minimization of entities the grid has to deal with.
+Attacking the stability issues of renewable DER, the TradeWind project \cite{Europe} simulated the impact of wind power on electricity exchange and cross-border congestions by using a flow-based market model.
+The idea revolves around identifying key european interconnections (already existant or not) in order "to make optimal use of the european spacial de-correlation of wind power".
+It was indeed shown that geographical aggregation provides smoothening effects and that the amount of prediction errors for wind power in a geographical region diminishes as the region size increases, especially for short forecast horizons.
+
+On a narrower scale, the authors of \cite{Kota2011} study the formation of virtual power plants (VPP) composed of multiple self-interested DER.
+On the grid side, two requirements for the formation of virtual power plants are considered, namely the reliability of supply and the minimization of entities the grid has to deal with.
 From this, \cite{Kota2011} builds a pricing mechanism that encourages VPP to report true estimates of their aggregated production and penalizes prediction errors.
@@
 In this paper and as will be explained in section \ref{sec:model}, coalitions utilities depend on statistical properties of historical values.
-The goal being to form coalitions that, statistically spealing, are more likely to exhibit stable behaviors.
+The goal being to form coalitions that, statistically speaking, are more likely to exhibit stable behaviors.
 We will see that, when it comes to the formation of the coalitions, a prosumer is completely interchangeable with the timeserie representing its available production.
@@
 
-The following section of this paper explains the model we built in order obtain sufficient historical values for our prosumers, such that forming coalitions based on statistics is actually meaningfull.
+The following section of this paper explains the model we built in order to obtain sufficient historical values for our prosumers, such that forming coalitions based on statistics is actually meaningfull.
 
@@
 Furthermore, it seems obvious that the utility should increase with the contract value.
-The $ 1/|S| $ term indicates that we favorise small coalitions, mainly because they are easier to maintain in term of communication.
+The $ 1/|S| $ term indicates that we favorise small coalitions, mainly because they are easier to maintain in terms of communications.
 
@@
 
-The strategy of $ \{ij\} $ consists in maximizing its contract values as long as it respects this inequality, e.g to annonce $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $.
+The strategy of $ \{ij\} $ consists in maximizing its contract value as long as it respects this inequality, e.g to annonce $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $.
 It thus appears (as it was intuitively understandable) that, for equivalent sizes, coalitions with low relative standard deviations ( $ \sigma_{ij} / \mu_{ij} $ ) are able to announce higher contract values.
@@
 
-The model we used in order to simulate timeseries of available production provide some diversity because of the energetic mix/climate vectors combo.
+The model we used in order to simulate timeseries of available production provide some diversity because of the combo between the energetic mix and the climate vectors.
 Nevertheless, as the number of agents grows, the timeseries tend to exhibit similar patterns.
 This is particularly visible when creating a correlation graph $ G_{1}(\mathcal{A},E_{1},\omega_{1}) $ with the same kind of metric as \cite{Garas2008} or \cite{Onnela2004} ($ d_{ij} = 1 - \rho_{ij}^{2} $).
-Very well defined clusters appear in the corresponding $ \epsilon $-graph for large range of $ \epsilon $.
+Indeed, very well defined clusters appear in the corresponding $ \epsilon $-graph for any values of $ \epsilon $.
 
@@
 As expected \cite{Onnela2004}, independently of the $ \epsilon $ parameter, the resulting $ \epsilon $-graph exhibits henceforth much less clustering than in $ G_{1} $ and communities seem hardly visible.
-Therefore, using classical clustering or community detection algorithms do not provide very good results.
+Therefore, using classical clustering or community detection algorithms seem to provide poor results.
 
@@
 
-In the following, we fix the reliability to a given empirical value and observe how the coalitions evolves for different values of $P_{\phi}^{MIN}$.
-Note that the opposite is also possible although a little less intuitive, but not shown here because it leads to similar conclusions.
+In the following, we fix the reliability to a given empirical value ($\phi = Cste $) and observe how the coalitions evolves for different values of $P_{\phi}^{MIN}$.
+Note that the opposite is also possible although a little less intuitive, but not shown here because of space limitations and because it leads to similar conclusions.
 
@@
 The reason is that increasing the number of coalitions allows more coalitions to reach stability and enter the market, but there is a point where nodes bringing stability are not sufficient inside the coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility.
-Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs where information is flooded, meaning that the algorithm performances decrease.
+Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs where information is flooded, meaning that the algorithm performances also decrease.
 Naturally, it is also visible that small values of $P_{\phi}^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.
@@
   \includegraphics[scale=0.8]{./figure5/coals.pdf}
-  \caption{Evolution of the coalitions for different values of grid requirement $ P_{MIN} $. Blue dots represents clique percolation results while red dots stands for random process results.}
+  \caption{Evolution of the coalitions for different values of grid requirement $ P_{\phi}^{MIN} $. Blue dots represents clique percolation results while red dots stands for random process results.}
   \label{Fig3}
@@
 For our 200 prosumers, we suppose that the grid fixes a reliability $ \phi = 0.1 $ (meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time).
-We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P_{MIN} $ requirement.
+We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P_{\phi}^{MIN} $ requirement.
 As a comparison, we use a completely random algorithm that only asks for a number of coalitions and partitions the agents in a random fashion.
@@
 The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility.
-The $ P_{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.
+The $ P_{\phi}^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.
 
 Looking at figure \ref{Fig3}, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements.
-The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither to low or to high ($ P_{MIN} = 5 $ in figure \ref{Fig3}).
+The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither to low or to high ($ P_{\phi}^{MIN} = 5 $ in figure \ref{Fig3}).
 
-In order to explore a little bit more in this direction, we have a look, on figure \ref{Fig4}, at the percentage of coalitions able to enter the market for different values of $ P_{MIN} $.
-For consistency, we always average the results of the random procedure over 100 realizations (5 realizations for our algorithm, but the results are relatively stable), and the errorbars in the plots stands for the standard deviations of the results.
+In more details, the left plot of figure \ref{Fig4} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for both random (red curve) and clique percolation (blue curve).
+As for the right plot, it shows the percentage of coalitions able to enter the market for different values of $ P_{\phi}^{MIN} $.
+For consistency, we average the results of both plots of the random procedure over 100 realizations (5 realizations for clique percolation, but the results are relatively stable), and the errorbars in the plots stands for the standard deviations of the results.
 
-When the grid requirements are very low ( $ P_{MIN} $ close to zero), all coalitions for both algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
-But as $ P_{MIN} $ increases, we see the percentage of the random procedure quickly dropping while it stays constant for our algorithm.
-For $ P_{MIN} = 4 $, we see that only a little more than half of the coalitions for the random case are able to enter while all of them enters for the clique percolation algorithm.
+When the grid requirements are constant (left plot), and the number of desired coalitions is low, clique percolation generally performs only a little better than random.
+Nevetheless, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidely while clique percolation social welfare decreases slowly.
+
+When the grid requirements vary, for very low $ P_{\phi}^{MIN} $ (close to zero), all coalitions for both algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
+But as $ P_{\phi}^{MIN} $ increases, we see the percentage of the random procedure quickly dropping while it stays constant for our algorithm.
+For $ P_{\phi}^{MIN} = 4 $, we see that only a little more than half of the coalitions for the random case are able to enter while all of them enters for the clique percolation algorithm.
 Finally, when the grid requirement becomes to high, the acceptance percentage of our algorithm tends slowly to zero (only partly shown in this plot for readability).
@@
 \begin{figure}
-   \includegraphics[scale=0.7]{./figure4/survival.eps}
+\setlength{\tabcolsep}{0em}
+\begin{tabular}{cc}
+   \includegraphics[scale=0.45]{./figure7/uti.eps} &
+   \includegraphics[scale=0.45]{./figure4/survival.eps} \\
+\end{tabular}
    \caption{Percentage of coalitions able to enter the market for different values of grid requirement $ P_{MIN} $.

### 2014-09-05 Article.tex
update

@@
 \author{\IEEEauthorblockN{Nicolas Gensollen}
-\IEEEauthorblockA{Institut Mines Telecom\\
-Email: http://www.michaelshell.org/contact.html}
+\IEEEauthorblockA{CNRS SAMOVAR, \\
+Institut Mines Telecom}
 \and
 \IEEEauthorblockN{Vincent Gauthier}
-\IEEEauthorblockA{Twentieth Century Fox\\
-Springfield, USA\\
-Email: homer@thesimpsons.com}
+\IEEEauthorblockA{CNRS SAMOVAR, \\
+Institut Mines Telecom}
 \and
 \IEEEauthorblockN{Michel Marot\\ and Monique Becker}
-\IEEEauthorblockA{Starfleet Academy\\
-San Francisco, California 96678-2391\\
-Telephone: (800) 555--1212\\
-Fax: (888) 555--1212}}
+\IEEEauthorblockA{CNRS SAMOVAR, \\
+Institut Mines Telecom}}
 
@@
 %\boldmath
-There is a growing interest for coalitional concepts in the smart grid, spanning multiple objectives such as power losses reduction, benefits maximization or market stability...
+
+In a smart grid environement, we study coalitions of prosumers (agents that produce but also consume depending on time) which aim at selling energy to the grid.
+We reckon that the grid is able to specify two requirements (stability and minimum production) that enable a coalition to enter the market.
+We then design an algorithmm that seeks for uncorrelated profiles among the prosumers in order to form coalitions that are statistically stable enough to pass the grid requirements.
+The proposed algorithm uses complex system tools such as correlation graphs or clique percolation, and remains simple enough such that coalitions are formed in reasonable time.
+We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare, but that it is also more robust against grid minimum production threshold variations.
+
+%There is a growing interest for coalitional concepts in the smart grid, spanning multiple objectives such as power losses reduction, benefits maximization or market stability...
 In this paper, we study coalitions of prosumers (agents that produce but also consume depending on time) which aim at selling energy to the grid.
@@
 
-In this paper, our focus is on how coalitions of prosumers, aiming at selling energy to the grid, should be formed in order to support it efficiently.
-Namely, improving statistically the production stability while decreasing the communication needs for staying in a stable state.
+In this paper, our focus is on how improving statistically the production stability of coalitions of prosumers while decreasing the communication needs for staying in a stable state.
 The idea is that, by using prediction techniques, coalitions can propose contract values to the grid, enabling it to schedule production on its side accordingly.
@@
 
-In this paper, our concern is on how prosumers should be aggregated in order to optimize this tradeoff, that is, how coalitions should be formed with the intention of announcing “high contract values with high reliability”.
+In this paper, we seek at optimizing this tradeoff, that is forming coalitions able to announce “high contract values with high reliability”. We thus aim at:
+
+%In this paper, our concern is on how prosumers should be aggregated in order to optimize this tradeoff, that is, how coalitions should be formed with the intention of announcing “high contract values with high reliability”.
 We thus aim at:
@@
 On a narrower scale, the authors of \cite{Kota2011} study the formation of virtual power plants (VPP) composed of multiple self-interested DER.
-On the grid side, two requirements for the formation of virtual power plants are considered, namely the reliability of supply and the minimization of entities the grid has to deal with.
+On the grid side, two requirements for the formation of virtual power plants are considered : the reliability of supply and the minimization of entities the grid has to deal with.
 From this, \cite{Kota2011} builds a pricing mechanism that encourages VPP to report true estimates of their aggregated production and penalizes prediction errors.
@@
 Nevertheless, in \cite{Mantegna1999} Mantegna introduced another approach where stocks, represented by the timeseries of their daily log returns, are organized in a graph such that stocks (vertices) exhibiting similar price fluctuation patterns are close to each other.
-This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ but sometimes also $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
+This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
 Because there is a distance between every two nodes, \cite{Mantegna1999} obtains a complete graph where precious information is flooded.
@@
 At this point, some coalitions may be partially overlapping, meaning that a node can belong to multiple communities.
-Detection of overlapping communities is actually a very active field of research, especially in social networks where persons often belongs to several communities (family, friends, work colleagues...).
+Detection of overlapping communities is actually a very active field of research, especially in social networks where persons often belongs to several communities (family, friends, colleagues...).
 
@@
 Largely because of these reasons (but also because they are easily, and sometimes freely, available online) we chose to use meteorological traces as inputs.
-In this paper we used namely both french data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data for the United States covering year 2010 can also be found at \cite{NCDC}).
+In this paper we used namely both french data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data for the United States covering 2010 can also be found at \cite{NCDC}).
 The weather stations already offer a quite good sampling of a given territory, and, with the growth of personal weather stations constantly updating data bases, the mesh becomes finer and finer.
@@
 Some noise is added so that prosumers have similar but not identical cycles.
-\item \textbf{Seasonal cycles} : Consumption is higher in the winter because of heating and low in the summer (we did not consider air conditioning in this paper).
+\item \textbf{Seasonal cycles} : Consumption is higher in the winter because of heating and low in the summer (air conditioning is not considered).
 Temperature traces were the principal ingredient for modeling these cycles.
@@
 At this point, modeling agents consists in fixing a few parameters (most of the time drawn from random distributions) such as the geographical position, the number of wind turbines, PV, appliances, the temperature of confort, and so on...
-The objective is that, depending on the weather of her zone, the DER and appliances she owns, and the way she decides to heat her home, a given prosumer is able to compute her production/consumption at any time.
+The objective is that, depending on the weather of his zone, the DER and appliances he owns, and the way he decides to heat his home, a given prosumer is able to compute his production/consumption at any time.
 We thus used simple and known power curve models for linking weather variables to output power, such that a wind turbine (or a PV for instance) takes a wind speed (a solar irradiance) as input and gives an instantaneous output power \cite{Kota2011} \cite{windturbinemodel}.
-Any agent is thus able to compute at any time (within the traces bounds) how much she produces and consumes.
+Any agent is thus able to compute at any time how much he produces and consumes.
 
@@
 The reliability stipulates that for any coalition S willing to join the market, the probability of S value (at any instant t) being below its contract value should at most be $ \phi $.
-That is, $ \forall S,\ \forall t,\ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] \leq \phi $.
-Furthermore, for consistency, we restrict $ \phi $ to small values ( $ \phi << 0.5 $).
+That is, $ \forall S,\ \forall t,\ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] \leq \phi $ (for consistency we restrict $ \phi $ to small values ( $ \phi << 0.5 $)).
 The minimum contract value under $ \phi $ states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P_{\phi}^{MIN} $) will be accepted.
@@
 For simplification, we consider that coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value that obeys the reliability rule (a value we denote by $ P_{\phi}(S) $).
-If $ P_{\phi}(S) \geq P_{\phi}^{MIN} $, meaning that it also obeys the minimum contract value rule, then S annonces this value on the market : $ P_{S}^{CRCT} = P_{\phi}(S) $.
+If $ P_{\phi}(S) \geq P_{\phi}^{MIN} $, meaning that it also obeys the minimum contract value rule, then S announces this value on the market : $ P_{S}^{CRCT} = P_{\phi}(S) $.
 Otherwise, coalition S is not able to enter.
 Basically, a coalition S is valid if and only if :
-\[ \left\{ \begin{array}{lll}
+\begin{equation}
+\left\{ \begin{array}{lll}
 		\forall t,\ Pr[ P_{S}(t) \leq P_{\phi}(S)] \leq \phi\ \textit{{\scriptsize (reliability rule)}} \\
@@
 
-\end{array} \right. \]
-
-We now choose a very simple utility function that derives directly from the above remarks.
+\end{array} \right. 
+\label{rules}
+\end{equation}
+We now choose a very simple utility function (eq.
+\ref{utility}) that derives directly from the above remarks.
 If a coalition cannot provide a valid contract value, it receives naturally a utility of zero.
 Furthermore, it seems obvious that the utility should increase with the contract value.
-The $ 1/|S| $ term indicates that we favorise small coalitions, mainly because they are easier to maintain in terms of communications.
+The $ 1/|S| $ term in eq.
+\ref{utility} indicates that we favorise small coalitions, mainly because they are easier to maintain in terms of communications.
 
-\[ \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} \]
+\begin{equation}
+ \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} 
+\label{utility}
+\end{equation}
 
@@
 
-\[ \left\{ \begin{array}{lll}
+\begin{equation}
+\left\{ \begin{array}{lll}
 		\mu_{ij} = \mu_{i} + \mu_{j} \\
 		\sigma_{ij} = \sqrt{\sigma_{i}^{2} + \sigma_{j}^{2} + \rho_{ij}\sigma_{i}\sigma_{j}}
-\end{array} \right. \]
+\end{array} \right.
+\label{parameters}
+\end{equation}
 
@@
 
-\[ \dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P_{\phi}(ij) - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] \leq \phi \]
+\begin{equation}
+\dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P_{\phi}(ij) - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] \leq \phi
+\label{reliability}
+\end{equation}
 
-The strategy of $ \{ij\} $ consists in maximizing its contract value as long as it respects this inequality, e.g to annonce $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $.
+The strategy of $ \{ij\} $ consists in maximizing its contract value as long as it respects the inequality \ref{reliability}, e.g to annonce $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $.
 It thus appears (as it was intuitively understandable) that, for equivalent sizes, coalitions with low relative standard deviations ( $ \sigma_{ij} / \mu_{ij} $ ) are able to announce higher contract values.
@@
 
-The model we used in order to simulate timeseries of available production provide some diversity because of the combo between the energetic mix and the climate vectors.
+The model we used in order to simulate timeseries of available production provides some diversity because of the combination between the energetic mix and the climate vectors.
 Nevertheless, as the number of agents grows, the timeseries tend to exhibit similar patterns.
-This is particularly visible when creating a correlation graph $ G_{1}(\mathcal{A},E_{1},\omega_{1}) $ with the same kind of metric as \cite{Garas2008} or \cite{Onnela2004} ($ d_{ij} = 1 - \rho_{ij}^{2} $).
-Indeed, very well defined clusters appear in the corresponding $ \epsilon $-graph for any values of $ \epsilon $.
+This is apparent when creating a correlation graph $ G_{1}(\mathcal{A},E_{1},\omega_{1}) $ with the same kind of metric as \cite{Garas2008} or \cite{Onnela2004} ($ d_{ij} = 1 - \rho_{ij}^{2} $), where well defined clusters appear in the $ \epsilon $-graph for any values of $ \epsilon $.
 
@@
 However, as seems intuitively understandable, cliques of this graph tend to exhibit very good utility values.
-Such  structures contain ideed a link between every two nodes, meaning that the overall correlation is quite small.
+Such  structures contain indeed a link between every two nodes, meaning that the overall correlation is quite small.
 Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques, if it is too low, the $\epsilon$-graph of $ G_{2} $ will not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly.
-For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even becomes computationally repulsive.
+For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally repulsive.
 Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ seems too brutal.
@@
 
-\[ \tau_{i}(S) = \dfrac{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) - \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S-\{i\}) }{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S)} \]
+\begin{equation}
+\tau_{i}(S) = \dfrac{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) - \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S-\{i\}) }{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S)}
+\label{tau}
+\end{equation}
 
@@
 
-At this point, we have three degrees of freedom, namely the reliability ($ \phi $), the required power to enter the market ($P_{\phi}^{MIN}$), and the prunning parameter ($\epsilon$).
-For conveniance, we introduce the number of desired coalitions ($ N_{COAL} $) and we fix $\epsilon $ as the smallest value such that the resulting $ \epsilon $-graph contains at least $ N_{COAL} $ non overlaping cliques.
+At this point, we have three degrees of freedom : the reliability ($ \phi $), the required power to enter the market ($P_{\phi}^{MIN}$), and the prunning parameter ($\epsilon$).
+For convenience, we introduce the number of desired coalitions ($ N_{COAL} $) and we fix $\epsilon $ as the smallest value such that the $ \epsilon $-graph contains at least $ N_{COAL} $ non overlaping cliques.
 The overall utility depends now on the grid requirements and on the number of desired coalitions.
@@
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P_{\phi}^{MIN}$.
-Obviously, the highest utility values are found for high $ \phi $, because coalitions are able then to announce higher contract values, yielding higer utilities.
+Obviously, the highest utility values are found for high $ \phi $, because coalitions are able then to announce higher contract values, yielding higher utilities.
 
@@
   \includegraphics[scale=0.8]{./figure5/coals.pdf}
-  \caption{Evolution of the coalitions for different values of grid requirement $ P_{\phi}^{MIN} $. Blue dots represents clique percolation results while red dots stands for random process results.}
+  \caption{Evolution of the coalitions for different values of $ P_{\phi}^{MIN} $ for clique percolation (blue dots) and random process (red dots).}
   \label{Fig3}
@@
 Looking at figure \ref{Fig3}, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements.
-The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither to low or to high ($ P_{\phi}^{MIN} = 5 $ in figure \ref{Fig3}).
+The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither too low or too high ($ P_{\phi}^{MIN} = 5 $ in figure \ref{Fig3}).
 
@@
 For $ P_{\phi}^{MIN} = 4 $, we see that only a little more than half of the coalitions for the random case are able to enter while all of them enters for the clique percolation algorithm.
-Finally, when the grid requirement becomes to high, the acceptance percentage of our algorithm tends slowly to zero (only partly shown in this plot for readability).
+Finally, when the grid requirement becomes too high, the acceptance percentage of our algorithm tends slowly to zero (only partly shown in this plot for readability).
 
@@
 \end{tabular}
-   \caption{Percentage of coalitions able to enter the market for different values of grid requirement $ P_{MIN} $.
-Blue curve represents clique percolation results while red curve stands for random process results}
+   \caption{Percentage of coalitions able to enter the market for different values of $ P_{MIN} $ for clique percolation (blue) and random process (red).}
    \label{Fig4}
@@
 
-In this paper we collected meteorological traces for weather stations sampling a given territory.
+We believe that the originality of this paper lies in its willingness to exploit the de-correlation of prosumer profiles in order to build stable coalitions.
+In this direction, we presented a model based on meteorological traces, that captures the complex "\textit{energetic mix/climate vectors}" combination and generates realistic production and consumption patterns.
+We then built a framework that enables the grid to specify stability and minimum production requirements for filtering the coalitions.
+On this basis, we proposed a simple algorithm that seeks for uncorrelated prosumer patterns as potential seeds and expand them in coalitions able to rise above the grid requirements.
+
+The algorithm is validated against a random process, where we show that it performs better (coalitions are more stable and the overall production is more important) and that it exhibits a higher robustness/flexibility against grid requirements changes.
+
+%In this paper we collected meteorological traces for weather stations sampling a given territory.
 After discetization of the space, each zone receives the climate vector corresponding to its mother weather station.
@@
 
-Through this paper, our interest was on how we could circumscribe groups of prosumers that can rise above the grid requirements.
+%Through this paper, our interest was on how we could circumscribe groups of prosumers that can rise above the grid requirements.
 In this direction, we removed seasonal variations from the prosumers timeseries and proposed to organize them over a correlation graph with a reversed metric.
@@
 
-An interresting lead for future work would be the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analysed.
+Interresting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analysed through game theory.
+Showing that maintaining the coalitions formed with our algorithm necessitates less communication and less storage capacity could also conduct to a stimulating project.
 Besides, we believe that not restricting the algorithm to non overlapping coalitions and studying the strategies and weights of nodes with multiple options could lead to interesting works.
@@
 % conference papers do not normally have an appendix
-
-% use section* for acknowledgement
-\section*{Acknowledgment}
-
-The authors would like to thank...
 

### 2014-09-11 Article.tex
update

@@
 
-In a smart grid environement, we study coalitions of prosumers (agents that produce but also consume depending on time) which aim at selling energy to the grid.
-We reckon that the grid is able to specify two requirements (stability and minimum production) that enable a coalition to enter the market.
-We then design an algorithmm that seeks for uncorrelated profiles among the prosumers in order to form coalitions that are statistically stable enough to pass the grid requirements.
-The proposed algorithm uses complex system tools such as correlation graphs or clique percolation, and remains simple enough such that coalitions are formed in reasonable time.
-We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare, but that it is also more robust against grid minimum production threshold variations.
+In a smart grid environment, we study coalitions formation of prosumers that aim at selling energy to the grid.
+It is paramount for the grid operation that the energy producers are able to sustain the grid requirements in term of stability and minimum production requirement in order to enter the energy market.
+We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalition and a production of an uncorrelated source of energy in order to have stable production level during the course of the bidding.
+We proposed an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalition that meet such complex constraints.
+We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare for the power grid, but also that it is more robust against unforeseen production variations to due to changing weather condition for instance.
 
@@
 % \fi
-%
+% Mantegna
 % For peerreview papers, this IEEEtran command inserts a page break and
@@
 % no \IEEEPARstart
-One of the most basic ideas in the smart grid revolves around the introduction of communication means in the power grids, which could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
-Distributed energy resources (DER) such as wind turbines or photovoltaic panels are supposed to emerge and populate not only spread out farms, but also classical residential neighborhoods.
-Together with electric vehicles, and demand side management tools, they constitute the building blocks which will help turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
-Such energy aware agents that consume, produce, and are ready to make concessions (appliances delays, V2G…) to ensure grid stability are commonly called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
+One of the key ideas in the smart grid revolves around the introduction of communication means inside the power grids.
+That could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
+Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge in only in solar farms, but also in residential areas.
+Together with electric vehicles, and demand side management tools, they will constitute the building blocks which will help to turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
+Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G…) to ensure grid stability are called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
 
-It is rather clear that there should be communication flows between prosumers and the grid as most of the demand side management concepts suppose such a link \cite{WuHamedHuangBook2011}.
-Nevertheless as the expected number of active prosumers in the future seems quite large, the literature assume more and more communication between prosumers as this is a key ingredient to maintain more complex structures (such as coalitions), and thereby decrease the communication load on the grid’s side.
-Furthermore, it has been shown that coalitions in the smart grid provide nice ways of improving stability while posing exciting challenges in terms of network infrastructure and management.
-Self sufficient microgrids that can disconnect from the main grid \cite{Pahwa}, virtual power plants (VPP) that map small DER in sizeable and adjustable plants \cite{Braun} \cite{Ramchurn}, or coalitions of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn} are just a few examples of how coalitional concepts can enhance grid reliability.
+There seems to have a clear consensus on the benefit of having a bidirectional communication flows between prosumers and the grid as most of the demand side management concepts have proposed \cite{WuHamedHuangBook2011}.
+Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication pattern (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
+The formation of coalition inside the smart grid environment could be applied to various agents such as the self sufficient micro-grids\cite{Pahwa}, the virtual power plants (VPP), sizable and adjustable plants \cite{Braun} \cite{Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}, are just a few use case where the coalition of multiple agents can enhance grid reliability and in the mean time reduce the communication burden.
 
-In this paper, our focus is on how improving statistically the production stability of coalitions of prosumers while decreasing the communication needs for staying in a stable state.
-The idea is that, by using prediction techniques, coalitions can propose contract values to the grid, enabling it to schedule production on its side accordingly.
-Actually, contract values are particularly relevant in power exchanges contexts where energy is traded based on predictions (day-ahead estimations for instance).
-It seems obvious that, in such systems, participants should try to minimize prediction errors in order to maintain a stable state of grid operation, and that some penalty rules should be applied to ensure that coalitions are willing to report contract values correctly.
-Furthermore, as the coalition productions are supposed to come from (uncertain) renewables, the notion of reliability comes into play with a visible tradeoff between contract values worth and reliability.
+In this paper, we focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probability of staying in acceptable range of production.
+By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly.
+Meeting contract values for the prosumer are especially relevant in power trade conditions, where energy is traded based on day-ahead predictions.
+In this context, participants should try to minimize its prediction error in order to maintain a stable state for the grid.
+They may even occur some penalty if the production deviate from its initial contract value.
+However, it is pertinent for both side to maintain the production as stable as possible, for the prosumer the stability in the context for renewable energy source will impact its net gain and for the power grid its help to maintain its reliability.
 
 In this paper, we seek at optimizing this tradeoff, that is forming coalitions able to announce “high contract values with high reliability”. We thus aim at:
-
-%In this paper, our concern is on how prosumers should be aggregated in order to optimize this tradeoff, that is, how coalitions should be formed with the intention of announcing “high contract values with high reliability”.
-We thus aim at:
-
 \begin{itemize}
-\item Realistic representations (prosumers productions are non independent variables and come (indirectly) from real traces)
+\item Realistic prosumers productions with renewable production sources  from real traces base from weather measure.
 \item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section \ref{sec:model})

### 2014-09-11 Article.tex
update

@@
 They may even occur some penalty if the production deviate from its initial contract value.
-However, it is pertinent for both side to maintain the production as stable as possible, for the prosumer the stability in the context for renewable energy source will impact its net gain and for the power grid its help to maintain its reliability.
+However, it is pertinent for both side to maintain the production as stable as possible, for the prosumer the stability (with renewable energy source) will impact its net gain and for the power grid its help to maintain its reliability.
 

### 2014-09-11 Article.tex
upate

@@
 They may even occur some penalty if the production deviate from its initial contract value.
-However, it is pertinent for both side to maintain the production as stable as possible, for the prosumer the stability (with renewable energy source) will impact its net gain and for the power grid its help to maintain its reliability.
+However, it is pertinent for both side to maintain the production as stable as possible, for the prosumer the stability (with renewable energy source) will impact its net gain and for the power grid it helps to maintain its reliability.
 

### 2014-09-11 Article.tex
update

@@
 Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication pattern (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
-The formation of coalition inside the smart grid environment could be applied to various agents such as the self sufficient micro-grids\cite{Pahwa}, the virtual power plants (VPP), sizable and adjustable plants \cite{Braun} \cite{Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}, are just a few use case where the coalition of multiple agents can enhance grid reliability and in the mean time reduce the communication burden.
+The formation of coalition inside the smart grid environment could be applied to various type of agent such as the self sufficient micro-grids\cite{Pahwa}, the virtual power plants (VPP), sizable and adjustable plants \cite{Braun} \cite{Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}, are just a few use case where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.
 

### 2014-09-11 Article.tex
update

--- 2014-09-11 Article.tex
@@
 One of the key ideas in the smart grid revolves around the introduction of communication means inside the power grids.
-That could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
-Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge in only in solar farms, but also in residential areas.
+This could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
+Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge only in remote farms, but also in residential areas.
 Together with electric vehicles, and demand side management tools, they will constitute the building blocks which will help to turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
-Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G…) to ensure grid stability are called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
+Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G…) to ensure grid stability, and are commonly called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
 
-There seems to have a clear consensus on the benefit of having a bidirectional communication flows between prosumers and the grid as most of the demand side management concepts have proposed \cite{WuHamedHuangBook2011}.
-Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication pattern (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
-The formation of coalition inside the smart grid environment could be applied to various agents such as the self sufficient micro-grids\cite{Pahwa}, the virtual power plants (VPP), sizable and adjustable plants \cite{Braun} \cite{Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}, are just a few use case where the coalition of multiple agents can enhance grid reliability and in the mean time reduce the communication burden.
+There seems to have a clear consensus on the benefit of having bidirectional communication flows between prosumers and the grid as most of the demand side management concepts have proposed such an architecture \cite{WuHamedHuangBook2011}.
+Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
+The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun} \cite{Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}.
+These are just a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.
 
@@
 Meeting contract values for the prosumer are especially relevant in power trade conditions, where energy is traded based on day-ahead predictions.
-In this context, participants should try to minimize its prediction error in order to maintain a stable state for the grid.
-They may even occur some penalty if the production deviate from its initial contract value.
-However, it is pertinent for both side to maintain the production as stable as possible, for the prosumer the stability (with renewable energy source) will impact its net gain and for the power grid it helps to maintain its reliability.
+In this context, participants should try to minimize their prediction errors in order to maintain a stable state for the grid.
+They may even occur some penalties if the productions deviate from the initial contract values.
+However, it is pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy source) will impact its net gain, and for the power grid, it will helps to maintain its reliability.
 

### 2014-09-11 Article.tex
update de image

@@
 In a smart grid environment, we study coalitions formation of prosumers that aim at selling energy to the grid.
-It is paramount for the grid operation that the energy producers are able to sustain the grid requirements in term of stability and minimum production requirement in order to enter the energy market.
+It is paramount for the grid operation that the energy producers are able to sustain the grid demand in term of stability and minimum production requirement in order to enter the energy market.
 We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalition and a production of an uncorrelated source of energy in order to have stable production level during the course of the bidding.
@@
 In more details, and as will be explained in section \ref{sec:model}, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
-We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
-With these ingredients, we are able to run simulations and record the different output profiles as timeseries.
+We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic geographical correlations.
+With these ingredients, we are able simulate the different production/consumption profiles.
 
-As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999} (see section \ref{sec:related}).
+In section \ref{sec:forming} As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999} (see section \ref{sec:related}).
 We will see that in such graphs, cliques represent structures of interest.
@@
 We thereby use a modified clique percolation algorithm (see section \ref{sec:results}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
-
-The rest of the paper is organized as follows, section \ref{sec:related} locates our work in the related litterature.
-Section \ref{sec:model} explains the model we used to obtain realistic prosumers available power timeseries and introduces some notations.
-Section \ref{sec:forming} relates the method for building coalitions based on the agent timeseries.
-Finally, section \ref{sec:results} provides some results.
+Finally, in section \ref{sec:results} provides some results.
 
@@
 
-\section{Forming coalitions}
+\section{Coalitions Formation}
 \label{sec:forming}

### 2014-09-12 Article.tex

@@
 It is of common use to represent the situation and assess the stability of the solution by using game theoretic tools.
-Some papers focus on finding an optimal coalition structure giving a pool of autonomous self interested agents.
-In this direction, \cite{Saad2009} and \cite{Luan2014} design a distributed merge and split algorithm based on modified Pareto order with utility functions aiming at optimizing a parameter of interest such as energy losses or communication costs.
+Some papers \cite{Saad2009} \cite{Luan2014} focus then on finding an optimal coalition structure giving a pool of autonomous self interested agents using distributed merge/split algorithms.
 
@@
 The problem of clustering is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering.
-Nevertheless, in \cite{Mantegna1999} Mantegna introduced another approach where stocks, represented by the timeseries of their daily log returns, are organized in a graph such that stocks (vertices) exhibiting similar price fluctuation patterns are close to each other.
+Nevertheless, in \cite{Mantegna1999} Mantegna introduced another approach where stocks, represented by the timeseries of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other.
 This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
-Because there is a distance between every two nodes, \cite{Mantegna1999} obtains a complete graph where precious information is flooded.
-In order to discard unsignificant information, \cite{Mantegna1999} used a minimum spanning tree oriented technique and achieves a hierarchical clustering of the stocks.
-Since then, several other filtering techniques have been proposed, such as k-nearest graphs, or $ \epsilon $-graphs, that relax the tree assumption \cite{Onnela2004}.
 
-The idea of $\epsilon$-graphs consists in filtering edges based on their weight, only keeping edges whose weights are less than $ \epsilon $.
-Despite the procedure simplicity, the choice of $ \epsilon $ (or k for k-nearest graphs) is not trivial and may influence strongly the results.
-There is indeed to our knowledge no optimal rule for chosing $ \epsilon $.
+The idea of $\epsilon$-graphs, in this context, consists in filtering edges based on their weight, only keeping edges whose weights are less than $ \epsilon $.
 In \cite{Garas2008} and \cite{Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques...) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
@@
 Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities \cite{Newmanb}.
-Although several techniques exists (based on different graph properties : modularity \cite{Newmanb}, edge betweeness \cite{Girvan2002}, spectrum of Laplacian \cite{Newman}...), the clique percolation algorithm \cite{Lancichinetti} uses directly this observation and the idea behind it is actually quite simple.
-It starts indeed by searching for cliques (a complete subgraph) in the given graph as potential community cores.
-Because other nodes can naturally belong to a community with a less strong connectivity requirement, the algorithm uses a fitness function that quantifies the quality (according to some criterions) of a community structure inside the graph.
-Expansed cliques alternatively look in their respective neighborhood, select the node that yields the best increase in fitness, and finally incorporate it.
-The percolation algorithm stops when seeds cease to grow or when all nodes are affected to at least one seed.
-At this point, some coalitions may be partially overlapping, meaning that a node can belong to multiple communities.
+Although several techniques exists (based on different graph properties : modularity \cite{Newmanb}, edge betweeness \cite{Girvan2002}, spectrum of Laplacian \cite{Newman}...), the clique percolation algorithm \cite{Lancichinetti} uses directly this observation and employs a greedy expansion rule based on a fitness function.
+When the algorithm stops, some coalitions may be partially overlapping, meaning that a node can belong to multiple communities.
 Detection of overlapping communities is actually a very active field of research, especially in social networks where persons often belongs to several communities (family, friends, colleagues...).
@@
 
-\begin{center}
 \begin{figure}
-\setlength{\tabcolsep}{-0.7em}
-\begin{tabular}{cc}
-   \includegraphics[scale=0.5]{./figure1/util.eps} &
-   \includegraphics[scale=0.5]{./figure6/util2.eps} \\
-\end{tabular}
-\caption{Utility function in parameters space. Left in $ (\phi, P_{\phi}^{MIN}) $ when $ N_{COAL} $ is fixed. Right in $ (N_{COAL}, P_{\phi}^{MIN}) $ for a fixed $ \phi $}
+ \centering
+  \subcaptionbox{$ (\phi, P_{\phi}^{MIN}) $ space\label{fig2:a}}{\includegraphics[width=1.6in]{./figure1/util.eps}}\hspace{1em}%
+  \subcaptionbox{$ (N_{COAL}, P_{\phi}^{MIN}) $ space\label{fig2:b}}{\includegraphics[width=1.6in]{./figure6/util2.eps}}
+\caption{Utility function in parameters space : In $ (\phi, P_{\phi}^{MIN}) $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ (N_{COAL}, P_{\phi}^{MIN}) $ for a fixed $ \phi $ (fig \ref{fig2:b})}
 \label{Fig2}
 \end{figure}
-\end{center}
 
-As visible on the left plot of figure \ref{Fig2}, the parameters $\phi$ and $P_{\phi}^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P_{\phi}^{MIN}$ could potentially lead to valid coalitions (and positive utility values).
+As visible on figure \ref{fig2:a}, the parameters $\phi$ and $P_{\phi}^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P_{\phi}^{MIN}$ could potentially lead to valid coalitions (and positive utility values).
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P_{\phi}^{MIN}$.
@@
 
-On the right plot of figure \ref{Fig2}, we can see that when the grid constraints are fixed ($ \phi = 0.08 $ and $ P_{\phi}^{MIN} = 2 $ for instance) and only the number of coalitions varies, there is an increase in social welfare (sum of coalitions utilities) up to a maximum point (for $ N_{COAL} = 3 $ here) before a decrease.
+On figure \ref{fig2:b}, we can see that when the grid constraints are fixed ($ \phi = 0.08 $ and $ P_{\phi}^{MIN} = 2 $ for instance) and only the number of coalitions varies, there is an increase in social welfare (sum of coalitions utilities) up to a maximum point (for $ N_{COAL} = 3 $ here) before a decrease.
 The reason is that increasing the number of coalitions allows more coalitions to reach stability and enter the market, but there is a point where nodes bringing stability are not sufficient inside the coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility.
@@
   \includegraphics[scale=0.8]{./figure5/coals.pdf}
-  \caption{Evolution of the coalitions for different values of $ P_{\phi}^{MIN} $ for clique percolation (blue dots) and random process (red dots).}
+  \caption{Evolution of the coalitions for different values of $ P_{\phi}^{MIN} $ for clique percolation (blue dots) and random process (red dots).
+The diameters of the dots are proportional to the corresponding coalition sizes.}
   \label{Fig3}
@@
 
-In more details, the left plot of figure \ref{Fig4} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for both random (red curve) and clique percolation (blue curve).
+In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for both random (red curve) and clique percolation (blue curve).
 As for the right plot, it shows the percentage of coalitions able to enter the market for different values of $ P_{\phi}^{MIN} $.
@@
 
-When the grid requirements are constant (left plot), and the number of desired coalitions is low, clique percolation generally performs only a little better than random.
+When the grid requirements are constant (figure \ref{fig4:b}), and the number of desired coalitions is low, clique percolation generally performs only a little better than random.
 Nevetheless, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidely while clique percolation social welfare decreases slowly.
 
-When the grid requirements vary, for very low $ P_{\phi}^{MIN} $ (close to zero), all coalitions for both algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
+When the grid requirements vary, for very low $ P_{\phi}^{MIN} $, all coalitions for both algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
 But as $ P_{\phi}^{MIN} $ increases, we see the percentage of the random procedure quickly dropping while it stays constant for our algorithm.
@@
 
-\begin{center}
 \begin{figure}
-\setlength{\tabcolsep}{0em}
-\begin{tabular}{cc}
-   \includegraphics[scale=0.45]{./figure7/uti.eps} &
-   \includegraphics[scale=0.45]{./figure4/survival.eps} \\
-\end{tabular}
-   \caption{Percentage of coalitions able to enter the market for different values of $ P_{MIN} $ for clique percolation (blue) and random process (red).}
+ \centering
+  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{./figure7/uti.eps}}\hspace{1em}%
+  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{./figure4/survival.eps}}
+\caption{Evolution of the social welfare (fig \ref{fig4:a}) and the percentage of coalitions able to enter the market (fig \ref{fig4:b}) for different values of $ P_{MIN} $.
+Blue curves represent clique percolation and the red ones, the random process.}
 \label{Fig4}
 \end{figure}
-\end{center}
 

### 2014-09-12 Article.tex
update

@@
 In a smart grid environment, we study coalitions formation of prosumers that aim at selling energy to the grid.
-It is paramount for the grid operation that the energy producers are able to sustain the grid demand in term of stability and minimum production requirement in order to enter the energy market.
+It is paramount for the grid operation that the energy producers are able to sustain the grid requirements in term of stability and minimum production requirement in order to enter the energy market.
 We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalition and a production of an uncorrelated source of energy in order to have stable production level during the course of the bidding.
@@
 In more details, and as will be explained in section \ref{sec:model}, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
-We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic geographical correlations.
-With these ingredients, we are able simulate the different production/consumption profiles.
+We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
+With these ingredients, we are able to run simulations and record the different output profiles as timeseries.
 
-In section \ref{sec:forming} As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999} (see section \ref{sec:related}).
+As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999} (see section \ref{sec:related}).
 We will see that in such graphs, cliques represent structures of interest.
@@
 We thereby use a modified clique percolation algorithm (see section \ref{sec:results}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
-Finally, in section \ref{sec:results} provides some results.
+
+The rest of the paper is organized as follows, section \ref{sec:related} locates our work in the related litterature.
+Section \ref{sec:model} explains the model we used to obtain realistic prosumers available power timeseries and introduces some notations.
+Section \ref{sec:forming} relates the method for building coalitions based on the agent timeseries.
+Finally, section \ref{sec:results} provides some results.
 
@@
 
-\section{Coalitions Formation}
+\section{Forming coalitions}
 \label{sec:forming}

### 2014-09-12 Article.tex
update figure 2

--- 2014-09-11 Article.tex
@@
 
-In more details, and as will be explained in section \ref{sec:model}, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
+In more details, and as will be explained in section \ref{sec:model}, we consider agents that depend on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
 We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic geographical correlations.
@@
 \begin{figure}
-\includegraphics[scale=0.45]{figure2.jpg}
+\includegraphics[scale=0.45]{figure2/Fig2.pdf}
 \caption{Process diagram}

### 2014-09-12 Article.tex
update the article

--- 2014-09-12 Article.tex
@@
 We thereby use a modified clique percolation algorithm (see section \ref{sec:results}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
-
-The rest of the paper is organized as follows, section \ref{sec:related} locates our work in the related litterature.
-Section \ref{sec:model} explains the model we used to obtain realistic prosumers available power timeseries and introduces some notations.
-Section \ref{sec:forming} relates the method for building coalitions based on the agent timeseries.
 Finally, section \ref{sec:results} provides some results.

### 2014-09-12 Article.tex

@@
 
-In more details, and as will be explained in section \ref{sec:model}, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
+In more details, and as will be explained in section \ref{sec:model}, we have consider agents that depend on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
 We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.

### 2014-09-12 Article.tex
update

@@
 \label{sec:model}
+\subsection{Prosumer model}
 
@@
 
+\subsection{Coalition formation model}
 More formally, we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value of agent i at instant t (its instantaneous production minus consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t).
@@
 \end{equation}
+
+\subsection{Utility function}
 We now choose a very simple utility function (eq.

### 2014-09-12 Article.tex
various update

--- 2014-09-12 Article.tex
@@
 \begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
+% Title
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \title{Allowing Stable Groups of Prosumers to Enter the Market through Proper Aggregation}
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{\IEEEauthorblockN{Nicolas Gensollen}
-\IEEEauthorblockA{CNRS SAMOVAR, \\
-Institut Mines Telecom}
-\and
-\IEEEauthorblockN{Vincent Gauthier}
-\IEEEauthorblockA{CNRS SAMOVAR, \\
-Institut Mines Telecom}
-\and
-\IEEEauthorblockN{Michel Marot\\ and Monique Becker}
-\IEEEauthorblockA{CNRS SAMOVAR, \\
-Institut Mines Telecom}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Authors
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author{Nicolas Gensollen, Vincent Gauthier, Michel Marot and Monique Becker \\
+\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
+Telecom SudParis/Institut Mines Telecom\\
+Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}
 
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Abstract
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
-%\boldmath
-
 In a smart grid environment, we study coalitions formation of prosumers that aim at selling energy to the grid.
@@
 We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare for the power grid, but also that it is more robust against unforeseen production variations to due to changing weather condition for instance.
+\end{abstract}
 
-%There is a growing interest for coalitional concepts in the smart grid, spanning multiple objectives such as power losses reduction, benefits maximization or market stability...
-In this paper, we study coalitions of prosumers (agents that produce but also consume depending on time) which aim at selling energy to the grid.
-We reckon that the grid is able to specify two requirements (stability and minimum production) that enable a coalition to enter the market.
-Based on probability distributions of units available power, coalitions which are able to achieve low standard deviations are considered as stable coalitions.
-As it seems intuitively understandable that maintaining stable coalitions would be less expensive in terms of communication load than highly versatible ones, our objective consists in forming coalitions that satisfy grid requirements both in terms of stability and minimum production.
-Our definition of stability being implicitely linked to correlation between agents timeseries (see section \ref{sec:model}), we organize them over a correlation graph and use a variation of clique percolation to construct valid coalitions in reasonable time.
-\end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the conference you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals/conferences frown on
-% math in the abstract anyway.
-
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-% Mantegna
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Section I: Introduction
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 \label{sec:introduction}
-% no \IEEEPARstart
+
 One of the key ideas in the smart grid revolves around the introduction of communication means inside the power grids.
@@
 
-In this paper, we seek at optimizing this tradeoff, that is forming coalitions able to announce “high contract values with high reliability”. We thus aim at:
+In this paper, we seek to form coalitions that are able to announce "high contract values with high reliability". We thus aim at:
 \begin{itemize}
-\item Realistic prosumers productions with renewable production sources  from real traces base from weather measure.
-\item Minimizing the standard deviations of coalitions production probability distributions through the use of a proper utility function (see section \ref{sec:model})
-\item Being (relatively) scalable with the number of prosumers
+\item Build a realistic prosumers model with renewable production sources based on weather data (see section \ref{subsec:ProsumerModel}).
+\item Define coalitions formation model that enale the grisd that minimize production probability distributions through the use of a proper utility function (see section \ref{subsec:Coalition})
+\item finaly we defined and overall utiliy function to account for the  (see section \ref{subsec:UtilityFunc})
+\item (see section \ref{sec:forming})
 \end{itemize}
 
-In more details, and as will be explained in section \ref{sec:model}, we consider agents that, depending on meteorological conditions, personal preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
+In more details, and as will be explained in section \ref{sec:model}, we have consider agents that depend on meteorological conditions, various energy mix and  preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
 We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
@@
 We thereby use a modified clique percolation algorithm (see section \ref{sec:results}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
-Finally, section \ref{sec:results} provides some results.
+Finally, in section \ref{sec:results} provides some results and a conclusion in section \ref{sec:conclusion}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Section II: Related Work
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Section III: Model
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Model}
 \label{sec:model}
+\subsection{Prosumer model}\label{subsec:ProsumerModel}
 
@@
 
-%Smart grid perspectives assume that smart meters populate the network, recording production/consumption values and reporting these to data centers.
-Any entities (DER, prosumers, microgrids, VPP and so on) would thereby be characterizable through timeseries expressing how much these entities produce or consume.
-
-%Through our model, we wish thus to be able to generate, for any prosumer, its available production over time, which depend both on its pure production and consumption.
-Naturally, these are far from random in the sense that there are recurring patterns (consumption or expected solar irradience for instance) or statistical properties (see wind speed studies) that shape these profiles.
-We thus used meteorological traces as base inputs because they naturaly provide these key ingredients (sources).
-On the production side, as we only consider wind turbines and photovoltaic panels as generators, we concentrate primarily on wind speed and solar irradience traces.
-Variables such as temperature or humidity that correlates very well with consumption were also of great use.
-
-%Basically, the user only has to indicate the number of prosumers and the simulation dates.
-Prosumers are then positioned randomly on a squared lattice.
-Any point of this lattice is linked to a meteorological profile derived from the input traces.
-A few parameters individualizing the prosumers (number and kind of DER owned, consumption habits, and so on...) are also picked randomly and account for diversity.
-
+\subsection{Coalition formation model}\label{subsec:Coalition}
 More formally, we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value of agent i at instant t (its instantaneous production minus consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t).
@@
 \end{equation}
+
+\subsection{Utility function}\label{subsec:UtilityFunc}
 We now choose a very simple utility function (eq.
@@
 
-\section{Forming coalitions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Section V: Coalition Formation
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Coalition Formation}
 \label{sec:forming}
@@
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
+% Section VI: Conclusion
 %
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals/conferences use top floats
-% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
-% command of the stfloats package.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
-\label{sec: conclusion}
+\label{sec:conclusion}
 
@@
 
-%In this paper we collected meteorological traces for weather stations sampling a given territory.
-After discetization of the space, each zone receives the climate vector corresponding to its mother weather station.
-Prosumers are then distributed randomly between zones and each prosumer receives a random number of DER and loads as well as diverse consumption habits.
-Simulations are then runned and allow us to obtain, for each prosumer, the timeserie describing the instantaneous power that he could potentially be wanted to sell on the market.
-Nevertheless, as entering on the energy market is constrained by the grid operator, such that only sufficiently reliable and sufficiently productive units are allowed to enter, prosumers have usually no other choice than forming coalitions.
-We then introduced a framework such that the grid requirements take a precise meaning and such that coalitions announce contract values accordingly.
-
-%Through this paper, our interest was on how we could circumscribe groups of prosumers that can rise above the grid requirements.
-In this direction, we removed seasonal variations from the prosumers timeseries and proposed to organize them over a correlation graph with a reversed metric.
-A greedy clique percolation process on the $ \epsilon $-graph enables us to finally establish coalitions of agents.
-We showed in section \ref{sec:results} that our algorithm was actually finding coalitions with high utility values but also that it was far more robust than a random search over coalition structures when the grid operator increases his requirements (Figure \ref{Fig4}).
-
-Interresting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analysed through game theory.
+Interesting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analyzed through game theory.
 Showing that maintaining the coalitions formed with our algorithm necessitates less communication and less storage capacity could also conduct to a stimulating project.
@@
 
-% conference papers do not normally have an appendix
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-%\begin{thebibliography}{1}
-
-%\bibitem{IEEEhowto:kopka}
-%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+% Bibliography
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \bibliography{Article}
-
-%\end{thebibliography}
-
-% that's all folks
 \end{document}

### 2014-09-12 Article.tex
update

@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
+\documentclass[conference]{IEEEtran}
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
+% Package
 %
-\documentclass[conference]{IEEEtran}
-% Add the compsoc option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{ifpdf}
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+  \usepackage[pdftex]{graphicx}
   % declare the path(s) where your graphic files are
@@
   % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
 \else
@@
   % driver is specified.
-  % \usepackage[dvips]{graphicx}
+  \usepackage[dvips]{graphicx}
   % declare the path(s) where your graphic files are
@@
   % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
+  \DeclareGraphicsExtensions{.eps}
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
+
 \usepackage[utf8]{inputenc}
@@
 \usepackage[french, english]{babel}
-
-\usepackage{graphicx}
+\usepackage{cite}
 \usepackage{caption}
@@
 
-% *** MATH PACKAGES ***
 %
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
 % correct bad hyphenation here
@@
 \begin{figure}
-\includegraphics[scale=0.45]{figure2/Fig2.pdf}
+\includegraphics[scale=0.45]{figure2/Fig2}
 \caption{Process diagram}
@@
  \centering
-  \subcaptionbox{$ (\phi, P_{\phi}^{MIN}) $ space\label{fig2:a}}{\includegraphics[width=1.6in]{./figure1/util.eps}}\hspace{1em}%
-  \subcaptionbox{$ (N_{COAL}, P_{\phi}^{MIN}) $ space\label{fig2:b}}{\includegraphics[width=1.6in]{./figure6/util2.eps}}
+  \subcaptionbox{$ (\phi, P_{\phi}^{MIN}) $ space\label{fig2:a}}{\includegraphics[width=1.6in]{./figure1/util}}\hspace{1em}%
+  \subcaptionbox{$ (N_{COAL}, P_{\phi}^{MIN}) $ space\label{fig2:b}}{\includegraphics[width=1.6in]{./figure6/util2}}
 \caption{Utility function in parameters space : In $ (\phi, P_{\phi}^{MIN}) $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ (N_{COAL}, P_{\phi}^{MIN}) $ for a fixed $ \phi $ (fig \ref{fig2:b})}
@@
 \begin{figure}[b]
-  \includegraphics[scale=0.8]{./figure5/coals.pdf}
+  \includegraphics[scale=0.8]{./figure5/coals}
   \caption{Evolution of the coalitions for different values of $ P_{\phi}^{MIN} $ for clique percolation (blue dots) and random process (red dots).
@@
  \centering
-  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{./figure7/uti.eps}}\hspace{1em}%
-  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{./figure4/survival.eps}}
-\caption{Evolution of the social welfare (fig \ref{fig4:a}) and the percentage of coalitions able to enter the market (fig \ref{fig4:b}) for different values of $ P_{MIN} $.
+  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{figure7/uti}}\hspace{1em}%
+  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{figure4/survival}}
+\caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.
 Blue curves represent clique percolation and the red ones, the random process.}

### 2014-09-12 Article.tex
update

@@
 Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
-The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun} \cite{Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}.
+The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}.
 These are just a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.
@@
 \begin{itemize}
-\item Build a realistic prosumers model with renewable production sources based on weather data (see section \ref{subsec:ProsumerModel}).
-\item Define coalitions formation model that enale the grisd that minimize production probability distributions through the use of a proper utility function (see section \ref{subsec:Coalition})
-\item finaly we defined and overall utiliy function to account for the  (see section \ref{subsec:UtilityFunc})
-\item (see section \ref{sec:forming})
+\item Building a realistic prosumers model with renewable production sources based on weather data (see section \ref{subsec:ProsumerModel}).
+\item Define a coalitions formation model that enable the grid to set some requirements to enable any group of prosumer to sell its production (see section \ref{subsec:Coalition})
+\item Define an utility function that will satisfy the grid requirement and maximize the stability of the prosumer coalition production (see section \ref{subsec:UtilityFunc})
+\item Define a coalition formation algorithm (see section \ref{sec:forming})
 \end{itemize}
 
-In more details, and as will be explained in section \ref{sec:model}, we have consider agents that depend on meteorological conditions, various energy mix and  preferences, and their appliances (loads and generators), consume and produce more or less energy (using only wind turbines and PV as generators, which can be quite simply and efficiently modeled).
-We used real meteorological traces (see section \ref{sec:model}) to account for seasonal and daily variations in the prosumers energetic output profiles as well as realistic correlations.
-With these ingredients, we are able to run simulations and record the different output profiles as timeseries.
+In section \ref{sec:model}, we considered agents production that depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, their appliances (loads and generators).
+We used real meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
+With these ingredients, we are able to run realistic simulations and record the different output profiles as time-series.
 
-As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999} (see section \ref{sec:related}).
+As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999}.
 We will see that in such graphs, cliques represent structures of interest.
@@
 In this paper and as will be explained in section \ref{sec:model}, coalitions utilities depend on statistical properties of historical values.
-The goal being to form coalitions that, statistically speaking, are more likely to exhibit stable behaviors.
-We will see that, when it comes to the formation of the coalitions, a prosumer is completely interchangeable with the timeserie representing its available production.
+The goal being to form coalitions that, statistically more likely to exhibit stable behaviors.
+We show, when it comes to the formation of the coalitions, a prosumer is completely interchangeable with the time-serie representing its available production.
 
-The setting is thus similar with some fiancial studies on stock exchanges, where researchers tried to find relevant clusters of stocks based on their daily prices variations.
+The setting is thus similar with some financial studies on stock exchanges, where researchers tried to find relevant clusters of stocks based on their daily prices variations.
 The problem of clustering is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering.
-Nevertheless, in \cite{Mantegna1999} Mantegna introduced another approach where stocks, represented by the timeseries of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other.
+Nevertheless, in \cite{Mantegna1999} the author introduced an approach where stocks time-series of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other.
 This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
@@
 The idea of $\epsilon$-graphs, in this context, consists in filtering edges based on their weight, only keeping edges whose weights are less than $ \epsilon $.
-In \cite{Garas2008} and \cite{Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques...) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
-One of the conclusions was that "strong links" (i.e links between strongly correlated timeseries) are responsible for clustering while "weak links" provide network connectivity.
+In \cite{Garas2008, Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques...) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
+One of the conclusions was that "strong links" (i.e links between strongly correlated time-series) are responsible for clustering while "weak links" provide network connectivity.
 
-Presented in this way, the timeseries clustering task seems very close to graph community detection.
+Presented in this way, the time-series clustering task seems very close to graph community detection.
 Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities \cite{Newmanb}.
-Although several techniques exists (based on different graph properties : modularity \cite{Newmanb}, edge betweeness \cite{Girvan2002}, spectrum of Laplacian \cite{Newman}...), the clique percolation algorithm \cite{Lancichinetti} uses directly this observation and employs a greedy expansion rule based on a fitness function.
+Although several techniques exists (based on different graph properties : modularity \cite{Newmanb}, edge betweenness \cite{Girvan2002}, spectrum of graph Laplacian \cite{Newman}), the clique percolation algorithm \cite{Lancichinetti} uses directly this observation and employs a greedy expansion rule based on a fitness function.
 When the algorithm stops, some coalitions may be partially overlapping, meaning that a node can belong to multiple communities.
 Detection of overlapping communities is actually a very active field of research, especially in social networks where persons often belongs to several communities (family, friends, colleagues...).
-
-The following section of this paper explains the model we built in order to obtain sufficient historical values for our prosumers, such that forming coalitions based on statistics is actually meaningfull.
 

### 2014-09-12 Article.tex
update the introduction

@@
 
-In section \ref{sec:model}, we considered agents production that depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, their appliances (loads and generators).
-We used real meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
+In section \ref{sec:model}, we considered agents production that depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, their appliances (loads).
+We used meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
 With these ingredients, we are able to run realistic simulations and record the different output profiles as time-series.
 
-As correlations play a significant role in our mechanism, we organize the prosumers in a correlation graph, a quite popular approach in market stocks clustering since \cite{Mantegna1999}.
-We will see that in such graphs, cliques represent structures of interest.
-In our settings, it means that cliques are “good” (from the utility point of view) foundations for high valued coalitions.
-We thereby use a modified clique percolation algorithm (see section \ref{sec:results}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill grid requirements.
+In order to select group prosumer that lower the volatility of the coalition's production, we define an algorithm based on a popular approach for stocks market clustering \cite{Mantegna1999}.
+In our settings, we claim that cliques are "good" from the utility point of view and thereby we use a modified clique percolation algorithm (see section \ref{sec:forming}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill both the grid requirements and lower the production volatility.
 Finally, in section \ref{sec:results} provides some results and a conclusion in section \ref{sec:conclusion}.

### 2014-09-12 Article.tex
I am done for today

@@
 It is paramount for the grid operation that the energy producers are able to sustain the grid demand in term of stability and minimum production requirement in order to enter the energy market.
-We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalition and a production of an uncorrelated source of energy in order to have stable production level during the course of the bidding.
+We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalition and a production of an uncorrelated source of energy in order to have a steady production level during the course of the bidding.
 We proposed an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalition that meet such complex constraints.
@@
 Together with electric vehicles, and demand side management tools, they will constitute the building blocks which will help to turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
-Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G…) to ensure grid stability, and are commonly called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
+Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G) to ensure grid stability, and are commonly called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
 
-There seems to have a clear consensus on the benefit of having bidirectional communication flows between prosumers and the grid as most of the demand side management concepts have proposed such an architecture \cite{WuHamedHuangBook2011}.
+There seems to have a clear consensus on the benefit of having a bidirectional communication flows between prosumers and the grid, as most of the demand side management concepts have proposed\cite{WuHamedHuangBook2011}.
 Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
@@
 
-In this paper, we focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probability of staying in acceptable range of production.
+In this paper, we focus on how to statistically improve the production stability by carefully forming coalitions of prosumers (probability of staying in acceptable range of production).
 By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly.
@@
 They may even occur some penalties if the productions deviate from the initial contract values.
-However, it is pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy source) will impact its net gain, and for the power grid, it will helps to maintain its reliability.
+However, it remains pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy source) will impact its net gain, and for the power grid, it will helps to maintain its reliability.
 
@@
 
-In order to select group prosumer that lower the volatility of the coalition's production, we define an algorithm based on a popular approach for stocks market clustering \cite{Mantegna1999}.
-In our settings, we claim that cliques are "good" from the utility point of view and thereby we use a modified clique percolation algorithm (see section \ref{sec:forming}) that will enable us to expand the cliques as much as needed in order to form proper coalitions that fulfill both the grid requirements and lower the production volatility.
+In order to select groups of prosumer that lower the volatility of the coalition's production, we define an algorithm based on a popular approach for stocks market clustering \cite{Mantegna1999}.
+We use a modified clique percolation algorithm (see section \ref{sec:forming}) that will enable us to expand the cliques as needed in order to form proper coalitions that fulfill both the grid requirements and lower the production volatility.
 Finally, in section \ref{sec:results} provides some results and a conclusion in section \ref{sec:conclusion}.

### 2014-09-12 Article.tex
a few corrections

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Allowing Stable Groups of Prosumers to Enter the Market through Proper Aggregation}
+\title{Coalitions of Uncorrelated Prosumers in a Smart Grid Environment}
 
@@
 In a smart grid environment, we study coalitions formation of prosumers that aim at selling energy to the grid.
-It is paramount for the grid operation that the energy producers are able to sustain the grid demand in term of stability and minimum production requirement in order to enter the energy market.
-We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalition and a production of an uncorrelated source of energy in order to have a steady production level during the course of the bidding.
-We proposed an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalition that meet such complex constraints.
-We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare for the power grid, but also that it is more robust against unforeseen production variations to due to changing weather condition for instance.
+It is paramount for the grid operation that the energy producers are able to sustain the grid demand in terms of stability and minimum production requirement in order to enter the energy market.
+We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalitions and a production of an uncorrelated source of energy in order to have a steady production level during the course of the bidding.
+We proposed an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalitions that meet such complex constraints.
+We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare for the power grid, but also that it is more robust against unforeseen production variations due to changing weather conditions for instance.
 \end{abstract}
@@
 
-There seems to have a clear consensus on the benefit of having a bidirectional communication flows between prosumers and the grid, as most of the demand side management concepts have proposed\cite{WuHamedHuangBook2011}.
+There seems to have a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as most of the demand side management concepts have proposed such an architecture \cite{WuHamedHuangBook2011}.
 Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
@@
 
-In this paper, we focus on how to statistically improve the production stability by carefully forming coalitions of prosumers (probability of staying in acceptable range of production).
+In this paper, we focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable range of production.
 By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly.
-Meeting contract values for the prosumer are especially relevant in power trade conditions, where energy is traded based on day-ahead predictions.
+Meeting contract values for the prosumers is especially relevant in power trade conditions, where energy is traded based on day-ahead predictions.
 In this context, participants should try to minimize their prediction errors in order to maintain a stable state for the grid.
@@
 \begin{itemize}
-\item Building a realistic prosumers model with renewable production sources based on weather data (see section \ref{subsec:ProsumerModel}).
-\item Define a coalitions formation model that enable the grid to set some requirements to enable any group of prosumer to sell its production (see section \ref{subsec:Coalition})
-\item Define an utility function that will satisfy the grid requirement and maximize the stability of the prosumer coalition production (see section \ref{subsec:UtilityFunc})
+\item Building a realistic prosumer model with renewable production sources based on weather data (see section \ref{subsec:ProsumerModel}).
+\item Define a coalition formation model that enables the grid to set some requirements under which any group of prosumers will be allowed to sell its production (see section \ref{subsec:Coalition})
+\item Define a utility function that will satisfy the grid requirements and maximize the stability of the prosumer coalition productions (see section \ref{subsec:UtilityFunc})
 \item Define a coalition formation algorithm (see section \ref{sec:forming})
@@
 
-In section \ref{sec:model}, we considered agents production that depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, their appliances (loads).
+In section \ref{sec:model}, we considered agents productions that depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, and their appliances (loads).
 We used meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
@@
 
-In order to select groups of prosumer that lower the volatility of the coalition's production, we define an algorithm based on a popular approach for stocks market clustering \cite{Mantegna1999}.
+In order to select groups of prosumers that lower the volatility of the coalition's productions, we define an algorithm based on a popular approach for stocks market clustering \cite{Mantegna1999}.
 We use a modified clique percolation algorithm (see section \ref{sec:forming}) that will enable us to expand the cliques as needed in order to form proper coalitions that fulfill both the grid requirements and lower the production volatility.
-Finally, in section \ref{sec:results} provides some results and a conclusion in section \ref{sec:conclusion}.
+Finally, section \ref{sec:results} provides some results and a conclusion in section \ref{sec:conclusion}.
 

### 2014-09-12 Article.tex
curve updates

@@
 However, these clusters of strongly correlated timeseries are the exact opposite of what we are seeking.
-We can indeed consider them directly as coalitions and compute their utilities, and the results show, as expected, terrible values (far worse than a random split of the agents in the same number of coalitions).
+We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves on figure \ref{Fig4}), as expected, terrible values (far worse than a random split of the agents in the same number of coalitions).
 
@@
 
-In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for both random (red curve) and clique percolation (blue curve).
-As for the right plot, it shows the percentage of coalitions able to enter the market for different values of $ P_{\phi}^{MIN} $.
-For consistency, we average the results of both plots of the random procedure over 100 realizations (5 realizations for clique percolation, but the results are relatively stable), and the errorbars in the plots stands for the standard deviations of the results.
+In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario.
+As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P_{\phi}^{MIN} $.
+For consistency, we average the results of both plots of the random procedure over 100 realizations, and the errorbars in the plots stands for the standard deviations of the results.
 
@@
 
-When the grid requirements vary, for very low $ P_{\phi}^{MIN} $, all coalitions for both algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
-But as $ P_{\phi}^{MIN} $ increases, we see the percentage of the random procedure quickly dropping while it stays constant for our algorithm.
-For $ P_{\phi}^{MIN} = 4 $, we see that only a little more than half of the coalitions for the random case are able to enter while all of them enters for the clique percolation algorithm.
-Finally, when the grid requirement becomes too high, the acceptance percentage of our algorithm tends slowly to zero (only partly shown in this plot for readability).
+When the grid requirements vary, for very low $ P_{\phi}^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
+But as $ P_{\phi}^{MIN} $ increases, we see the percentage of the correlated coalitions crashing down very quickly.
+After a few increases in $ P_{\phi}^{MIN} $, the percentage of the random procedure starts droping while it stays almost constant for our algorithm.
+For $ P_{\phi}^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm.
+Finally, when the grid requirement becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).
 
@@
  \centering
-  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{figure7/uti}}\hspace{1em}%
-  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{figure4/survival}}
+  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{figure7_with_green_curve_v1/uti5}}\hspace{1em}%
+  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{figure4_with_green_curve/survival2_old}}
 \caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.
-Blue curves represent clique percolation and the red ones, the random process.}
+Blue curves represent clique percolation, red curves, the random process, and the green ones, correlated coalitions.}
 \label{Fig4}

### 2014-09-15 Article.tex
update Vincent + Monique

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Coalitions of Uncorrelated Prosumers in a Smart Grid Environment}
+\title{Coalition Formation Algorithm of Prosumers in a Smart Grid Environment}
 
@@
 \begin{abstract}
-In a smart grid environment, we study coalitions formation of prosumers that aim at selling energy to the grid.
+In a smart grid environment, we study coalition formation of prosumers that aim at selling energy to the grid.
 It is paramount for the grid operation that the energy producers are able to sustain the grid demand in terms of stability and minimum production requirement in order to enter the energy market.
-We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minium energy level for the coalitions and a production of an uncorrelated source of energy in order to have a steady production level during the course of the bidding.
-We proposed an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalitions that meet such complex constraints.
+We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minimum energy level for the coalitions and a steady production level which leads to finding uncorrelated source of energy  to form a coalition.
+We propose an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalitions that meet such complex constraints.
 We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare for the power grid, but also that it is more robust against unforeseen production variations due to changing weather conditions for instance.
@@
 
-In section \ref{sec:model}, we considered agents productions that depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, and their appliances (loads).
-We used meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
+In section \ref{sec:model}, we consider that the agent's production depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, and their appliances (loads).
+We use meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
 With these ingredients, we are able to run realistic simulations and record the different output profiles as time-series.
@@
 
-Attacking the stability issues of renewable DER, the TradeWind project \cite{Europe} simulated the impact of wind power on electricity exchange and cross-border congestions by using a flow-based market model.
+Tackling the stability issues of renewable DER, the TradeWind project \cite{Europe} simulated the impact of wind power on electricity exchange and cross-border congestions by using a flow-based market model.
 The idea revolves around identifying key european interconnections (already existant or not) in order "to make optimal use of the european spacial de-correlation of wind power".
@@
 
-In this paper and as will be explained in section \ref{sec:model}, coalitions utilities depend on statistical properties of historical values.
-The goal being to form coalitions that, statistically more likely to exhibit stable behaviors.
-We show, when it comes to the formation of the coalitions, a prosumer is completely interchangeable with the time-serie representing its available production.
-
+In this paper we ought to form coalitions utilities that depend on statistical properties of time-series values.
 The setting is thus similar with some financial studies on stock exchanges, where researchers tried to find relevant clusters of stocks based on their daily prices variations.
@@
 
-The idea of $\epsilon$-graphs, in this context, consists in filtering edges based on their weight, only keeping edges whose weights are less than $ \epsilon $.
-In \cite{Garas2008, Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques...) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
-One of the conclusions was that "strong links" (i.e links between strongly correlated time-series) are responsible for clustering while "weak links" provide network connectivity.
+$\epsilon$-graphs, consists in filtering edges based on their weight, only keeping edges whose weights are less than $ \epsilon $.
+In \cite{Garas2008, Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
+However, there is not well defined method to select of right tradeoff $ \epsilon $ as function of the network topology.
 
@@
 When the algorithm stops, some coalitions may be partially overlapping, meaning that a node can belong to multiple communities.
-Detection of overlapping communities is actually a very active field of research, especially in social networks where persons often belongs to several communities (family, friends, colleagues...).
+Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities in the mean time.
 
@@
 
-A major concern while designing our model was its ability to translate correctly realistic patterns (consumption and production) as well as realistic geographical correlations between patterns.
-Largely because of these reasons (but also because they are easily, and sometimes freely, available online) we chose to use meteorological traces as inputs.
-In this paper we used namely both french data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data for the United States covering 2010 can also be found at \cite{NCDC}).
-The weather stations already offer a quite good sampling of a given territory, and, with the growth of personal weather stations constantly updating data bases, the mesh becomes finer and finer.
+Our major concern was to design a prosumer model with realistic patterns of consumption, production as well as realistic geographical correlations between them.
+We based our prosumer model on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).
+
 As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering traces for these stations.
@@
 
-We thus collect for all chosen weather stations three kind of traces (average wind speed, cloudiness, and average temperature), which we will call a climate vector in the following.
-We consider that cimate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector.
+We thus collect for all chosen weather stations three kinds of data (average wind speed, cloudiness, and average temperature), which we will call a climate vector in the following.
+We consider that climate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector.
 
-At this point, modeling agents consists in fixing a few parameters (most of the time drawn from random distributions) such as the geographical position, the number of wind turbines, PV, appliances, the temperature of confort, and so on...
+At this point, modeling agents consists in fixing a few parameters (most of the time drawn from random distributions) such as the geographical position, the number of wind turbines, PV, appliances, the temperature of confort, and so on.
 The objective is that, depending on the weather of his zone, the DER and appliances he owns, and the way he decides to heat his home, a given prosumer is able to compute his production/consumption at any time.
@@
 \subsection{Coalition formation model}\label{subsec:Coalition}
-More formally, we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value of agent i at instant t (its instantaneous production minus consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t).
+More formally, we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value at agent i at instant t (its instantaneous production minus his instantaneous consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t).
 During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
-As expected (because we introduced them), the timeseries exhibit high seasonal patterns, but completely different from agent to agent because they depend both on the energetic mix and habits of a given prosumer.
+As expected (because we introduced them), the time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both of their energetic mix and habits of a given prosumer.
 Nevertheless, these macro variations can hide the interesting information in the correlation coefficients and twist the rest of our process.
 Thereby, we remove these seasonal effects for all agents (see second block of figure \ref{Fig1}).
-We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting timeserie for  agent i.
+We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent i.
 For any agent i, we note $ \mathcal{P}_{i} $ the probability distribution drawn from $ \mathcal{T}_{i} $ and by $ P_{i} $ a random variable following $ \mathcal{P}_{i} $.
 
-We now extand these notations to any coalition $ S \subset \mathcal{A} $ : 
+We now extend these notations to any coalition $ S \subset \mathcal{A} $ : 
 \begin{itemize}
@@
 
-At this point, if a coalition S whishes to join the market, it has to choose a contract value that fulfill the two conditions.
+At this point, if a coalition S wishes to join the market, it has to choose a contract value that fulfill the two conditions.
 For simplification, we consider that coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value that obeys the reliability rule (a value we denote by $ P_{\phi}(S) $).

### 2014-09-15 Article.tex
update Nico

@@
 It is paramount for the grid operation that the energy producers are able to sustain the grid demand in terms of stability and minimum production requirement in order to enter the energy market.
-We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minimum energy level for the coalitions and a steady production level which leads to finding uncorrelated source of energy  to form a coalition.
+We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minimum energy level for the coalitions and a steady production level which leads to finding uncorrelated sources of energy  to form a coalition.
 We propose an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalitions that meet such complex constraints.
@@
 They may even occur some penalties if the productions deviate from the initial contract values.
-However, it remains pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy source) will impact its net gain, and for the power grid, it will helps to maintain its reliability.
+However, it remains pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy sources) will impact its net gain, and for the power grid, it will helps to maintain its reliability.
 
@@
 In \cite{Garas2008, Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
-However, there is not well defined method to select of right tradeoff $ \epsilon $ as function of the network topology.
+However, there is no well defined method to select of right tradeoff $ \epsilon $ as function of the network topology.
 
@@
 The objective is that, depending on the weather of his zone, the DER and appliances he owns, and the way he decides to heat his home, a given prosumer is able to compute his production/consumption at any time.
-We thus used simple and known power curve models for linking weather variables to output power, such that a wind turbine (or a PV for instance) takes a wind speed (a solar irradiance) as input and gives an instantaneous output power \cite{Kota2011} \cite{windturbinemodel}.
+
+We thus used existing models enabling conversion between weather variables and output powers (see power curves in \cite{Kota2011} \cite{windturbinemodel}).
+\cite{Dans2007} provides also a convenient way of using cloudiness traces as realistic degradation factors in a \textit{"clear blue sky"} solar model, which enables us to rebuild realistic solar irradiance traces for stations that do not provide such information directly.
 Any agent is thus able to compute at any time how much he produces and consumes.
@@
 Thereby, we remove these seasonal effects for all agents (see second block of figure \ref{Fig1}).
-We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent i.
-For any agent i, we note $ \mathcal{P}_{i} $ the probability distribution drawn from $ \mathcal{T}_{i} $ and by $ P_{i} $ a random variable following $ \mathcal{P}_{i} $.
+We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent i, and by $ \mathcal{P}_{i} $ the corresponding probability distribution.
 
@@
 \item $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $
-\item $ \mathcal{T}_{S} = \{ P_{S}(t_{0}),...,P_{S}(t_{K}) \} $ with $ \mathcal{P}_{S} $ its probability distribution
+\item $ \mathcal{T}_{S} = \{ P_{S}(t_{0}),...,P_{S}(t_{K}) \} $ with $ \mathcal{P}_{S} $ the corresponding probability distribution
 \item A coalition S has the possibility to announce a contract value $ P_{S}^{CRCT} $ on the market.
@@
 
-In this paper, we consider that coalition announcements are constrained by the grid. Namely, the grid has the possibility of fixing two parameters :
+Due to the volatile nature of DER productions or user consumptions, the available power of a coalition will oscillates around $ P_{S}^{CRCT} $.
+It will be the responsability of the coalition, that is the information system maintaining the coalition, to compensate for deviations.
+In this paper, we consider then that the smaller these deviations, the better.
+Based on the $ \mathcal{T}_{S} $ distribution, the probability of S value (at any instant t) being below its contract value $ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] $ appears therefore as a quantity of interest that we would want to keep as low as possible.
 
-\begin{itemize}
-\item the \textbf{reliability} (denoted by $ \phi \in [0,1] $)
-\item the \textbf{minimum contract value} ( $ P_{\phi}^{MIN} $). 
-\end{itemize}
+In this model, we consider that only the coalitions that have this probability value sufficiently low are acceptable.
+That this, we denote by $ \phi \in [0,1] $ the threshold imposed to the coalitions over this quantity, and we call it the \textbf{reliability} in the following.
+The reliability stipulates then that for any coalition S willing to join the market, the probability of S value (at any instant t) being below its contract value should at most be $ \phi $.
+That is, $ \forall S,\ \forall t,\ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] \leq \phi $ ($ \phi $ being restricted to small values for consistency).
 
-The reliability stipulates that for any coalition S willing to join the market, the probability of S value (at any instant t) being below its contract value should at most be $ \phi $.
-That is, $ \forall S,\ \forall t,\ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] \leq \phi $ (for consistency we restrict $ \phi $ to small values ( $ \phi << 0.5 $)).
-The minimum contract value under $ \phi $ states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P_{\phi}^{MIN} $) will be accepted.
+We then consider another constraint denoted by $ P^{MIN} $, as the \textbf{minimum contract value}, which states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P^{MIN} $) will be accepted.
 
@@
 For simplification, we consider that coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value that obeys the reliability rule (a value we denote by $ P_{\phi}(S) $).
-If $ P_{\phi}(S) \geq P_{\phi}^{MIN} $, meaning that it also obeys the minimum contract value rule, then S announces this value on the market : $ P_{S}^{CRCT} = P_{\phi}(S) $.
+If $ P_{\phi}(S) \geq P^{MIN} $, meaning that it also obeys the minimum contract value rule, then S announces this value on the market : $ P_{S}^{CRCT} = P_{\phi}(S) $.
 Otherwise, coalition S is not able to enter.
@@
 		\forall t,\ Pr[ P_{S}(t) \leq P_{\phi}(S)] \leq \phi\ \textit{{\scriptsize (reliability rule)}} \\
-		and\ P_{\phi}(S) \geq P_{\phi}^{MIN}\ \textit{{\scriptsize (min value rule)}}
+		and\ P_{\phi}(S) \geq P^{MIN}\ \textit{{\scriptsize (min value rule)}}
 
@@
 \begin{equation}
-\tau_{i}(S) = \dfrac{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) - \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S-\{i\}) }{\mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S)}
+\tau_{i}(S) = \dfrac{\mathcal{U}_{\phi,\ P^{MIN}}(S) - \mathcal{U}_{\phi,\ P^{MIN}}(S-\{i\}) }{\mathcal{U}_{\phi,\ P^{MIN}}(S)}
 \label{tau}
@@
 
-At this point, we have three degrees of freedom : the reliability ($ \phi $), the required power to enter the market ($P_{\phi}^{MIN}$), and the prunning parameter ($\epsilon$).
+At this point, we have three degrees of freedom : the reliability ($ \phi $), the required power to enter the market ($P^{MIN}$), and the prunning parameter ($\epsilon$).
 For convenience, we introduce the number of desired coalitions ($ N_{COAL} $) and we fix $\epsilon $ as the smallest value such that the $ \epsilon $-graph contains at least $ N_{COAL} $ non overlaping cliques.
@@
  \centering
-  \subcaptionbox{$ (\phi, P_{\phi}^{MIN}) $ space\label{fig2:a}}{\includegraphics[width=1.6in]{./figure1/util}}\hspace{1em}%
-  \subcaptionbox{$ (N_{COAL}, P_{\phi}^{MIN}) $ space\label{fig2:b}}{\includegraphics[width=1.6in]{./figure6/util2}}
-\caption{Utility function in parameters space : In $ (\phi, P_{\phi}^{MIN}) $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ (N_{COAL}, P_{\phi}^{MIN}) $ for a fixed $ \phi $ (fig \ref{fig2:b})}
+  \subcaptionbox{$ (\phi, P^{MIN}) $ space\label{fig2:a}}{\includegraphics[width=1.6in]{./figure1/util}}\hspace{1em}%
+  \subcaptionbox{$ (N_{COAL}, P^{MIN}) $ space\label{fig2:b}}{\includegraphics[width=1.6in]{./figure6/util2}}
+\caption{Utility function in parameters space : In $ (\phi, P^{MIN}) $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ (N_{COAL}, P^{MIN}) $ for a fixed $ \phi $ (fig \ref{fig2:b})}
 \label{Fig2}
@@
 
-As visible on figure \ref{fig2:a}, the parameters $\phi$ and $P_{\phi}^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P_{\phi}^{MIN}$ could potentially lead to valid coalitions (and positive utility values).
-Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P_{\phi}^{MIN}$.
+As visible on figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P^{MIN}$ could potentially lead to valid coalitions (and positive utility values).
+Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P^{MIN}$.
 Obviously, the highest utility values are found for high $ \phi $, because coalitions are able then to announce higher contract values, yielding higher utilities.
 
-In the following, we fix the reliability to a given empirical value ($\phi = Cste $) and observe how the coalitions evolves for different values of $P_{\phi}^{MIN}$.
+In the following, we fix the reliability to a given empirical value ($\phi = Cste $) and observe how the coalitions evolves for different values of $P^{MIN}$.
 Note that the opposite is also possible although a little less intuitive, but not shown here because of space limitations and because it leads to similar conclusions.
 
-On figure \ref{fig2:b}, we can see that when the grid constraints are fixed ($ \phi = 0.08 $ and $ P_{\phi}^{MIN} = 2 $ for instance) and only the number of coalitions varies, there is an increase in social welfare (sum of coalitions utilities) up to a maximum point (for $ N_{COAL} = 3 $ here) before a decrease.
+On figure \ref{fig2:b}, we can see that when the grid constraints are fixed ($ \phi = 0.08 $ and $ P^{MIN} = 2 $ for instance) and only the number of coalitions varies, there is an increase in social welfare (sum of coalitions utilities) up to a maximum point (for $ N_{COAL} = 3 $ here) before a decrease.
 The reason is that increasing the number of coalitions allows more coalitions to reach stability and enter the market, but there is a point where nodes bringing stability are not sufficient inside the coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility.
 Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs where information is flooded, meaning that the algorithm performances also decrease.
-Naturally, it is also visible that small values of $P_{\phi}^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.
+Naturally, it is also visible that small values of $P^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.
 
@@
   \includegraphics[scale=0.8]{./figure5/coals}
-  \caption{Evolution of the coalitions for different values of $ P_{\phi}^{MIN} $ for clique percolation (blue dots) and random process (red dots).
+  \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots).
 The diameters of the dots are proportional to the corresponding coalition sizes.}
@@
 For our 200 prosumers, we suppose that the grid fixes a reliability $ \phi = 0.1 $ (meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time).
-We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P_{\phi}^{MIN} $ requirement.
+We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P^{MIN} $ requirement.
 As a comparison, we use a completely random algorithm that only asks for a number of coalitions and partitions the agents in a random fashion.
@@
 The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility.
-The $ P_{\phi}^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.
+The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.
 
 Looking at figure \ref{Fig3}, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements.
-The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither too low or too high ($ P_{\phi}^{MIN} = 5 $ in figure \ref{Fig3}).
+The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither too low or too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).
 
 In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario.
-As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P_{\phi}^{MIN} $.
+As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
 For consistency, we average the results of both plots of the random procedure over 100 realizations, and the errorbars in the plots stands for the standard deviations of the results.
@@
 
-When the grid requirements vary, for very low $ P_{\phi}^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
-But as $ P_{\phi}^{MIN} $ increases, we see the percentage of the correlated coalitions crashing down very quickly.
-After a few increases in $ P_{\phi}^{MIN} $, the percentage of the random procedure starts droping while it stays almost constant for our algorithm.
-For $ P_{\phi}^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm.
+When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
+But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions crashing down very quickly.
+After a few increases in $ P^{MIN} $, the percentage of the random procedure starts droping while it stays almost constant for our algorithm.
+For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm.
 Finally, when the grid requirement becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).

### 2014-09-15 Article.tex
update Nico

@@
 In this model, we consider that only the coalitions that have this probability value sufficiently low are acceptable.
-That this, we denote by $ \phi \in [0,1] $ the threshold imposed to the coalitions over this quantity, and we call it the \textbf{reliability} in the following.
+That is, we denote by $ \phi \in [0,1] $ the threshold imposed to the coalitions over this quantity, and we call it the \textbf{reliability} in the following.
 The reliability stipulates then that for any coalition S willing to join the market, the probability of S value (at any instant t) being below its contract value should at most be $ \phi $.

### 2014-09-17 Article.tex
update Vincent + Monique

@@
 
-As visible on figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P^{MIN}$ could potentially lead to valid coalitions (and positive utility values).
-Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and valid ones can arise for a larger spectrum of $ P^{MIN}$.
-Obviously, the highest utility values are found for high $ \phi $, because coalitions are able then to announce higher contract values, yielding higher utilities.
+As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P^{MIN}$.
+This could potentially lead to valid coalitions (and positive utility values).
+Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$.
+Obviously, the highest utility values are found for high $ \phi $, because the coalitions formed are able then to announce higher contract values, yielding higher utilities.
 
-In the following, we fix the reliability to a given empirical value ($\phi = Cste $) and observe how the coalitions evolves for different values of $P^{MIN}$.
-Note that the opposite is also possible although a little less intuitive, but not shown here because of space limitations and because it leads to similar conclusions.
+In figure \ref{fig2:b}, we fix the reliability value to a given empirical value ($\phi = 0.08 $).
+We can observe how the social welfare $\sum_{i \in S} \tau_i(S)$ evolves according to $\phi$ and $P^{MIN}$.
+In our setting we can observe  a maximum point for $ N_{COAL} = 3 $, followed a decrease.
+There is clearly a maximum number of coalitions sustainable for given set of parameters.
+After that point we will form more coalitions that would be unable to pass the grid requirements.
+Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs, meaning that the algorithm performance will also decrease.
 
-On figure \ref{fig2:b}, we can see that when the grid constraints are fixed ($ \phi = 0.08 $ and $ P^{MIN} = 2 $ for instance) and only the number of coalitions varies, there is an increase in social welfare (sum of coalitions utilities) up to a maximum point (for $ N_{COAL} = 3 $ here) before a decrease.
-The reason is that increasing the number of coalitions allows more coalitions to reach stability and enter the market, but there is a point where nodes bringing stability are not sufficient inside the coalitions to make them pass the grid requirements, and some coalitions start to fail with zero utility.
-Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs where information is flooded, meaning that the algorithm performances also decrease.
-Naturally, it is also visible that small values of $P^{MIN}$ lead to higher utilities because coalitions are able to announce higher contract values.
-
-\begin{center}
-\begin{figure}[b]
+\begin{figure}[htbp]
+  \centering
   \includegraphics[scale=0.8]{./figure5/coals}
@@
 \end{figure}
-\end{center}
 
-For our 200 prosumers, we suppose that the grid fixes a reliability $ \phi = 0.1 $ (meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time).
-We also fix the number of coalitions to 10 and have a look at how these coalitions evolve as the grid changes the $ P^{MIN} $ requirement.
-As a comparison, we use a completely random algorithm that only asks for a number of coalitions and partitions the agents in a random fashion.
+In figure \ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time.
+We observe the result of our algorithm for a fixed number of coalitions to $N_{coal} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $.
+We compare our algorithm with a random process that forms coalition.
 Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots).
@@
 
-Looking at figure \ref{Fig3}, we can see first that our algorithm seems to make generally a better job at finding high valued coalitions, but also that the results seem far more robust against how the grid positions its requirements.
-The blue dots allowed to enter the market seem indeed to outnumber the red dots, especially when the grid requirements are neither too low or too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).
+Looking at figure \ref{Fig3}, we can see first that our algorithm performs better at finding high valued coalitions.
+The blue dots allowed to enter the market are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).
 
@@
 As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
-For consistency, we average the results of both plots of the random procedure over 100 realizations, and the errorbars in the plots stands for the standard deviations of the results.
+For consistency, we average the results of both plots of the random procedure over 100 realizations with 95\% of confidence interval.
 
-When the grid requirements are constant (figure \ref{fig4:b}), and the number of desired coalitions is low, clique percolation generally performs only a little better than random.
-Nevetheless, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidely while clique percolation social welfare decreases slowly.
+When the grid requirements are constant (figure \ref{fig4:b}), and the number of desired coalitions is low, clique percolation generally performs better than random case.
+Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed.
 
 When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
-But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions crashing down very quickly.
-After a few increases in $ P^{MIN} $, the percentage of the random procedure starts droping while it stays almost constant for our algorithm.
+But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions is going down very quickly.
+After a few increases in $ P^{MIN} $, the percentage of the random procedure starts dropping while it stays almost constant for our algorithm.
 For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm.
-Finally, when the grid requirement becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).
+Finally, when the grid requirements becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).
 

### 2014-09-17 Article.tex
update caption

@@
  \centering
-  \subcaptionbox{$ (\phi, P^{MIN}) $ space\label{fig2:a}}{\includegraphics[width=1.6in]{./figure1/util}}\hspace{1em}%
-  \subcaptionbox{$ (N_{COAL}, P^{MIN}) $ space\label{fig2:b}}{\includegraphics[width=1.6in]{./figure6/util2}}
-\caption{Utility function in parameters space : In $ (\phi, P^{MIN}) $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ (N_{COAL}, P^{MIN}) $ for a fixed $ \phi $ (fig \ref{fig2:b})}
+  \subcaptionbox{$ (\phi, P^{MIN}) $ subspace\label{fig2:a}}{\includegraphics[width=1.7in]{./figure1/util}}
+  \subcaptionbox{$ (N_{COAL}, P^{MIN}) $ subspace\label{fig2:b}}{\includegraphics[width=1.70in]{./figure6/util2}}
+\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: In $ \{\phi, P^{MIN}\} $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $ (fig \ref{fig2:b})}
 \label{Fig2}

### 2014-09-18 Article.tex
update captions and few modif in V

@@
 \begin{equation}
- \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} 
+ \mathcal{U}_{\phi,\ P^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} 
 \label{utility}
@@
   \subcaptionbox{$ (N_{COAL}, P^{MIN}) $ subspace\label{fig2:b}}{\includegraphics[width=1.70in]{./figure6/util2}}
-\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: In $ \{\phi, P^{MIN}\} $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $ (fig \ref{fig2:b})}
+\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: In $ \{\phi \in [0,0.6], P^{MIN}\} $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $ (fig \ref{fig2:b}).
+For readability, $ P^{MIN} $ is expressed in tenth of MW.}
 \label{Fig2}
@@
 
-As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P^{MIN}$.
-This could potentially lead to valid coalitions (and positive utility values).
+As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high.
+Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utility values).
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$.
@@
 In figure \ref{fig2:b}, we fix the reliability value to a given empirical value ($\phi = 0.08 $).
-We can observe how the social welfare $\sum_{i \in S} \tau_i(S)$ evolves according to $\phi$ and $P^{MIN}$.
-In our setting we can observe  a maximum point for $ N_{COAL} = 3 $, followed a decrease.
+We can observe how the social welfare $\sum_{S \in CS} \mathcal{U}_{\phi,\ P^{MIN}}(S)$ evolves according to $P^{MIN}$ and $ N_{COAL} = |CS| $.
+In our setting we can observe  a maximum point for $ N_{COAL} = 3 $, followed by a decrease.
 There is clearly a maximum number of coalitions sustainable for given set of parameters.
@@
   \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots).
-The diameters of the dots are proportional to the corresponding coalition sizes.}
+The diameters of the dots are proportional to the corresponding coalition sizes.
+The hatched zone corresponds to the \textit{"under-requirement space"}, and results in zero utility for any coalition in this area.}
   \label{Fig3}
@@
 The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility.
-The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.
+The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the \textit{"under-requirement space"}, meaning that whenever a coalition is in this zone, it has a null utility.
 
 Looking at figure \ref{Fig3}, we can see first that our algorithm performs better at finding high valued coalitions.
-The blue dots allowed to enter the market are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).
+The blue dots allowed to enter the market (non hatched zone) are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).
 
@@
 As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
-For consistency, we average the results of both plots of the random procedure over 100 realizations with 95\% of confidence interval.
+For consistency, we average the results of both plots of the random procedure over 100 realizations.
 

### 2014-09-18 Article.tex
update notation array (pas beau)

@@
 \usepackage{subcaption}
+\usepackage{array}
 
@@
 
+\newcolumntype{h}{>{\centering}p{40mm}<{}}
+
+\begin{center}
+\begin{figure}
+\scriptsize
+\begin{tabular}{|h|h|}
+  \hline
+  \textbf{Notations} & \textbf{Meaning} \tabularnewline
+  \hline 
+  $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \tabularnewline
+  $ P_{i}(t) \in \mathbb{R}\ \ (P_{S}(t)) $ & Instantaneous available production of agent i (coalition S) at time t \tabularnewline
+  $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time serie of agent i available production values \tabularnewline
+  $ \mathcal{P}_{i}\ \ (\mathcal{P}_{S}) $ & Probability distribution of agent i (coalition S) available production values \tabularnewline
+  $ \phi \in [0,1] $ & Reliability threshold (fixed by the grid operator) \tabularnewline
+  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold (fixed by the grid operator) \tabularnewline
+  $ P_{S}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition S \tabularnewline
+  $ P_{\phi}(S) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \tabularnewline
+  $ \mathcal{U}_{\phi,\ P^{MIN}}(S) $ & Utility function \tabularnewline
+  $ \tau_{i}(S) \in \mathbb{R}$ & Descision rule for mapping overlaping into disjoint coalitions \tabularnewline
+  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \tabularnewline
+  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition (entry of the algorithms) \tabularnewline
+  $ CS = \{S_{1},...,S_{N_{COAL}}\} $ & Coalition structure \tabularnewline
+  \hline
+\end{tabular}
+\caption{Model main notations}
+\label{Fig89}
+\end{figure}
+\end{center}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-09-18 Article.tex
update notation array (pas beau)

@@
 They may even occur some penalties if the productions deviate from the initial contract values.
-However, it remains pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy sources) will impact its net gain, and for the power grid, it will helps to maintain its reliability.
+However, it remains pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy sources) will impact its net gain, and for the power grid, it will help to maintain its reliability.
 
@@
 We then consider another constraint denoted by $ P^{MIN} $, as the \textbf{minimum contract value}, which states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P^{MIN} $) will be accepted.
+Figure \ref{Fig89} summurizes the main notations of this paper.
 
@@
 What this paper investigates in the following is the developement of an algorithm that organizes prosumers such that the synergy term of the standard deviation ( $ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) is minimized.
-In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with low relative standard deviation, e.g high utility coalitions, can be computed.
+In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with low relative standard deviations, e.g high utility coalitions, can be computed.
 
@@
 We can observe how the social welfare $\sum_{S \in CS} \mathcal{U}_{\phi,\ P^{MIN}}(S)$ evolves according to $P^{MIN}$ and $ N_{COAL} = |CS| $.
-In our setting we can observe  a maximum point for $ N_{COAL} = 3 $, followed by a decrease.
-There is clearly a maximum number of coalitions sustainable for given set of parameters.
+In our setting we can observe  a maximum point around $ N_{COAL} = 3 $, followed by a decrease.
+There is clearly a maximum number of coalitions sustainable for a given set of parameters.
 After that point we will form more coalitions that would be unable to pass the grid requirements.
@@
 In figure \ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time.
-We observe the result of our algorithm for a fixed number of coalitions to $N_{coal} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $.
-We compare our algorithm with a random process that forms coalition.
+We observe the result of our algorithm for a fixed number of coalitions of $N_{coal} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $.
+We compare our algorithm with a random process that splits the agents in $ N_{COAL} $ coalitions.
 Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots).

### 2014-09-18 figure4_with_green_curve/Python/plot_fig.py
figure 4 Python plot code

@@ -1,0 +1,60 @@
+import json
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+
+#matplotlib config
+def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
+    import matplotlib as mpl
+    cmap = cm.jet
+    mpl.rcParams['axes.labelsize'] = 9
+    mpl.rcParams['xtick.labelsize'] = 9
+    mpl.rcParams['ytick.labelsize'] = 9
+    mpl.rcParams['legend.fontsize'] = 9
+    #rcParams['font.family'] = 'serif'
+    #rcParams['font.serif'] = ['Computer Modern Roman']
+    #mpl.rcParams['text.usetex'] = True
+    mpl.rcParams['figure.figsize'] = 7.3, 4.2
+    mpl.rcParams['axes.labelweight'] = 'bold'
+    mpl.rcParams['font.weight'] = 'bold'
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['lines.markersize'] = 8
+    mpl.rcParams['legend.numpoints'] = 1
+    # figure size in inch
+    mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
+    # figure dots per inch
+    mpl.rcParams['figure.dpi'] = 300
+    mpl.rcParams['figure.autolayout'] = True
+    
+#Load curve data
+perco = []; rand = []; corre = []; std_perco = []; std_rand = []; std_corre = []
+with open('.\percentage_percolation.json', "r") as fichier:
+    perco = json.load(fichier)
+with open('.\std_percolation.json', "r") as fichier:
+    std_perco = json.load(fichier)
+with open('.\percentage_random.json', "r") as fichier:
+    rand = json.load(fichier)
+with open('.\std_random.json', "r") as fichier:
+    std_rand = json.load(fichier)
+with open('.\percentage_correlated.json', "r") as fichier:
+    corre = json.load(fichier)
+with open('.\std_correlated.json', "r") as fichier:
+    std_corre = json.load(fichier)	
+
+#make the plot
+matplotlib_setup(4,3)
+Pmax = 12; phi = 0.15; nb_coal = 7
[…]

### 2014-09-19 figure7_with_green_curve_v1/Python/plot_fig.py
figure 7 Python plot code

@@ -1,0 +1,61 @@
+import json
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+
+#matplotlib config
+def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
+    import matplotlib as mpl
+    cmap = cm.jet
+    mpl.rcParams['axes.labelsize'] = 9
+    mpl.rcParams['xtick.labelsize'] = 9
+    mpl.rcParams['ytick.labelsize'] = 9
+    mpl.rcParams['legend.fontsize'] = 9
+    #rcParams['font.family'] = 'serif'
+    #rcParams['font.serif'] = ['Computer Modern Roman']
+    #mpl.rcParams['text.usetex'] = True
+    mpl.rcParams['figure.figsize'] = 7.3, 4.2
+    mpl.rcParams['axes.labelweight'] = 'bold'
+    mpl.rcParams['font.weight'] = 'bold'
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['lines.markersize'] = 8
+    mpl.rcParams['legend.numpoints'] = 1
+    # figure size in inch
+    mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
+    # figure dots per inch
+    mpl.rcParams['figure.dpi'] = 300
+    mpl.rcParams['figure.autolayout'] = True
+    
+#Load curve data
+perco = []; rand = []; corre = []; std_perco = []; std_rand = []; std_corre = []
+with open('.\percentage_percolation.json', "r") as fichier:
+    perco = json.load(fichier)
+with open('.\std_percolation.json', "r") as fichier:
+    std_perco = json.load(fichier)
+with open('.\percentage_random.json', "r") as fichier:
+    rand = json.load(fichier)
+with open('.\std_random.json', "r") as fichier:
+    std_rand = json.load(fichier)
+with open('.\percentage_correlated.json', "r") as fichier:
+    corre = json.load(fichier)
+with open('.\std_correlated.json', "r") as fichier:
+    std_corre = json.load(fichier)	
+
+#make the plot
+matplotlib_setup(4,3)
+Pmax = 12; phi = 0.15
+
+axes = plt.gca()
[…]

### 2014-09-22 figure8/plot_fig.py
another design for last figure

@@ -1,0 +1,118 @@
+import json
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+
+#matplotlib config
+def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
+    import matplotlib as mpl
+    cmap = cm.jet
+    mpl.rcParams['axes.labelsize'] = 9
+    mpl.rcParams['xtick.labelsize'] = 9
+    mpl.rcParams['ytick.labelsize'] = 9
+    mpl.rcParams['legend.fontsize'] = 9
+    #rcParams['font.family'] = 'serif'
+    #rcParams['font.serif'] = ['Computer Modern Roman']
+    mpl.rcParams['text.usetex'] = True
+    mpl.rcParams['figure.figsize'] = 7.3, 4.2
+    mpl.rcParams['axes.labelweight'] = 'bold'
+    mpl.rcParams['font.weight'] = 'bold'
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['lines.markersize'] = 8
+    mpl.rcParams['legend.numpoints'] = 1
+    # figure size in inch
+    mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
+    # figure dots per inch
+    mpl.rcParams['figure.dpi'] = 300
+    mpl.rcParams['figure.autolayout'] = True
+	
+#Load curve data subplot a
+perco = []; rand = []; corre = []; std_perco = []; std_rand = []; std_corre = []
+with open('fig_percentage\percentage_percolation.json', "r") as fichier:
+    perco = json.load(fichier)
+with open('fig_percentage\std_percolation.json', "r") as fichier:
+    std_perco = json.load(fichier)
+with open('fig_percentage\percentage_random.json', "r") as fichier:
+    rand = json.load(fichier)
+with open('fig_percentage\std_random.json', "r") as fichier:
+    std_rand = json.load(fichier)
+with open('fig_percentage\percentage_correlated.json', "r") as fichier:
+    corre = json.load(fichier)
+with open('fig_percentage\std_correlated.json', "r") as fichier:
[…]

### 2014-09-22 Article.tex
update table and figure 8

@@
 
-\newcolumntype{h}{>{\centering}p{40mm}<{}}
-
-\begin{center}
-\begin{figure}
+\begin{table}[!t]
+\label{Fig89}
+\centering
 \scriptsize
-\begin{tabular}{|h|h|}
+\begin{tabular}{l|p{4.5cm}}
   \hline
-  \textbf{Notations} & \textbf{Meaning} \tabularnewline
+  \textbf{Symbols} &  \\
   \hline 
-  $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \tabularnewline
-  $ P_{i}(t) \in \mathbb{R}\ \ (P_{S}(t)) $ & Instantaneous available production of agent i (coalition S) at time t \tabularnewline
-  $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time serie of agent i available production values \tabularnewline
-  $ \mathcal{P}_{i}\ \ (\mathcal{P}_{S}) $ & Probability distribution of agent i (coalition S) available production values \tabularnewline
-  $ \phi \in [0,1] $ & Reliability threshold (fixed by the grid operator) \tabularnewline
-  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold (fixed by the grid operator) \tabularnewline
-  $ P_{S}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition S \tabularnewline
-  $ P_{\phi}(S) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \tabularnewline
-  $ \mathcal{U}_{\phi,\ P^{MIN}}(S) $ & Utility function \tabularnewline
-  $ \tau_{i}(S) \in \mathbb{R}$ & Descision rule for mapping overlaping into disjoint coalitions \tabularnewline
-  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \tabularnewline
-  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition (entry of the algorithms) \tabularnewline
-  $ CS = \{S_{1},...,S_{N_{COAL}}\} $ & Coalition structure \tabularnewline
+  $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
+  $ P_{i}(t) \in \mathbb{R}\ \ (P_{S}(t)) $ & Instantaneous production of agent $i$ in coalition $S$ at time $t$ \\
+  $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time series of agent i available production values \\
+  $ \mathcal{P}_{i}\ \ (\mathcal{P}_{S}) $ & Probability distribution of the production of the agent i in coalition S\\
+  $ \phi \in [0,1] $ & Reliability threshold \\
+  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
+  $ P_{S}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition S \\
+  $ P_{\phi}(S) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
+  $ \mathcal{U}_{\phi,\ P^{MIN}}(S) $ & Utility function \\
+  $ \tau_{i}(S) \in \mathbb{R}$ & Decision rule for mapping overlapping into disjoint coalitions \\
+  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
+  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
+  $ CS = \{S_{1},...,S_{N_{COAL}}\} $ & Coalition structure \\
   \hline
 \end{tabular}
-\caption{Model main notations}
-\label{Fig89}
-\end{figure}
-\end{center}
+\caption{Notations}
+\end{table}
 

### 2014-09-22 figure8/plot_fig.py

@@ -16,9 +16,9 @@ def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
     mpl.rcParams['figure.figsize'] = 7.3, 4.2
     mpl.rcParams['axes.labelweight'] = 'bold'
     mpl.rcParams['font.weight'] = 'bold'
-    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['axes.linewidth'] = 0.75
     mpl.rcParams['lines.linewidth'] = 2
-    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['axes.linewidth'] = 0.75
     mpl.rcParams['lines.linewidth'] = 2
     mpl.rcParams['lines.markersize'] = 8
     mpl.rcParams['legend.numpoints'] = 1
@@ -27,7 +27,7 @@ def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
     # figure dots per inch
     mpl.rcParams['figure.dpi'] = 300
     mpl.rcParams['figure.autolayout'] = True
-	
+
 #Load curve data subplot a
 perco = []; rand = []; corre = []; std_perco = []; std_rand = []; std_corre = []
 with open('fig_percentage\percentage_percolation.json', "r") as fichier:
@@ -41,8 +41,8 @@ with open('fig_percentage\std_random.json', "r") as fichier:
 with open('fig_percentage\percentage_correlated.json', "r") as fichier:
     corre = json.load(fichier)
 with open('fig_percentage\std_correlated.json', "r") as fichier:
-    std_corre = json.load(fichier)	
-	
+    std_corre = json.load(fichier)
+
 #Transform data
 perco_sup = [a+b for a,b in zip(perco,std_perco)]
 perco_inf = [a-b for a,b in zip(perco,std_perco)]
@@ -64,8 +64,8 @@ with open('fig_SW\std_random.json', "r") as fichier:
 with open('fig_SW\percentage_correlated.json', "r") as fichier:
     corre2 = json.load(fichier)
 with open('fig_SW\std_correlated.json', "r") as fichier:
-    std_corre2 = json.load(fichier)	
-	
+    std_corre2 = json.load(fichier)
+
 #Transform data
 perco_sup2 = [a+b for a,b in zip(perco2,std_perco2)]
[…]

### 2014-09-22 figure8/plot_fig.py
update the figure 8

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import matplotlib.cm as cm
 
 #matplotlib config
-def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
+def matplotlib_setup(figsize_x=15, figsize_y=5):
     import matplotlib as mpl
     cmap = cm.jet
     mpl.rcParams['axes.labelsize'] = 9
@@ -13,7 +13,6 @@ def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
     #rcParams['font.family'] = 'serif'
     #rcParams['font.serif'] = ['Computer Modern Roman']
     mpl.rcParams['text.usetex'] = True
-    mpl.rcParams['figure.figsize'] = 7.3, 4.2
     mpl.rcParams['axes.labelweight'] = 'bold'
     mpl.rcParams['font.weight'] = 'bold'
     mpl.rcParams['axes.linewidth'] = 0.75
@@ -26,21 +25,21 @@ def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
     mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
     # figure dots per inch
     mpl.rcParams['figure.dpi'] = 300
-    mpl.rcParams['figure.autolayout'] = True
+    #mpl.rcParams['figure.autolayout'] = True
 
 #Load curve data subplot a
 perco = []; rand = []; corre = []; std_perco = []; std_rand = []; std_corre = []
-with open('fig_percentage\percentage_percolation.json', "r") as fichier:
+with open('fig_percentage/percentage_percolation.json', "r") as fichier:
     perco = json.load(fichier)
-with open('fig_percentage\std_percolation.json', "r") as fichier:
+with open('fig_percentage/std_percolation.json', "r") as fichier:
     std_perco = json.load(fichier)
-with open('fig_percentage\percentage_random.json', "r") as fichier:
+with open('fig_percentage/percentage_random.json', "r") as fichier:
     rand = json.load(fichier)
-with open('fig_percentage\std_random.json', "r") as fichier:
+with open('fig_percentage/std_random.json', "r") as fichier:
     std_rand = json.load(fichier)
-with open('fig_percentage\percentage_correlated.json', "r") as fichier:
[…]

### 2014-09-23 Article.tex
update table 1
update

@@
 Due to the volatile nature of DER productions or user consumptions, the available power of a coalition will oscillates around $ P_{S}^{CRCT} $.
-It will be the responsability of the coalition, that is the information system maintaining the coalition, to compensate for deviations.
+It will be the responsibility of the coalition, that is the information system maintaining the coalition, to compensate for deviations.
 In this paper, we consider then that the smaller these deviations, the better.
@@
 We then consider another constraint denoted by $ P^{MIN} $, as the \textbf{minimum contract value}, which states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P^{MIN} $) will be accepted.
-Figure \ref{Fig89} summurizes the main notations of this paper.
+In table \ref{table1} summarizes the main notations of this paper.
 
@@
 \begin{table}[!t]
-\label{Fig89}
 \centering
@@
 \begin{tabular}{l|p{4.5cm}}
-  \hline
-  \textbf{Symbols} &  \\
   \hline 
@@
 \end{tabular}
-\caption{Notations}
+\caption{Notations\label{table1}}
 \end{table}

### 2014-09-23 Article.tex
A few corrections

--- 2014-09-22 Article.tex
@@
 One of the key ideas in the smart grid revolves around the introduction of communication means inside the power grids.
-This could enable complex improvements in the energy management and lead progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
+This could enable complex improvements in the energy management and leads progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
 Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge only in remote farms, but also in residential areas.
@@
 The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}.
-These are just a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.
+These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.
 
@@
 
-In section \ref{sec:model}, we consider that the agent's production depend on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, and their appliances (loads).
+In section \ref{sec:model}, we consider that the agent's production depends on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, and their appliances (loads).
 We use meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
@@
 In \cite{Garas2008, Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
-However, there is no well defined method to select of right tradeoff $ \epsilon $ as function of the network topology.
+However, there is no well defined method to select a right tradeoff $ \epsilon $ as a function of the network topology.
 
@@
 When the algorithm stops, some coalitions may be partially overlapping, meaning that a node can belong to multiple communities.
-Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities in the mean time.
+Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities.
 
@@
 During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
-As expected (because we introduced them), the time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both of their energetic mix and habits of a given prosumer.
+As expected (because we introduced them), the time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both of the energetic mix and the habits of a given prosumer.
 Nevertheless, these macro variations can hide the interesting information in the correlation coefficients and twist the rest of our process.
@@
 We then consider another constraint denoted by $ P^{MIN} $, as the \textbf{minimum contract value}, which states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P^{MIN} $) will be accepted.
-Figure \ref{Fig89} summurizes the main notations of this paper.
+Table \ref{Fig89} summurizes the main notations of this paper.
 
@@
   \hline
-  \textbf{Symbols} &  \\
+  \multicolumn{2}{l}{\textbf{Symbols}} \\
   \hline 
   $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
-  $ P_{i}(t) \in \mathbb{R}\ \ (P_{S}(t)) $ & Instantaneous production of agent $i$ in coalition $S$ at time $t$ \\
+  $ P_{i}(t) \in \mathbb{R}\ \ (P_{S}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
   $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time series of agent i available production values \\
-  $ \mathcal{P}_{i}\ \ (\mathcal{P}_{S}) $ & Probability distribution of the production of the agent i in coalition S\\
+  $ \mathcal{P}_{i}\ \ (\mathcal{P}_{S}) $ & Probability distribution of the production of the agent i \\
   $ \phi \in [0,1] $ & Reliability threshold \\
@@
 \caption{Notations}
+\label{Fig89}
 \end{table}
@@
 
-In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario.
-As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
+In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario.
+As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
 For consistency, we average the results of both plots of the random procedure over 100 realizations.
 
-When the grid requirements are constant (figure \ref{fig4:b}), and the number of desired coalitions is low, clique percolation generally performs better than random case.
+When the grid requirements are constant (figure \ref{Fig4}b), and the number of desired coalitions is low, clique percolation generally performs better than random case.
 Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed.
@@
 
+%\begin{figure}
+% \centering
+%  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{figure7_with_green_curve_v1/uti5}}\hspace{1em}%
+%  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{figure4_with_green_curve/survival2_old}}
+%\caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.
+Blue curves represent clique percolation, red curves, the random process, and the green ones, correlated coalitions.}
+%\label{Fig4}
+%\end{figure}
+
 \begin{figure}
  \centering
-  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{figure7_with_green_curve_v1/uti5}}\hspace{1em}%
-  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{figure4_with_green_curve/survival2_old}}
+  \includegraphics[scale=0.35]{figure8/fig8.pdf}
   \caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.

### 2014-09-23 Article.tex

@@
 We then consider another constraint denoted by $ P^{MIN} $, as the \textbf{minimum contract value}, which states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P^{MIN} $) will be accepted.
-Table \ref{Fig89} summurizes the main notations of this paper.
+In table \ref{table1} summarizes the main notations of this paper.
 
@@
   \hline 
-  \multicolumn{2}{l}{\textbf{Symbols}} \\
-  \hline 
   $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
@@
 \end{tabular}
-\caption{Notations}
-\label{Fig89}
+
+\caption{Notations\label{table1}}
 \end{table}

### 2014-09-23 Article.tex
A few corrections
optimization problem equation and other updates

@@
 \begin{abstract}
-In a smart grid environment, we study coalition formation of prosumers that aim at selling energy to the grid.
-It is paramount for the grid operation that the energy producers are able to sustain the grid demand in terms of stability and minimum production requirement in order to enter the energy market.
+In a smart grid environment, we study coalition formation of prosumers that aim at entering the energy market.
+It is paramount for the grid operation that the energy producers are able to sustain the grid demand in terms of stability and minimum production requirement.
 We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minimum energy level for the coalitions and a steady production level which leads to finding uncorrelated sources of energy  to form a coalition.
@@
 
-There seems to have a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as most of the demand side management concepts have proposed such an architecture \cite{WuHamedHuangBook2011}.
-Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and in the mean time satisfy the multiple requirements of the power grid management.
+There seems to have a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}.
+Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
 The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}.
@@
 As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering traces for these stations.
-In this paper, we consider that prosumers can only produce through wind turbines and photovoltaic panels (PV).
+We consider here that prosumers can only produce through wind turbines and photovoltaic panels (PV).
 On the other hand, we built their consumption patterns according to two major cycles :
@@
 During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
-As expected (because we introduced them), the time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both of the energetic mix and the habits of a given prosumer.
+As expected, the prosumer's time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both on the energetic mix and the habits.
 Nevertheless, these macro variations can hide the interesting information in the correlation coefficients and twist the rest of our process.
 Thereby, we remove these seasonal effects for all agents (see second block of figure \ref{Fig1}).
-We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent i, and by $ \mathcal{P}_{i} $ the corresponding probability distribution.
+We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent i, and by $ \mathcal{F}(P_{i}) $ the corresponding probability distribution.
 
-We now extend these notations to any coalition $ S \subset \mathcal{A} $ : 
+We now extend these notations to any coalition $ s \subset \mathcal{A} $ : 
 \begin{itemize}
-\item $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $
-\item $ \mathcal{T}_{S} = \{ P_{S}(t_{0}),...,P_{S}(t_{K}) \} $ with $ \mathcal{P}_{S} $ the corresponding probability distribution
-\item A coalition S has the possibility to announce a contract value $ P_{S}^{CRCT} $ on the market.
+\item $ P_{s}(t) = \sum_{i \in s} P_{i}(t) $
+\item $ \mathcal{T}_{s} = \{ P_{s}(t_{0}),...,P_{s}(t_{K}) \} $ with $ \mathcal{F}(P_{s}) $ the corresponding probability distribution
+\item A coalition S has the possibility to announce a contract value $ P_{s}^{CRCT} $ on the market.
 \end{itemize}
 
-Due to the volatile nature of DER productions or user consumptions, the available power of a coalition will oscillates around $ P_{S}^{CRCT} $.
-It will be the responsibility of the coalition, that is the information system maintaining the coalition, to compensate for deviations.
-In this paper, we consider then that the smaller these deviations, the better.
-Based on the $ \mathcal{T}_{S} $ distribution, the probability of S value (at any instant t) being below its contract value $ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] $ appears therefore as a quantity of interest that we would want to keep as low as possible.
+Due to the volatile nature of DER productions or user consumptions, the available power of a coalition will oscillates around $ P_{s}^{CRCT} $.
+It is of the responsibility of the information system that forms coalitions to select agents in order to provide a production as stable as possible.
+Based on the $ \mathcal{F}(P_{s}) $ distribution, the probability of s value being below its contract value $ Pr[P_{s} \leq P_{s}^{CRCT} ] $ appears therefore as a quantity of interest that we would want to keep as low as possible.
 
-In this model, we consider that only the coalitions that have this probability value sufficiently low are acceptable.
-That is, we denote by $ \phi \in [0,1] $ the threshold imposed to the coalitions over this quantity, and we call it the \textbf{reliability} in the following.
-The reliability stipulates then that for any coalition S willing to join the market, the probability of S value (at any instant t) being below its contract value should at most be $ \phi $.
-That is, $ \forall S,\ \forall t,\ Pr[P_{S}(t) \leq P_{S}^{CRCT} ] \leq \phi $ ($ \phi $ being restricted to small values for consistency).
+More formally, the rest of the paper derives a heuristic that aims at solving the following problem (Table \ref{table1} summarizes the main notations):
+ \begin{equation}
+  \min_{\substack{S \subset CS\\ |S| = N_{COAL} \\ 
+     \forall s \in S,\ |s| \neq 0 \\ P_{s}^{CRCT} \geq P^{MIN}}}
+  \Bigg\{ \sum_{s \in S} Pr[P_{s} \leq P_{s}^{CRCT}] \Bigg\}
+  \label{problem}
+\end{equation}
 
-We then consider another constraint denoted by $ P^{MIN} $, as the \textbf{minimum contract value}, which states that only coalitions with higher contract values ($ P_{S}^{CRCT} \geq P^{MIN} $) will be accepted.
-In table \ref{table1} summarizes the main notations of this paper.
+We explain now the two requirements that will help us solve this problem.
+Let $ \phi \in [0,1] $ be the \textbf{reliability} threshold imposed to the coalitions.
+It stipulates that for any coalition s willing to join the market, the probability of s value (at any instant t) being below its contract value should at most be $ \phi $.
+That is, $ \forall s \in 2^{N},\ Pr[P_{s} \leq P_{s}^{CRCT} ] \leq \phi $ ($ \phi $ being restricted to small values for consistency).
 
-At this point, if a coalition S wishes to join the market, it has to choose a contract value that fulfill the two conditions.
-For simplification, we consider that coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value that obeys the reliability rule (a value we denote by $ P_{\phi}(S) $).
-If $ P_{\phi}(S) \geq P^{MIN} $, meaning that it also obeys the minimum contract value rule, then S announces this value on the market : $ P_{S}^{CRCT} = P_{\phi}(S) $.
-Otherwise, coalition S is not able to enter.
-Basically, a coalition S is valid if and only if :
+We also consider the \textbf{minimum contract value} constraint denoted by $ P^{MIN} $, which states that only coalitions with higher contract values ($ P_{s}^{CRCT} \geq P^{MIN} $) will be accepted.
+
+At this point, if a coalition s wishes to join the market, it has to choose a contract value that fulfill the two conditions.
+For simplification, we consider that coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value that obeys the reliability rule (a value we denote by $ P_{\phi}(s) $).
+If $ P_{\phi}(s) \geq P^{MIN} $, meaning that it also obeys the minimum contract value rule, then s announces this value on the market : $ P_{s}^{CRCT} = P_{\phi}(s) $.
+Otherwise, coalition s is not able to enter.
+Basically, a coalition s is valid if and only if :
 \begin{equation}
 \left\{ \begin{array}{lll}
-		\forall t,\ Pr[ P_{S}(t) \leq P_{\phi}(S)] \leq \phi\ \textit{{\scriptsize (reliability rule)}} \\
-		and\ P_{\phi}(S) \geq P^{MIN}\ \textit{{\scriptsize (min value rule)}}
+		Pr[ P_{s} \leq P_{\phi}(s)] \leq \phi\ \textit{{\scriptsize (reliability rule)}} \\
+		and\ P_{\phi}(s) \geq P^{MIN}\ \textit{{\scriptsize (min value rule)}}
 
@@
 \subsection{Utility function}\label{subsec:UtilityFunc}
-We now choose a very simple utility function (eq.
+We now choose a utility function (eq.
 \ref{utility}) that derives directly from the above remarks.
@@
 Furthermore, it seems obvious that the utility should increase with the contract value.
-The $ 1/|S| $ term in eq.
-\ref{utility} indicates that we favorise small coalitions, mainly because they are easier to maintain in terms of communications.
+The $ 1/|s| $ term in eq.
+\ref{utility} indicates that we favor small coalitions, mainly because they are easier to maintain in terms of communications.
 
 \begin{equation}
- \mathcal{U}_{\phi,\ P^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} 
+ \mathcal{U}_{\phi,\ P^{MIN}}(s) = \mathbf{1}_{\textit{s\ valid}} \dfrac{P_{\phi}(s)}{|s|} 
 \label{utility}
@@
 
-In order to illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with gaussian value probability distributions $ \mathcal{P}_{i} = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{P}_{j} = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{P}_{ij} $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
+In order to illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
 
@@
 
-with $ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the reliability condition $ Pr[P_{ij}(t) \leq P_{\phi}(ij) ] \leq \phi $ as :
+with $ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the reliability condition $ Pr[P_{ij} \leq P_{\phi}(ij) ] \leq \phi $ as :
 
@@
 
-What this paper investigates in the following is the developement of an algorithm that organizes prosumers such that the synergy term of the standard deviation ( $ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) is minimized.
+What this paper investigates in the following is the developement of an algorithm that organizes prosumers such that the synergy term of the standard deviation ($ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) is minimized.
 In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with low relative standard deviations, e.g high utility coalitions, can be computed.
@@
   \hline 
+  \\
   $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
-  $ P_{i}(t) \in \mathbb{R}\ \ (P_{S}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
+  $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
   $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time series of agent i available production values \\
-  $ \mathcal{P}_{i}\ \ (\mathcal{P}_{S}) $ & Probability distribution of the production of the agent i \\
+  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of the production of the agent $i$ \\
   $ \phi \in [0,1] $ & Reliability threshold \\
   $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
-  $ P_{S}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition S \\
-  $ P_{\phi}(S) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
-  $ \mathcal{U}_{\phi,\ P^{MIN}}(S) $ & Utility function \\
-  $ \tau_{i}(S) \in \mathbb{R}$ & Decision rule for mapping overlapping into disjoint coalitions \\
+  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition s \\
+  $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
+  $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
+  $ \tau_{i}(s) \in \mathbb{R}$ & Decision rule for mapping overlapping into disjoint coalitions \\
   $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
   $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
-  $ CS = \{S_{1},...,S_{N_{COAL}}\} $ & Coalition structure \\
+  $ CS $ & Set of all coalition structures \\
+  \\
   \hline
@@
 
-The model we used in order to simulate timeseries of available production provides some diversity because of the combination between the energetic mix and the climate vectors.
+The model we used in order to simulate timeseries of available production provides some diversity because of the combination between the energetic mix (renewable generators combination) and the climate vectors.
 Nevertheless, as the number of agents grows, the timeseries tend to exhibit similar patterns.
@@
 Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ seems too brutal.
-Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(S) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore.
+Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore.
 
-Cliques for small values of $ \epsilon $ appear thus as good seeds for stable coalitions, but we still have to find a way to bring them above the grid requirements as well as increasing the social welfare, both by incorporating more nodes.
-This is exactly where clique percolation proves useful for our concerns.
-The main idea of making the seeds grow as long as it increases a fitness function is exactly what we need.
-The only difference being that our fitness function is simply chosen as the utility function.
+Cliques for small values of $ \epsilon $ appear thus as good seeds for stable coalitions.
+However we have to find a way to bring the coalitions above the grid requirements as well as increasing the social welfare by incorporating more nodes to the coalitions.
+Clique percolation algorithms have proven to be a useful heuristic for such complex problems.
+The main idea behind this algorithm is to make the seeds grow as long as they increase a fitness function based on internal and external degrees of the seeds \cite{Lancichinetti}.
+The main difference between our case study and that algorithm is that we chose this fitness function as our utility function (cf.
+eq.\eqref{utility}) that both maximizes the production and minimizes its volatility.
 
 As explained in section \ref{sec:related}, clique percolation leads generally to overlapping communities.
-For simplicity, in this paper, we wish to keep the coalitions separated and  leave the management of overlapping coalitions for future work.
-We thus implemented a simple heuristic that consider nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss.
-More formally, for a coalition S and a node $ i \in S $, we define :
+For simplicity, in this paper, we wish to keep the coalitions disjoint and  leave the management of overlapping coalitions for future work.
+We thus implemented a simple heuristic that considers nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss.
+More formally, for a coalition s and a node $ i \in s $, we define :
 
 \begin{equation}
-\tau_{i}(S) = \dfrac{\mathcal{U}_{\phi,\ P^{MIN}}(S) - \mathcal{U}_{\phi,\ P^{MIN}}(S-\{i\}) }{\mathcal{U}_{\phi,\ P^{MIN}}(S)}
+\tau_{i}(s) = \dfrac{\mathcal{U}_{\phi,\ P^{MIN}}(s) - \mathcal{U}_{\phi,\ P^{MIN}}(s-\{i\}) }{\mathcal{U}_{\phi,\ P^{MIN}}(s)}
 \label{tau}
@@
 
-If node i belongs to  multiple coalitions, the only coalition retaining node i is the one that maximizes $ \tau_{i} $.
+If node i belongs to  multiple coalitions, the only coalition including node i is the one that maximizes $ \tau_{i} $.
 
 At this point, we have three degrees of freedom : the reliability ($ \phi $), the required power to enter the market ($P^{MIN}$), and the prunning parameter ($\epsilon$).
-For convenience, we introduce the number of desired coalitions ($ N_{COAL} $) and we fix $\epsilon $ as the smallest value such that the $ \epsilon $-graph contains at least $ N_{COAL} $ non overlaping cliques.
+For convenience, we introduce the number of desired coalitions ($ N_{COAL} $) and we choose $\epsilon $ as the smallest value such that the $ \epsilon $-graph contains at least $ N_{COAL} $ non overlaping cliques.
 The overall utility depends now on the grid requirements and on the number of desired coalitions.
 
-The next section shows how the utility behave within the parameters space and provides the results of our 200 agents test.
+Next section shows the utility behavior within the parameters space and provides the results of our 200 agents test.
 
@@
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$.
-Obviously, the highest utility values are found for high $ \phi $, because the coalitions formed are able then to announce higher contract values, yielding higher utilities.
+Obviously, the highest utility values are found for high $ \phi $, because the coalitions formed are able to announce higher contract values, yielding higher utilities.
 
 In figure \ref{fig2:b}, we fix the reliability value to a given empirical value ($\phi = 0.08 $).
-We can observe how the social welfare $\sum_{S \in CS} \mathcal{U}_{\phi,\ P^{MIN}}(S)$ evolves according to $P^{MIN}$ and $ N_{COAL} = |CS| $.
+We can observe how the social welfare $\sum_{s \in S} \mathcal{U}_{\phi,\ P^{MIN}}(s)$ evolves according to $P^{MIN}$ and $ N_{COAL} $.
 In our setting we can observe  a maximum point around $ N_{COAL} = 3 $, followed by a decrease.
 There is clearly a maximum number of coalitions sustainable for a given set of parameters.
-After that point we will form more coalitions that would be unable to pass the grid requirements.
+After this value we will form more coalitions that would be unable to pass the grid requirements.
 Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs, meaning that the algorithm performance will also decrease.
@@
 In figure \ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time.
-We observe the result of our algorithm for a fixed number of coalitions of $N_{coal} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $.
+We observe the result of our algorithm for a fixed number of coalitions of $N_{COAL} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $.
 We compare our algorithm with a random process that splits the agents in $ N_{COAL} $ coalitions.
@@
 
-Looking at figure \ref{Fig3}, we can see first that our algorithm performs better at finding high valued coalitions.
+Looking at figure \ref{Fig3}, we can see first that our algorithm performs better than the random formation algorithm at finding high valued coalitions.
 The blue dots allowed to enter the market (non hatched zone) are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).

### 2014-09-24 Article.tex
update the eq1
update the section 4

@@
 \hyphenation{op-tical net-works semi-conduc-tor}
+
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
  \begin{equation}
-  \min_{\substack{S \subset CS\\ |S| = N_{COAL} \\ 
-     \forall s \in S,\ |s| \neq 0 \\ P_{s}^{CRCT} \geq P^{MIN}}}
-  \Bigg\{ \sum_{s \in S} Pr[P_{s} \leq P_{s}^{CRCT}] \Bigg\}
+  \argmin _{\substack{ S \subset CS \\ 
+                      |S| = N_{COAL} \\ 
+                      \forall s \in S,\ |s| \neq 0 
+                      \\ P_{s}^{CRCT} \geq P^{MIN}}}
+   \sum_{s \in S}  \Pr \bigl[ P_{s} \leq P_{s}^{CRCT} \bigr]
   \label{problem}
@@
 This section explains the process with which we form the coalitions (see the third block of Figure \ref{Fig1}).
-First, we need to simulate the timeseries of available power (first two blocks of Figure \ref{Fig1}).
+First, we need to simulate the time-series of available power (first two blocks of Figure \ref{Fig1}).
 We consider a pool $ \mathcal{A} $ of 200 agents, whose parameters were chosen randomly.
 The prosumers are positioned (also randomly) on a square lattice previously filled with climate vectors obtained from the french data sets (see section \ref{sec:model}).
-Simulations were run from february 2006 to december 2010 such that we are dealing with 200 hourly sampled timeseries of available power over this period.
+Simulations were run from february 2006 to december 2010 such that we are dealing with 200 hourly sampled time-series of available power over this period.
 Removing season trends finally leads us to the formation of coalitions.
 
-The model we used in order to simulate timeseries of available production provides some diversity because of the combination between the energetic mix (renewable generators combination) and the climate vectors.
-Nevertheless, as the number of agents grows, the timeseries tend to exhibit similar patterns.
-This is apparent when creating a correlation graph $ G_{1}(\mathcal{A},E_{1},\omega_{1}) $ with the same kind of metric as \cite{Garas2008} or \cite{Onnela2004} ($ d_{ij} = 1 - \rho_{ij}^{2} $), where well defined clusters appear in the $ \epsilon $-graph for any values of $ \epsilon $.
+The model we used in order to simulate time-series of available production provides some diversity because of the combination between the energetic mix (renewable generators combination) and the climate vectors.
+Nevertheless, as the number of agents grows, the time-series tend to exhibit similar patterns.
+This is apparent when creating a correlation graph with similar metric suc as defined in \cite{Garas2008} or \cite{Onnela2004} ($ d_{ij} = 1 - \rho_{ij}^{2} $), where well defined clusters appear in the $ \epsilon $-graph for any values of $ \epsilon $.
 
-However, these clusters of strongly correlated timeseries are the exact opposite of what we are seeking.
+However, these clusters of strongly correlated time-series are the exact opposite of what we are seeking.
 We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves on figure \ref{Fig4}), as expected, terrible values (far worse than a random split of the agents in the same number of coalitions).
 
-We thus opt for reversing the metric ($ d_{ij}^{'} = \rho_{ij}^{2} $) such that uncorrelated timeseries are close to each other and (anti)correlated timeseries are distant in $ G_{2}(\mathcal{A},E_{2},\omega_{2}) $.
-As expected \cite{Onnela2004}, independently of the $ \epsilon $ parameter, the resulting $ \epsilon $-graph exhibits henceforth much less clustering than in $ G_{1} $ and communities seem hardly visible.
+We thus opt for reversing the metric ($ d_{ij}^{'} = \rho_{ij}^{2} $) such that uncorrelated time-series are close to each other and (anti)correlated time-series are distant.
+We defined a Graph $G_\epsilon(\mathcal{A},E)$ where $\mathcal{A}$ are the agent and a edge $E$ is present if $ \rho_{ij}^{2} > \epsilon $.
+As expected \cite{Onnela2004}, independently of the $ \epsilon $ parameter, the resulting $ \epsilon $-graph exhibits henceforth much less clustering and communities seem hardly visible.
 Therefore, using classical clustering or community detection algorithms seem to provide poor results.
@@
 
-%\begin{figure}
-% \centering
-%  \subcaptionbox{Social Welfare\label{fig4:a}}{\includegraphics[width=1.6in]{figure7_with_green_curve_v1/uti5}}\hspace{1em}%
-%  \subcaptionbox{Market entering percentage\label{fig4:b}}{\includegraphics[width=1.6in]{figure4_with_green_curve/survival2_old}}
-%\caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.
-Blue curves represent clique percolation, red curves, the random process, and the green ones, correlated coalitions.}
-%\label{Fig4}
-%\end{figure}
-
 \begin{figure}

### 2014-09-24 Article.tex
update section4 and 5

@@
 \begin{figure}
-\includegraphics[scale=0.45]{figure2/Fig2}
+\includegraphics[scale=0.43]{figure2/Fig2}
 \caption{Process diagram}
@@
   $ CS $ & Set of all coalition structures \\
+  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
   \\
@@
 
-The model we used in order to simulate time-series of available production provides some diversity because of the combination between the energetic mix (renewable generators combination) and the climate vectors.
+The model we used in order to simulate time-series of available production provides some diversity because of the combination between the energetic mix (renewable generators combination) and climate vectors.
 Nevertheless, as the number of agents grows, the time-series tend to exhibit similar patterns.
-This is apparent when creating a correlation graph with similar metric suc as defined in \cite{Garas2008} or \cite{Onnela2004} ($ d_{ij} = 1 - \rho_{ij}^{2} $), where well defined clusters appear in the $ \epsilon $-graph for any values of $ \epsilon $.
+This is apparent when creating a correlation graph with similar metric such as defined in \cite{Garas2008} or \cite{Onnela2004} ($ 1 - \rho_{ij}^{2} $), where well defined clusters appear in the $ \epsilon $-graph for any values of $ \epsilon $.
 
@@
 
-We thus opt for reversing the metric ($ d_{ij}^{'} = \rho_{ij}^{2} $) such that uncorrelated time-series are close to each other and (anti)correlated time-series are distant.
-We defined a Graph $G_\epsilon(\mathcal{A},E)$ where $\mathcal{A}$ are the agent and a edge $E$ is present if $ \rho_{ij}^{2} > \epsilon $.
-As expected \cite{Onnela2004}, independently of the $ \epsilon $ parameter, the resulting $ \epsilon $-graph exhibits henceforth much less clustering and communities seem hardly visible.
+We thus opt for reversing the metric ($ \rho_{ij}^{2} $) such that uncorrelated time-series are close to each other in the graph and (anti)correlated time-series are distant.
+We defined a graph $G_\epsilon(\mathcal{A},E)$ where $\mathcal{A}$ is the set of agents and $ E $ the edge set.
+A given link between $ i $ and $ j $ is present if $ \rho_{ij}^{2} \leq \epsilon $, that is, $ e_{i,j} = \mathbf{1}_{\{\rho_{ij}^{2} \leq \epsilon \}}\rho_{ij}^{2} $.
+As expected \cite{Onnela2004}, independently of the $ \epsilon $ parameter, $ G_{\epsilon } $ exhibits henceforth much less clustering and communities seem hardly visible.
 Therefore, using classical clustering or community detection algorithms seem to provide poor results.
@@
 Such  structures contain indeed a link between every two nodes, meaning that the overall correlation is quite small.
-Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques, if it is too low, the $\epsilon$-graph of $ G_{2} $ will not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly.
-For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally repulsive.
+Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques : if it is too low, $ G_{\epsilon} $ does not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly.
+For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally intensive.
 Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ seems too brutal.
@@
 Cliques for small values of $ \epsilon $ appear thus as good seeds for stable coalitions.
-However we have to find a way to bring the coalitions above the grid requirements as well as increasing the social welfare by incorporating more nodes to the coalitions.
-Clique percolation algorithms have proven to be a useful heuristic for such complex problems.
-The main idea behind this algorithm is to make the seeds grow as long as they increase a fitness function based on internal and external degrees of the seeds \cite{Lancichinetti}.
-The main difference between our case study and that algorithm is that we chose this fitness function as our utility function (cf.
+We define $ N_{COAL} $ as the desired number of coalitions fixed by the user.
+Furthermore, let $ \Theta(G_{\epsilon}) $ be the set of non overlaping cliques in $ G_{\epsilon} $, we choose the $ \epsilon $ parameter as the smallest value such that $ G_{\epsilon} $ contains at least $ N_{COAL} $ non overlaping cliques :
+
+\begin{equation}
+\epsilon^{\star} = \min_{\substack{ \epsilon \in [0,1] }} \Big\{ \epsilon\,\ s.t\ |\Theta(G_{\epsilon})| \geq N_{COAL} \Big\}
+\end{equation}
+
+We construct $ G_{\epsilon^{\star}} $ such that it contains at least $ N_{COAL} $ potential seeds that may or may not be above the grid requirements, but exhibit low correlations among the members.
+Naturally, grid requirements could eventually be reached and social welfare increased by incorporating more nodes into the seeds.
+For this purpose, we use a clique percolation algorithm on $ G_{\epsilon^{\star}} $ that will extend the seeds over a correlation constrained environement.
+The main idea behind this algorithm is to make the seeds grow by local optimization of a fitness function, usually based on internal and external degrees of the seeds \cite{Lancichinetti}.
+The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (cf.
 eq.\eqref{utility}) that both maximizes the production and minimizes its volatility.
@@
 If node i belongs to  multiple coalitions, the only coalition including node i is the one that maximizes $ \tau_{i} $.
-
-At this point, we have three degrees of freedom : the reliability ($ \phi $), the required power to enter the market ($P^{MIN}$), and the prunning parameter ($\epsilon$).
-For convenience, we introduce the number of desired coalitions ($ N_{COAL} $) and we choose $\epsilon $ as the smallest value such that the $ \epsilon $-graph contains at least $ N_{COAL} $ non overlaping cliques.
-The overall utility depends now on the grid requirements and on the number of desired coalitions.
-
-Next section shows the utility behavior within the parameters space and provides the results of our 200 agents test.
+At this point, we have three degrees of freedom : the reliability ($ \phi $), the required power to enter the market ($P^{MIN}$), and the number of desired coalitions ($ N_{COAL} $).
+Next section shows the utility behavior within the parameters space $ \{\phi,P^{MIN},N_{COAL} \} $ and provides results of our 200 agents test.
 
@@
 As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high.
-Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utility values).
+Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utilities).
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$.
@@
 
-In figure \ref{fig2:b}, we fix the reliability value to a given empirical value ($\phi = 0.08 $).
+In figure \ref{fig2:b}, we fix the reliability to a given empirical value ($\phi = 0.08 $).
 We can observe how the social welfare $\sum_{s \in S} \mathcal{U}_{\phi,\ P^{MIN}}(s)$ evolves according to $P^{MIN}$ and $ N_{COAL} $.
-In our setting we can observe  a maximum point around $ N_{COAL} = 3 $, followed by a decrease.
+In our setting we notice a maximum point around $ N_{COAL} = 3 $, followed by a decrease.
 There is clearly a maximum number of coalitions sustainable for a given set of parameters.
@@
   \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots).
-The diameters of the dots are proportional to the corresponding coalition sizes.
-The hatched zone corresponds to the \textit{"under-requirement space"}, and results in zero utility for any coalition in this area.}
+The diameters of the dots are proportional to the coalition sizes.
+The hatched zero utility zone corresponds to the \textit{"under-requirement space"}.}
   \label{Fig3}
@@
 As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
-For consistency, we average the results of both plots of the random procedure over 100 realizations.
+For consistency, we average the results of both plots over 100 realizations (colored zones around the curves shows the standard deviations).
 

### 2014-09-24 figure1/Python/plot_fig.py
utility left plot python code

@@ -1,0 +1,55 @@
+import json
+import numpy as np
+import matplotlib
+import matplotlib.cm as cm
+import matplotlib.pyplot as plt
+
+def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
+    import matplotlib as mpl
+    cmap = cm.jet
+    mpl.rcParams['axes.labelsize'] = 9
+    mpl.rcParams['xtick.labelsize'] = 9
+    mpl.rcParams['ytick.labelsize'] = 9
+    mpl.rcParams['legend.fontsize'] = 9
+    #rcParams['font.family'] = 'serif'
+    #rcParams['font.serif'] = ['Computer Modern Roman']
+    mpl.rcParams['text.usetex'] = True
+    mpl.rcParams['figure.figsize'] = 7.3, 4.2
+    mpl.rcParams['axes.labelweight'] = 'bold'
+    mpl.rcParams['font.weight'] = 'bold'
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['lines.markersize'] = 8
+    mpl.rcParams['legend.numpoints'] = 1
+    # figure size in inch
+    mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
+    # figure dots per inch
+    mpl.rcParams['figure.dpi'] = 300
+    
+matplotlib_setup(4,3)
+
+with open("phi_Pmin_utility.json", "r") as fichier:
+    M = json.load(fichier)
+
+M = np.array(M)	
+maxi = M.max()
+normalized_M = np.vectorize(lambda x:float(x) / float(maxi))(M)
+
+fig = plt.figure()
+levels = np.arange(normalized_M.T.min(), normalized_M.T.max(), 0.2) 
+extent = (0,6,0,7)
+norm = cm.colors.Normalize(vmax=normalized_M.T.max(), vmin=normalized_M.T.min())
+cmap = cm.jet
+im = plt.imshow(normalized_M.T, origin='lower',extent=extent, cmap=cmap, norm=norm)
+v = plt.axis()
+plt.contour(normalized_M.T, levels, hold='on', colors = 'k', origin='lower', extent=extent)
+plt.axis(v)
[…]

### 2014-09-25 figure6/Python/plot_fig.py
utility right plot python code

@@ -1,0 +1,55 @@
+import json
+import numpy as np
+import matplotlib
+import matplotlib.cm as cm
+import matplotlib.pyplot as plt
+
+def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
+    import matplotlib as mpl
+    cmap = cm.jet
+    mpl.rcParams['axes.labelsize'] = 9
+    mpl.rcParams['xtick.labelsize'] = 9
+    mpl.rcParams['ytick.labelsize'] = 9
+    mpl.rcParams['legend.fontsize'] = 9
+    #rcParams['font.family'] = 'serif'
+    #rcParams['font.serif'] = ['Computer Modern Roman']
+    mpl.rcParams['text.usetex'] = True
+    mpl.rcParams['figure.figsize'] = 7.3, 4.2
+    mpl.rcParams['axes.labelweight'] = 'bold'
+    mpl.rcParams['font.weight'] = 'bold'
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['lines.markersize'] = 8
+    mpl.rcParams['legend.numpoints'] = 1
+    # figure size in inch
+    mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
+    # figure dots per inch
+    mpl.rcParams['figure.dpi'] = 300
+    
+matplotlib_setup(4,3)
+
+with open("Ncoal_Pmin_utility.json", "r") as fichier:
+    M = json.load(fichier)
+
+M = np.array(M)	
+maxi = M.max()
+normalized_M = np.vectorize(lambda x:float(x) / float(maxi))(M)
+
+fig = plt.figure()
+levels = np.arange(normalized_M.T.min(), normalized_M.T.max(), 0.15) 
+extent = (1,9,1,10)
+norm = cm.colors.Normalize(vmax=normalized_M.T.max(), vmin=normalized_M.T.min())
+cmap = cm.jet
+im = plt.imshow(normalized_M.T, origin='lower',extent=extent, cmap=cmap, norm=norm)
+v = plt.axis()
+plt.contour(normalized_M.T, levels, hold='on', colors = 'k', origin='lower', extent=extent)
+plt.axis(v)
[…]

### 2014-09-26 figure9/plot_fig.py
new version of last figure + python code

@@ -1,0 +1,121 @@
+import json
+import numpy as np
+import matplotlib
+import matplotlib.cm as cm
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import ImageGrid
+
+def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
+    import matplotlib as mpl
+    cmap = cm.jet
+    mpl.rcParams['axes.labelsize'] = 9
+    mpl.rcParams['xtick.labelsize'] = 9
+    mpl.rcParams['ytick.labelsize'] = 9
+    mpl.rcParams['legend.fontsize'] = 9
+    #rcParams['font.family'] = 'serif'
+    #rcParams['font.serif'] = ['Computer Modern Roman']
+    mpl.rcParams['text.usetex'] = True
+    mpl.rcParams['figure.figsize'] = 7.3, 4.2
+    mpl.rcParams['axes.labelweight'] = 'bold'
+    mpl.rcParams['font.weight'] = 'bold'
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['axes.linewidth'] = 0.75    
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['lines.markersize'] = 8
+    mpl.rcParams['legend.numpoints'] = 1
+    # figure size in inch
+    mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
+    # figure dots per inch
+    mpl.rcParams['figure.dpi'] = 300
+    
+matplotlib_setup(4,3)
+
+
+
+
+fig = plt.figure()
+grid = ImageGrid(fig, (1, 1, 1), nrows_ncols=(1, 2), axes_pad=0.2,
+                 add_all=True, label_mode='L', cbar_mode='single',
+                 cbar_location='right', cbar_pad=0.1)
+
+with open("phi_Pmin_utility.json", "r") as fichier:
+    M = json.load(fichier)
+M = np.array(M)	
+maxi = M.max()
+normalized_M = np.vectorize(lambda x:float(x) / float(maxi))(M)
+levels = np.arange(normalized_M.T.min(), normalized_M.T.max(), 0.2) 
+extent = (0,6,0,7)
[…]

### 2014-09-26 Article.tex
new version of utility plot

@@
 
+\begin{table}[t]
+\centering
+\scriptsize
+\begin{tabular}{l|p{4.5cm}}
+  \hline 
+  \\
+  $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
+  $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
+  $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time series of agent i available production values \\
+  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of the production of the agent $i$ \\
+  $ \phi \in [0,1] $ & Reliability threshold \\
+  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
+  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition s \\
+  $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
+  $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
+  $ \tau_{i}(s) \in \mathbb{R}$ & Decision rule for mapping overlapping into disjoint coalitions \\
+  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
+  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
+  $ CS $ & Set of all coalition structures \\
+  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
+  \\
+  \hline
+\end{tabular}
+
+\caption{Notations\label{table1}}
+\end{table}
+
 \subsection{Utility function}\label{subsec:UtilityFunc}
@@
 In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with low relative standard deviations, e.g high utility coalitions, can be computed.
-
-\begin{table}[!t]
-\centering
-\scriptsize
-\begin{tabular}{l|p{4.5cm}}
-  \hline 
-  \\
-  $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
-  $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
-  $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time series of agent i available production values \\
-  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of the production of the agent $i$ \\
-  $ \phi \in [0,1] $ & Reliability threshold \\
-  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
-  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition s \\
-  $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
-  $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
-  $ \tau_{i}(s) \in \mathbb{R}$ & Decision rule for mapping overlapping into disjoint coalitions \\
-  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
-  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
-  $ CS $ & Set of all coalition structures \\
-  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
-  \\
-  \hline
-\end{tabular}
-
-\caption{Notations\label{table1}}
-\end{table}
 
@@
  \centering
-  \subcaptionbox{$ (\phi, P^{MIN}) $ subspace\label{fig2:a}}{\includegraphics[width=1.7in]{./figure1/util}}
-  \subcaptionbox{$ (N_{COAL}, P^{MIN}) $ subspace\label{fig2:b}}{\includegraphics[width=1.70in]{./figure6/util2}}
-\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: In $ \{\phi \in [0,0.6], P^{MIN}\} $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $ (fig \ref{fig2:b}).
+  \includegraphics[scale=0.9]{./figure9/util}
+\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: a) In $ \{\phi \in [0,0.6], P^{MIN}\} $ when $ N_{COAL} $ is fixed.
+b) in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $.
 For readability, $ P^{MIN} $ is expressed in tenth of MW.}
@@
 
-As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high.
+As shown in figure \ref{Fig2}a, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high.
 Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utilities).
@@
 
-In figure \ref{fig2:b}, we fix the reliability to a given empirical value ($\phi = 0.08 $).
+In figure \ref{Fig2}b, we fix the reliability to a given empirical value ($\phi = 0.08 $).
 We can observe how the social welfare $\sum_{s \in S} \mathcal{U}_{\phi,\ P^{MIN}}(s)$ evolves according to $P^{MIN}$ and $ N_{COAL} $.

### 2014-09-26 figure9/plot_fig.py

@@ -29,7 +29,7 @@ def matplotlib_setup(figsize_x=8.3, figsize_y=4.2):
     # figure dots per inch
     mpl.rcParams['figure.dpi'] = 300
     
-matplotlib_setup(4,3)
+matplotlib_setup(4,2)
 
 
 
@@ -54,7 +54,7 @@ grid[0].contour(normalized_M.T, levels, hold='on', colors = 'k', origin='lower',
 grid[0].axis(v)
 #ax2.title("Utility", fontsize=9)
 grid[0].text(0.05, 0.95, 'a', transform=grid[0].transAxes, color='w', fontsize=16, fontweight='bold', va='top')
-grid[0].set_xlabel(r" \textbf{ $ \phi $ ( $ N_{COAL} = 10 $ ) }",fontsize=9)
+grid[0].set_xlabel(r" \textbf{ $ \phi*10 $ ( $ N_{COAL} = 10 $ ) }",fontsize=9)
 grid[0].set_ylabel(r" \textbf{ $ P^{MIN}$ }",fontsize=10)
 
 with open("Ncoal_Pmin_utility.json", "r") as fichier:

### 2014-09-26 Article.tex
modif michel (in progress)

@@
 
-There seems to have a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}.
+There seems to be a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}.
 Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
@@
 
-In section \ref{sec:model}, we consider that the agent's production depends on meteorological conditions (wind turbines and PV as generators), various energy mix and preferences, and their appliances (loads).
+In section \ref{sec:model}, we consider that the agent's production depends on meteorological conditions (wind turbines and photovoltaic panels (PV) as generators), various energy mix and preferences, and their appliances (loads).
 We use meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
@@
 
-On a narrower scale, the authors of \cite{Kota2011} study the formation of virtual power plants (VPP) composed of multiple self-interested DER.
-On the grid side, two requirements for the formation of virtual power plants are considered : the reliability of supply and the minimization of entities the grid has to deal with.
+On a narrower scale, \cite{Kota2011} study the formation, in a game theoretic setting, of virtual power plants (VPP) composed of multiple self-interested DER.
+Two requirements for the formation of virtual power plants are considered : the reliability of supply and the minimization of entities the grid has to deal with.
 From this, \cite{Kota2011} builds a pricing mechanism that encourages VPP to report true estimates of their aggregated production and penalizes prediction errors.
@@
 
-In this paper we ought to form coalitions utilities that depend on statistical properties of time-series values.
-The setting is thus similar with some financial studies on stock exchanges, where researchers tried to find relevant clusters of stocks based on their daily prices variations.
-The problem of clustering is usually approached by means of a similarity measure and completed by a clustering technique such as K-means or hierarchical clustering.
-Nevertheless, in \cite{Mantegna1999} the author introduced an approach where stocks time-series of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other.
+In this paper we do not consider stability against player defection but focus on a statistics-oriented definition of coalition stability.
+We ough indeed to form coalitions utilities that depend on statistical properties of time-series values.
+The setting is thus similar with some financial studies on stock exchanges, where researchers tried to find relevant correlated clusters of stocks based on their daily prices variations.
+There exists well-known algorithms such as K-means or hierarchical clustering that are traditionally used for such purposes.
+Nevertheless, despite their popularity, they do not appear particularly well suited for meaningfully translating complex correlation relationships between time series while clustering them, especially when these series exhibit complicated underlying patterns.
+In \cite{Mantegna1999} the author introduced an approach where stocks time-series of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other.
 This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
@@
 Presented in this way, the time-series clustering task seems very close to graph community detection.
+Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities, and several topology oriented techniques for finding communities are present in the litterature (\cite{Newmanb} \cite{Girvan2002} \cite{Newman}).
+For our purpose where decorrelation is the closeness notion, such algorithms tend to have some difficulties because only a few inclusions of very high correlations can strongly affect the stability of a coalition.
+However, local cliques, where all nodes are linked, provide uncorrelated groups of nodes.
+There exists greedy heuristics based on cliques in the litterature such as clique percolation \cite{Lancichinetti} which realizes local optimizations of a fitness function and results in overlaping community structures.
+Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities.
+
+%Presented in this way, the time-series clustering task seems very close to graph community detection.
 Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities \cite{Newmanb}.
@@
 
-Our major concern was to design a prosumer model with realistic patterns of consumption, production as well as realistic geographical correlations between them.
-We based our prosumer model on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).
+Our major concern was to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them.
+Only wind turbines and PV will be used as generators throughout this paper, and we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of prosumers.
+To account for the production of a prosumer, we relied on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).
+As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering three kinds of traces (wind speed, cloudiness, and temperature) that we organize in a structure we will refer to as "climate vector" in the following.
+We consider that climate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector.
+We used existing DER models enabling conversion between weather variables and output powers (see power curves in \cite{Kota2011} \cite{windturbinemodel}).
+\cite{Dans2007} provides also a convenient way of using cloudiness traces as realistic degradation factors in a \textit{"clear blue sky"} solar model, which enables us to rebuild realistic solar irradiance traces for stations that do not provide such information directly.
 
-As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering traces for these stations.
-We consider here that prosumers can only produce through wind turbines and photovoltaic panels (PV).
-On the other hand, we built their consumption patterns according to two major cycles :
+Consumption behaviors are characterized by a model we designed, which takes into account two major cycles :
+
 \begin{itemize}
@@
 \item \textbf{Seasonal cycles} : Consumption is higher in the winter because of heating and low in the summer (air conditioning is not considered).
-Temperature traces were the principal ingredient for modeling these cycles.
+Temperature traces are used for modeling these cycles.
 \end{itemize}
@@
 
-We thus collect for all chosen weather stations three kinds of data (average wind speed, cloudiness, and average temperature), which we will call a climate vector in the following.
-We consider that climate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector.
-
 At this point, modeling agents consists in fixing a few parameters (most of the time drawn from random distributions) such as the geographical position, the number of wind turbines, PV, appliances, the temperature of confort, and so on.
@@
 
-We thus used existing models enabling conversion between weather variables and output powers (see power curves in \cite{Kota2011} \cite{windturbinemodel}).
-\cite{Dans2007} provides also a convenient way of using cloudiness traces as realistic degradation factors in a \textit{"clear blue sky"} solar model, which enables us to rebuild realistic solar irradiance traces for stations that do not provide such information directly.
-Any agent is thus able to compute at any time how much he produces and consumes.
-
-\subsection{Coalition formation model}\label{subsec:Coalition}
-More formally, we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of agents (the prosumers) and $ P_{i}(t) $ the instantaneous power value at agent i at instant t (its instantaneous production minus his instantaneous consumption, meaning that $ P_{i}(t) $ represents agent i available power at instant t).
+Let $ P_{i}(t) $ be the instantaneous available power of agent i at instant t (his instantaneous production minus his instantaneous consumption).
 During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
@@
 
-We now extend these notations to any coalition $ s \subset \mathcal{A} $ : 
+\subsection{Coalition formation model}\label{subsec:Coalition}
+
+We extend the agent notations to any coalition $ s \subset \mathcal{A} $ : 
 \begin{itemize}

### 2014-09-26 Article.tex
modif michel

@@
 It is of the responsibility of the information system that forms coalitions to select agents in order to provide a production as stable as possible.
-Based on the $ \mathcal{F}(P_{s}) $ distribution, the probability of s value being below its contract value $ Pr[P_{s} \leq P_{s}^{CRCT} ] $ appears therefore as a quantity of interest that we would want to keep as low as possible.
+The probability of s value being below its contract value $ Pr[P_{s} \leq P_{s}^{CRCT} ] $ appears therefore as a quantity of interest that we would want to keep as low as possible.
 
-More formally, the rest of the paper derives a heuristic that aims at solving the following problem (Table \ref{table1} summarizes the main notations):
+In order to illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
+
+\begin{equation}
+\left\{ \begin{array}{lll}
+		\mu_{ij} = \mu_{i} + \mu_{j} \\
+		\sigma_{ij} = \sqrt{\sigma_{i}^{2} + \sigma_{j}^{2} + \rho_{ij}\sigma_{i}\sigma_{j}}
+\end{array} \right.
+\label{parameters}
+\end{equation}
+
+with $ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the probability $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] $ as :
+
+\begin{equation}
+\dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P_{ij}^{CRCT} - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right]
+\label{reliability}
+\end{equation}
+
+Let $ \phi \in [0,1] $ be the \textbf{reliability} constraint imposed to the coalition $\{ij\} $.
+It stipulates that if $\{ij\} $ wants to join the market, the probability of $\{ij\} $ value (at any instant t) being below its contract value should at most be $ \phi $.
+That is, $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] \leq \phi $.
+We also impose to $ \{ij\} $ a \textbf{minimum contract value} constraint denoted by $ P^{MIN} $, which states that the coalition should announce more than $ P^{MIN} $ to be accepted ($ P_{s}^{CRCT} \geq P^{MIN} $).
+
+Coalition $ \{ij\} $ has thus potentially a set of possible contract values that respects both rules.
+The strategy of $ \{ij\} $ (and all other coalitions in the paper) consists then in choosing, if it exists, the maximum value of this constrained set.
+We denote by $ P_{\phi}(s) $ this maximum value for a coalition s.
+For $ \{ij\} $, $ P_{\phi}(s) $ can be computed using eq.
+\ref{reliability} : $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $.
+If  $ P_{\phi}(ij) \geq P^{MIN} $, then $ \{ij\} $ can announce $ P_{ij}^{CRCT} = P_{\phi}(ij)$ on the market, else it is not allowed to enter.
+It thus appears (as it was intuitively understandable) that, for equivalent sizes, coalitions with low relative standard deviations ( $ \sigma_{ij} / \mu_{ij} $ ) are able to announce higher contract values.
+
+What this paper investigates in the following is the developement of a heuristic that organizes prosumers such that the synergy terms of the standard deviations ($ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) are minimized.
+More formally, we consider the following problem (Table \ref{table1} summarizes the main notations):
  \begin{equation}
@@
 
-We explain now the two requirements that will help us solve this problem.
-Let $ \phi \in [0,1] $ be the \textbf{reliability} threshold imposed to the coalitions.
-It stipulates that for any coalition s willing to join the market, the probability of s value (at any instant t) being below its contract value should at most be $ \phi $.
-That is, $ \forall s \in 2^{N},\ Pr[P_{s} \leq P_{s}^{CRCT} ] \leq \phi $ ($ \phi $ being restricted to small values for consistency).
-
-We also consider the \textbf{minimum contract value} constraint denoted by $ P^{MIN} $, which states that only coalitions with higher contract values ($ P_{s}^{CRCT} \geq P^{MIN} $) will be accepted.
-
-At this point, if a coalition s wishes to join the market, it has to choose a contract value that fulfill the two conditions.
-For simplification, we consider that coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value that obeys the reliability rule (a value we denote by $ P_{\phi}(s) $).
-If $ P_{\phi}(s) \geq P^{MIN} $, meaning that it also obeys the minimum contract value rule, then s announces this value on the market : $ P_{s}^{CRCT} = P_{\phi}(s) $.
-Otherwise, coalition s is not able to enter.
+As explained in the example above, if a coalition s wishes to join the market, it has to choose a contract value within a constrained set.
+Only coalitions whose set is not empty are able to enter the market (valid coalitions).
+For simplification, we consider that valid coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value of this set ( $ P_{s}^{CRCT} = P_{\phi}(s) $).
 Basically, a coalition s is valid if and only if :
@@
 Obviously, maximising this utility function amounts to maximizing the coalition contract value with the minimum possible number of agents.
-
-In order to illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
-
-\begin{equation}
-\left\{ \begin{array}{lll}
-		\mu_{ij} = \mu_{i} + \mu_{j} \\
-		\sigma_{ij} = \sqrt{\sigma_{i}^{2} + \sigma_{j}^{2} + \rho_{ij}\sigma_{i}\sigma_{j}}
-\end{array} \right.
-\label{parameters}
-\end{equation}
-
-with $ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the reliability condition $ Pr[P_{ij} \leq P_{\phi}(ij) ] \leq \phi $ as :
-
-\begin{equation}
-\dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P_{\phi}(ij) - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] \leq \phi
-\label{reliability}
-\end{equation}
-
-The strategy of $ \{ij\} $ consists in maximizing its contract value as long as it respects the inequality \ref{reliability}, e.g to annonce $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $.
-It thus appears (as it was intuitively understandable) that, for equivalent sizes, coalitions with low relative standard deviations ( $ \sigma_{ij} / \mu_{ij} $ ) are able to announce higher contract values.
-
-What this paper investigates in the following is the developement of an algorithm that organizes prosumers such that the synergy term of the standard deviation ($ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) is minimized.
-In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with low relative standard deviations, e.g high utility coalitions, can be computed.
+In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with high utilities can be computed.
 

### 2014-09-26 Article.tex
update

@@
 There seems to be a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}.
-Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (such as coalitions) that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
+Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
 The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}.
@@
 
-In this paper, we focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable range of production.
+In this paper, we consider the case of prosumers grouping together into coalitions to be able to produce sufficient power which may be sold to the grid.
+If the coalition's production is sufficiently high and stable, the coalition can negociate on a day by day basis with the grid by announcing each day an expected production for the next day.
+We focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable range of production.
 By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly.
@@
 
-We extend the agent notations to any coalition $ s \subset \mathcal{A} $ : 
+Let us recall that our goal is to group prosumer into coalitions so that the global power production resulting of the superposition of individual prosumer's productions be sufficiently prdictible and above a minimum power level.
+We thus extend the agent notations to any coalition $ s \subset \mathcal{A} $ :
 \begin{itemize}
@@
 \item $ \mathcal{T}_{s} = \{ P_{s}(t_{0}),...,P_{s}(t_{K}) \} $ with $ \mathcal{F}(P_{s}) $ the corresponding probability distribution
-\item A coalition S has the possibility to announce a contract value $ P_{s}^{CRCT} $ on the market.
 \end{itemize}
 
-Due to the volatile nature of DER productions or user consumptions, the available power of a coalition will oscillates around $ P_{s}^{CRCT} $.
+Furthermore, we consider that a coalition S has the possibility to announce a contract value $ P_{s}^{CRCT} $ on the market.
+Due to the volatile nature of DER productions or user consumptions, the available power of a coalition will oscillate around $ P_{s}^{CRCT} $.
 It is of the responsibility of the information system that forms coalitions to select agents in order to provide a production as stable as possible.
@@
 
-In order to illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
+To illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with a production following a gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
 
@@
 
-with $ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the probability $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] $ as :
+$ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the probability $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] $ as :
 
@@
   $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
-  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value of coalition s \\
+  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value announced by coalition s \\
   $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
@@
 \subsection{Utility function}\label{subsec:UtilityFunc}
-We now choose a utility function (eq.
+Adressing the problem of eq.\ref{problem}, we choose a utility function (eq.
 \ref{utility}) that derives directly from the above remarks.

### 2014-09-29 Article.tex
6 pages update

@@
 Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
-The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}.
+Formation of coalitions inside a smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn}.
 These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.
@@
 \begin{figure}
-\includegraphics[scale=0.43]{figure2/Fig2}
+\includegraphics[scale=0.37]{figure2/Fig2}
 \caption{Process diagram}
@@
   \\
-  $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
   $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
-  $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time series of agent i available production values \\
-  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of the production of the agent $i$ \\
+  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of $i$'s production \\
   $ \phi \in [0,1] $ & Reliability threshold \\
@@
   $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
-  $ \tau_{i}(s) \in \mathbb{R}$ & Decision rule for mapping overlapping into disjoint coalitions \\
+  $ \tau_{i}(s) \in \mathbb{R}$ & Overlapping to disjoint coalitions mapping \\
   $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
@@
  \centering
-  \includegraphics[scale=0.9]{./figure9/util}
-\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: a) In $ \{\phi \in [0,0.6], P^{MIN}\} $ when $ N_{COAL} $ is fixed.
+  \includegraphics[scale=0.8]{./figure9/util}
+\caption{Utility as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: a) In $ \{\phi, P^{MIN}\} $ when $ N_{COAL} $ is fixed.
 b) in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $.
@@
   \centering
-  \includegraphics[scale=0.8]{./figure5/coals}
+  \includegraphics[scale=0.7]{./figure5/coals}
   \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots).
@@
  \centering
-  \includegraphics[scale=0.35]{figure8/fig8.pdf}
+  \includegraphics[scale=0.3]{figure8/fig8.pdf}
   \caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.

### 2014-09-29 figure8/plot_fig.py
fig markers

@@ -81,8 +81,8 @@ fig = plt.figure()
 ax = fig.add_subplot(131)
 ax.set_ylim([0,3])
 ax.plot( range(Pmax)[1:], perco2, 'o--', color='b', label='percolation')
-ax.plot( range(Pmax)[1:], rand2, 'o--', color='r', label='random')
-ax.plot( range(Pmax)[1:], corre2, 'o--', color='g', label='correlated')
+ax.plot( range(Pmax)[1:], rand2, 'g^--', color='r', label='random')
+ax.plot( range(Pmax)[1:], corre2, 'bs--', color='g', label='correlated')
 ax.fill_between(range(Pmax)[1:], perco_sup2, perco2, facecolor='blue', alpha=alph)
 ax.fill_between(range(Pmax)[1:], perco2, perco_inf2, facecolor='blue', alpha=alph)
 ax.fill_between(range(Pmax)[1:], rand_sup2, rand2, facecolor='red', alpha=alph)
@@ -100,8 +100,8 @@ ax2 = fig.add_subplot(132)
 ax2.set_ylim([0,1.2])
 ax2.set_xlim([0,11])
 ax2.plot( range(Pmax), perco, 'o--', color='b', label='percolation')
-ax2.plot( range(Pmax), rand, 'o--', color='r', label='random')
-ax2.plot( range(Pmax), corre, 'o--', color='g', label='correlated')
+ax2.plot( range(Pmax), rand, 'g^--', color='r', label='random')
+ax2.plot( range(Pmax), corre, 'bs--', color='g', label='correlated')
 ax2.fill_between(range(Pmax), perco_sup, perco, facecolor='blue', alpha=alph)
 ax2.fill_between(range(Pmax), perco, perco_inf, facecolor='blue', alpha=alph)
 ax2.fill_between(range(Pmax), rand_sup, rand, facecolor='red', alpha=alph)

### 2014-09-29 Article.tex
corrections

@@
 
-In this paper, we consider the case of prosumers grouping together into coalitions to be able to produce sufficient power which may be sold to the grid.
+In this paper, we consider the case where prosumers wanted to be clustered together into coalitions in order to be able to produce enough power which may be sold to the grid.
 If the coalition's production is sufficiently high and stable, the coalition can negociate on a day by day basis with the grid by announcing each day an expected production for the next day.
@@
 In this paper we do not consider stability against player defection but focus on a statistics-oriented definition of coalition stability.
-We ough indeed to form coalitions utilities that depend on statistical properties of time-series values.
+We want to form coalitions based on utilities that depend on statistical properties of time-series values.
 The setting is thus similar with some financial studies on stock exchanges, where researchers tried to find relevant correlated clusters of stocks based on their daily prices variations.
@@
 
-Our major concern was to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them.
+Our major concern is to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them.
 Only wind turbines and PV will be used as generators throughout this paper, and we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of prosumers.
-To account for the production of a prosumer, we relied on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).
+To account for the production of a prosumer, we rely on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).
 As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering three kinds of traces (wind speed, cloudiness, and temperature) that we organize in a structure we will refer to as "climate vector" in the following.
 We consider that climate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector.
-We used existing DER models enabling conversion between weather variables and output powers (see power curves in \cite{Kota2011} \cite{windturbinemodel}).
+We use existing DER models enabling conversion between weather variables and output powers (see power curves in \cite{Kota2011} \cite{windturbinemodel}).
 \cite{Dans2007} provides also a convenient way of using cloudiness traces as realistic degradation factors in a \textit{"clear blue sky"} solar model, which enables us to rebuild realistic solar irradiance traces for stations that do not provide such information directly.
@@
 We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves on figure \ref{Fig4}), as expected, terrible values (far worse than a random split of the agents in the same number of coalitions).
-
-We thus opt for reversing the metric ($ \rho_{ij}^{2} $) such that uncorrelated time-series are close to each other in the graph and (anti)correlated time-series are distant.
+We thus opt for reversing the metric ($ \rho_{ij}^{2} $) such that uncorrelated time-series are close to each other in the graph and correlated time-series are distant.
 We defined a graph $G_\epsilon(\mathcal{A},E)$ where $\mathcal{A}$ is the set of agents and $ E $ the edge set.
@@
 Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore.
-
 Cliques for small values of $ \epsilon $ appear thus as good seeds for stable coalitions.
@@
 Naturally, grid requirements could eventually be reached and social welfare increased by incorporating more nodes into the seeds.
-For this purpose, we use a clique percolation algorithm on $ G_{\epsilon^{\star}} $ that will extend the seeds over a correlation constrained environement.
+For this purpose, we use a clique percolation algorithm on $ G_{\epsilon^{\star}} $ that will extend the seeds over a correlation constrained environment.
 The main idea behind this algorithm is to make the seeds grow by local optimization of a fitness function, usually based on internal and external degrees of the seeds \cite{Lancichinetti}.
@@
 eq.\eqref{utility}) that both maximizes the production and minimizes its volatility.
-
 As explained in section \ref{sec:related}, clique percolation leads generally to overlapping communities.
@@
  \centering
-  \includegraphics[scale=0.8]{./figure9/util}
+  \includegraphics[scale=0.75]{./figure9/util}
 \caption{Utility as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: a) In $ \{\phi, P^{MIN}\} $ when $ N_{COAL} $ is fixed.
@@
 There is clearly a maximum number of coalitions sustainable for a given set of parameters.
-After this value we will form more coalitions that would be unable to pass the grid requirements.
+After that value the coalitions newly formed won't be able to pass the grid requirements.
 Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs, meaning that the algorithm performance will also decrease.
@@
   \centering
-  \includegraphics[scale=0.7]{./figure5/coals}
+  \includegraphics[scale=0.65]{./figure5/coals}
   \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots).
@@
 
-In figure \ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time.
+In fig.
+\ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time.
 We observe the result of our algorithm for a fixed number of coalitions of $N_{COAL} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $.
 We compare our algorithm with a random process that splits the agents in $ N_{COAL} $ coalitions.
-Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots).
-The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility.
+Fig.
+\ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots).
+The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot, the higher its utility.
 The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the \textit{"under-requirement space"}, meaning that whenever a coalition is in this zone, it has a null utility.
-
-Looking at figure \ref{Fig3}, we can see first that our algorithm performs better than the random formation algorithm at finding high valued coalitions.
-The blue dots allowed to enter the market (non hatched zone) are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).
+Looking at fig.
+\ref{Fig3}, we can see first that our algorithm performs better than the random formation algorithm at finding high valued coalitions.
+The blue dots allowed to enter the market (non hatched zone) are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in fig.
+\ref{Fig3}).
 
@@
 For consistency, we average the results of both plots over 100 realizations (colored zones around the curves shows the standard deviations).
-
 When the grid requirements are constant (figure \ref{Fig4}b), and the number of desired coalitions is low, clique percolation generally performs better than random case.
 Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed.
-
 When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
@@
 On this basis, we proposed a simple algorithm that seeks for uncorrelated prosumer patterns as potential seeds and expand them in coalitions able to rise above the grid requirements.
-
-The algorithm is validated against a random process, where we show that it performs better (coalitions are more stable and the overall production is more important) and that it exhibits a higher robustness/flexibility against grid requirements changes.
+We validated our algorithm against a random choice of coalitions, which can be considered as an average case.
+We showed  that it performs better (coalitions are more stable and the overall production is more important) and that it exhibits a higher robustness/flexibility against grid requirements changes.
+The worst case scenario is represented by coalitions of correlated prosumers.
 

### 2014-09-29 Article.tex
more recent newman refs

@@
 Presented in this way, the time-series clustering task seems very close to graph community detection.
-Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities, and several topology oriented techniques for finding communities are present in the litterature (\cite{Newmanb} \cite{Girvan2002} \cite{Newman}).
+Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities, and several topology oriented techniques for finding communities are present in the litterature (\cite{Newman2013} \cite{Girvan2002} \cite{Newman2013_2}).
 For our purpose where decorrelation is the closeness notion, such algorithms tend to have some difficulties because only a few inclusions of very high correlations can strongly affect the stability of a coalition.

### 2014-09-29 Article.tex
ICC 2014 refs

@@
 Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge only in remote farms, but also in residential areas.
-Together with electric vehicles, and demand side management tools, they will constitute the building blocks which will help to turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
-Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G) to ensure grid stability, and are commonly called prosumers \cite{Rathnayaka2012} \cite{Ramchurn}.
+Together with electric vehicles, and demand side management programs \cite{Samadi2014}, they will constitute the building blocks which will help to turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
+Such agents that both consume and produce energy are ready to make concessions (appliances delays, V2G) to ensure grid stability, and are commonly called prosumers \cite{6883384,Ramchurn}.
 
@@
 The worst case scenario is represented by coalitions of correlated prosumers.
-
 Interesting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analyzed through game theory.

### 2014-10-01 Article.tex
corrections

@@
 
-Let us recall that our goal is to group prosumer into coalitions so that the global power production resulting of the superposition of individual prosumer's productions be sufficiently prdictible and above a minimum power level.
+Let us recall that our goal is to group prosumers into coalitions so that the global power production resulting of the superposition of individual prosumer's productions be sufficiently predictible and above a minimum power level.
 We thus extend the agent notations to any coalition $ s \subset \mathcal{A} $ :
@@
 
-To illustrate what is done in the following, lets consider a very simple example with two agents, say i and j, with a production following a gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
+To illustrate what is done in the following, let us consider a very simple example with two agents, say i and j, with a production following a gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
 
@@
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$.
-Obviously, the highest utility values are found for high $ \phi $, because the coalitions formed are able to announce higher contract values, yielding higher utilities.
+Obviously, the highest utility values are found for high $ \phi $, because the formed coalitions are able to announce higher contract values, yielding higher utilities.
 

### 2015-01-14 Article.tex
update

@@
   \hline 
-  \\
   $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
@@
   $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
-  \\
   \hline
@@
   \centering
-  \includegraphics[scale=0.65]{./figure5/coals}
+  \includegraphics[scale=0.63]{./figure5/coals}
   \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots).

### 2015-01-14 Article.tex
typos_corrections

@@
 One of the key ideas in the smart grid revolves around the introduction of communication means inside the power grids.
-This could enable complex improvements in the energy management and leads progressively to a greener energetic system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
+This could enable complex improvements in the energy management and leads progressively to a greener energy system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
 Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge only in remote farms, but also in residential areas.
@@
 In section \ref{sec:model}, we consider that the agent's production depends on meteorological conditions (wind turbines and photovoltaic panels (PV) as generators), various energy mix and preferences, and their appliances (loads).
-We use meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energetic profiles as well as realistic geographical correlations between different agents.
+We use meteorological data (see section \ref{subsec:ProsumerModel}) to account for seasonal and daily variations in the prosumer's energy profiles as well as realistic geographical correlations between different agents.
 With these ingredients, we are able to run realistic simulations and record the different output profiles as time-series.
@@
 
-At this point, modeling agents consists in fixing a few parameters (most of the time drawn from random distributions) such as the geographical position, the number of wind turbines, PV, appliances, the temperature of confort, and so on.
+At this point, modeling agents consists in fixing a few parameters (most of the time drawn from random distributions) such as the geographical position, the number of wind turbines, PV, appliances, the temperature of comfort, and so on.
 The objective is that, depending on the weather of his zone, the DER and appliances he owns, and the way he decides to heat his home, a given prosumer is able to compute his production/consumption at any time.
@@
 During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
-As expected, the prosumer's time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both on the energetic mix and the habits.
+As expected, the prosumer's time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both on the energy mix and the habits.
 Nevertheless, these macro variations can hide the interesting information in the correlation coefficients and twist the rest of our process.
@@
 
-The model we used in order to simulate time-series of available production provides some diversity because of the combination between the energetic mix (renewable generators combination) and climate vectors.
+The model we used in order to simulate time-series of available production provides some diversity because of the combination between the energy mix (renewable generators combination) and climate vectors.
 Nevertheless, as the number of agents grows, the time-series tend to exhibit similar patterns.
@@
 For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally intensive.
-Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ seems too brutal.
+Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ does not seem elegant.
 Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore.
@@
 We believe that the originality of this paper lies in its willingness to exploit the de-correlation of prosumer profiles in order to build stable coalitions.
-In this direction, we presented a model based on meteorological traces, that captures the complex "\textit{energetic mix/climate vectors}" combination and generates realistic production and consumption patterns.
+In this direction, we presented a model based on meteorological traces, that captures the complex "\textit{energy mix/climate vectors}" combination and generates realistic production and consumption patterns.
 We then built a framework that enables the grid to specify stability and minimum production requirements for filtering the coalitions.
@@
 Interesting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analyzed through game theory.
-Showing that maintaining the coalitions formed with our algorithm necessitates less communication and less storage capacity could also conduct to a stimulating project.
-Besides, we believe that not restricting the algorithm to non overlapping coalitions and studying the strategies and weights of nodes with multiple options could lead to interesting works.
+Showing that maintaining the coalitions formed with our algorithm necessitates less communication and less storage capacity could also lead to a stimulating project.
+Besides, we believe that not restricting the algorithm to non overlapping coalitions and studying the strategies and weights of nodes with multiple options could lead to interesting research.
 

### 2015-01-14 Article.tex
bigger figs and few corrections

@@
 \DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 
@@
 \begin{figure}
-\includegraphics[scale=0.37]{figure2/Fig2}
+\includegraphics[scale=0.45]{figure2/Fig2}
 \caption{Process diagram}
@@
 
-Let $ P_{i}(t) $ be the instantaneous available power of agent i at instant t (his instantaneous production minus his instantaneous consumption).
+Let $ P_{i}(t) $ be the instantaneous available power of agent \textit{i} at instant \textit{t} (his instantaneous production minus his instantaneous consumption).
 During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
@@
 Thereby, we remove these seasonal effects for all agents (see second block of figure \ref{Fig1}).
-We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent i, and by $ \mathcal{F}(P_{i}) $ the corresponding probability distribution.
+We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent \textit{i}, and by $ \mathcal{F}(P_{i}) $ the corresponding probability distribution.
 
@@
 
-Furthermore, we consider that a coalition S has the possibility to announce a contract value $ P_{s}^{CRCT} $ on the market.
+Furthermore, we consider that a coalition \textit{s} has the possibility to announce a contract value $ P_{s}^{CRCT} $ on the market.
 Due to the volatile nature of DER productions or user consumptions, the available power of a coalition will oscillate around $ P_{s}^{CRCT} $.
 It is of the responsibility of the information system that forms coalitions to select agents in order to provide a production as stable as possible.
-The probability of s value being below its contract value $ Pr[P_{s} \leq P_{s}^{CRCT} ] $ appears therefore as a quantity of interest that we would want to keep as low as possible.
+The probability of \textit{s} value being below its contract value $ Pr[P_{s} \leq P_{s}^{CRCT} ] $ appears therefore as a quantity of interest that we would want to keep as low as possible.
 
-To illustrate what is done in the following, let us consider a very simple example with two agents, say i and j, with a production following a gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
+To illustrate what is done in the following, let us consider a very simple example with two agents, say \textit{i} and \textit{j}, with a production following a gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
 
@@
 Let $ \phi \in [0,1] $ be the \textbf{reliability} constraint imposed to the coalition $\{ij\} $.
-It stipulates that if $\{ij\} $ wants to join the market, the probability of $\{ij\} $ value (at any instant t) being below its contract value should at most be $ \phi $.
+It stipulates that if $\{ij\} $ wants to join the market, the probability of $\{ij\} $ value (at any instant \textit{t}) being below its contract value should at most be $ \phi $.
 That is, $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] \leq \phi $.
@@
 The strategy of $ \{ij\} $ (and all other coalitions in the paper) consists then in choosing, if it exists, the maximum value of this constrained set.
-We denote by $ P_{\phi}(s) $ this maximum value for a coalition s.
+We denote by $ P_{\phi}(s) $ this maximum value for a coalition \textit{s}.
 For $ \{ij\} $, $ P_{\phi}(s) $ can be computed using eq.
@@
 
-What this paper investigates in the following is the developement of a heuristic that organizes prosumers such that the synergy terms of the standard deviations ($ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) are minimized.
+%What this paper investigates in the following is the developement of a heuristic that organizes prosumers such that the synergy terms of the standard deviations ($ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) are minimized.
 More formally, we consider the following problem (Table \ref{table1} summarizes the main notations):
+
+In the following, we do not make any gaussian assumption as in the previous toy example.
+We investigate then the developement of a heuristic that tries to find a compromise between maximizing the production sold on the market (benefits) on one hand and minimizing the probability that the coalitions fail to sustain their contract values (stability) on the other hand (Table \ref{table1} summarizes the main notations) :
+
+\begin{equation}
+  \argmax _{\substack{ S \subset CS \\ 
+                      |S| = N_{COAL} \\ 
+                      \forall s \in S,\ |s| \neq 0 
+                      \\ P_{s}^{CRCT} \geq P^{MIN}}}
+   \sum_{s \in S}  P_{s}^{CRCT}
+  \label{problem}
+\end{equation}
+
  \begin{equation}
@@
 
-As explained in the example above, if a coalition s wishes to join the market, it has to choose a contract value within a constrained set.
+This last minimization objective is explored by organizing prosumers such that the synergy terms of the standard deviations ($ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) are minimized.
+As explained in the example above, if a coalition \textit{s} wishes to join the market, it has to choose a contract value within a constrained set.
 Only coalitions whose set is not empty are able to enter the market (valid coalitions).
 For simplification, we consider that valid coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value of this set ( $ P_{s}^{CRCT} = P_{\phi}(s) $).
-Basically, a coalition s is valid if and only if :
+Let us define a valid coalition more formally as a coalition satisfying the following rules :
 \begin{equation}
@@
   \hline 
+  \\
   $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
@@
   $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
+  \\
   \hline
@@
 We thus implemented a simple heuristic that considers nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss.
-More formally, for a coalition s and a node $ i \in s $, we define :
+More formally, for a coalition \textit{s} and a node $ i \in s $, we define\footnote{\label{note1}Not to be mingled with the Shapley value} :
 
@@
 
-If node i belongs to  multiple coalitions, the only coalition including node i is the one that maximizes $ \tau_{i} $.
+If node \textit{i} belongs to  multiple coalitions, the only coalition including node \textit{i} is the one that maximizes $ \tau_{i} $.
 At this point, we have three degrees of freedom : the reliability ($ \phi $), the required power to enter the market ($P^{MIN}$), and the number of desired coalitions ($ N_{COAL} $).
@@
  \centering
-  \includegraphics[scale=0.75]{./figure9/util}
+  \includegraphics[scale=0.9]{./figure9/util}
 \caption{Utility as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: a) In $ \{\phi, P^{MIN}\} $ when $ N_{COAL} $ is fixed.
@@
   \centering
-  \includegraphics[scale=0.63]{./figure5/coals}
+  \includegraphics[scale=0.9]{./figure5/coals}
   \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots).
@@
 
-In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario.
+In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve with squares), clique percolation (blue curve with circles), and correlated coalitions (green curves with triangles) that stands for a worst case scenario.
 As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
@@
  \centering
-  \includegraphics[scale=0.3]{figure8/fig8.pdf}
+  \includegraphics[scale=0.36]{figure8/fig8_bis.pdf}
   \caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.
-Blue curves represent clique percolation, red curves, the random process, and the green ones, correlated coalitions.}
+Blue curves with circles represent clique percolation, red curves with squares, the random process, and the green ones with triangles, correlated coalitions.}
  \label{Fig4}

### 2015-01-14 figure8/plot_fig.py

@@ -3,13 +3,13 @@ import matplotlib.pyplot as plt
 import matplotlib.cm as cm
 
 #matplotlib config
-def matplotlib_setup(figsize_x=15, figsize_y=5):
+def matplotlib_setup(figsize_x=12, figsize_y=5):
     import matplotlib as mpl
     cmap = cm.jet
-    mpl.rcParams['axes.labelsize'] = 9
-    mpl.rcParams['xtick.labelsize'] = 9
-    mpl.rcParams['ytick.labelsize'] = 9
-    mpl.rcParams['legend.fontsize'] = 9
+    mpl.rcParams['axes.labelsize'] = 13
+    mpl.rcParams['xtick.labelsize'] = 16
+    mpl.rcParams['ytick.labelsize'] = 16
+    mpl.rcParams['legend.fontsize'] = 14
     #rcParams['font.family'] = 'serif'
     #rcParams['font.serif'] = ['Computer Modern Roman']
     mpl.rcParams['text.usetex'] = True
@@ -78,7 +78,7 @@ matplotlib_setup()
 Pmax = 11; phi = 0.15; alph = .3
 
 fig = plt.figure()
-ax = fig.add_subplot(131)
+ax = fig.add_subplot(121)
 ax.set_ylim([0,3])
 ax.plot( range(Pmax)[1:], perco2, 'o--', color='b', label='percolation')
 ax.plot( range(Pmax)[1:], rand2, 'g^--', color='r', label='random')
@@ -91,12 +91,13 @@ ax.fill_between(range(Pmax)[1:], corre_sup2, corre2, facecolor='green', alpha=al
 ax.fill_between(range(Pmax)[1:], corre2, corre_inf2, facecolor='green', alpha=alph)
 ax.text(0.05, 0.95, 'a', transform=ax.transAxes, fontsize=16, fontweight='bold', va='top')
 ax.set_xlabel(r" \textbf{ $ N_{COAL}$ }",fontsize=14)
-ax.set_ylabel('Social Welfare')
+ax.set_ylabel('Social Welfare', fontsize = 15)
 ax.grid()
+ax.legend()
 
 Pmax = 12; phi = 0.15; nb_coal = 7; alph = .3
 
-ax2 = fig.add_subplot(132)
+ax2 = fig.add_subplot(122)
 ax2.set_ylim([0,1.2])
 ax2.set_xlim([0,11])
[…]

### 2015-01-29 Article.tex
few text additions

@@
 Nevertheless, despite their popularity, they do not appear particularly well suited for meaningfully translating complex correlation relationships between time series while clustering them, especially when these series exhibit complicated underlying patterns.
+We will show later in this paper that using K-means as a clustering technique for grouping prosumers together leads to less good results.
 In \cite{Mantegna1999} the author introduced an approach where stocks time-series of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other.
-This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
+This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient (see notations in table \ref{table1}) ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.
 
@@
 For our purpose where decorrelation is the closeness notion, such algorithms tend to have some difficulties because only a few inclusions of very high correlations can strongly affect the stability of a coalition.
-However, local cliques, where all nodes are linked, provide uncorrelated groups of nodes.
+Furthermore, these graphs tend to exhibit far less clustering than their correlation-based counterparts, which can increase the difficulty of finding communities.
+However, we will see in the following, that a clique (a complete subgraph) in our filtered decoralation-based graph exhibit, by construction, low correlation.
+We indeed removed in the filtering (see $\epsilon$-graphs) all links that stands for high correlation.
 There exists greedy heuristics based on cliques in the litterature such as clique percolation \cite{Lancichinetti} which realizes local optimizations of a fitness function and results in overlaping community structures.
 Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities.
+Section \ref{sec:forming} will explain how we used such techniques on the decorelation graph in order to form coalitions of prosumers.
 
@@
 Our major concern is to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them.
-Only wind turbines and PV will be used as generators throughout this paper, and we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of prosumers.
+Only wind turbines and PV (photovoltaic panels) will be used as generators throughout this paper, and we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of prosumers.
 To account for the production of a prosumer, we rely on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).
@@
 
-To illustrate what is done in the following, let us consider a very simple example with two agents, say \textit{i} and \textit{j}, with a production following a gaussian value probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
+To illustrate what is done in the following, let us consider a very simple toy example with two agents, say \textit{i} and \textit{j}, with production values following gaussian probability distributions $ \mathcal{F}(P_{i}) = \mathcal{N}(\mu_{i}, \sigma_{i}) $ and $ \mathcal{F}(P_{j}) = \mathcal{N}(\mu_{j}, \sigma_{j}) $, such that the joint probability distribution  $ \mathcal{F}(P_{ij}) $ of the coalition $\{ij\}$ is also a gaussian with the following parameters :
 
@@
 
-$ \rho_{ij} $ the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $. We can easily write the probability $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] $ as :
+$ \rho_{ij} $ is the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $ (see table \ref{table1} for notations).
+Because all distributions are gaussians, we can easily write the probability $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] $ as :
 
@@
 We denote by $ P_{\phi}(s) $ this maximum value for a coalition \textit{s}.
-For $ \{ij\} $, $ P_{\phi}(s) $ can be computed using eq.
+For coalition $ \{ij\} $ in the toy example, $ P_{\phi}(s) $ can be computed using eq.
 \ref{reliability} : $ P_{\phi}(ij) = \mu_{ij} - \sigma_{ij}\sqrt{2}erf^{-1}(1-2 \phi ) $.
@@
    \sum_{s \in S}  P_{s}^{CRCT}
-  \label{problem}
+  \label{problem_1}
 \end{equation}
@@
    \sum_{s \in S}  \Pr \bigl[ P_{s} \leq P_{s}^{CRCT} \bigr]
-  \label{problem}
+  \label{problem_2}
 \end{equation}
@@
   \\
-  $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
-  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of $i$'s production \\
+  $ d_{ij} $ & Distance metric between timeseries i and j \\
+  $ \rho_{ij} $ & Pearson's correlation coefficient between timeseries i and j \\
+  $ \mathcal{A} $ & Agent set \\
+  $ P_{i}(t) \in \mathbb{R} $ & Instantaneous production of agent $i$ at time $t$ \\
+  $ P_{s}(t) $ & Instantaneous production of coalition s \\
+  $ \mathcal{F}(P_{i}) $ & Probability distribution of $i$'s production \\
+  $ \mathcal{F}(P_{S}) $ & Probability distribution of coalition s production \\
+  $ \mathcal{N}(\mu, \sigma) $ & Normal distribution with mean $ \mu $ and standard deviation $ \sigma $ \\
   $ \phi \in [0,1] $ & Reliability threshold \\
@@
   $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
+  $ G_{\epsilon}(\mathcal{A}, E) $ & Filtered correlation graph \\
   $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
@@
 \subsection{Utility function}\label{subsec:UtilityFunc}
-Adressing the problem of eq.\ref{problem}, we choose a utility function (eq.
+Adressing the problem of eq.\ref{problem_1} and \ref{problem_2}, we choose a utility function (eq.
 \ref{utility}) that derives directly from the above remarks.
 If a coalition cannot provide a valid contract value, it receives naturally a utility of zero.
-Furthermore, it seems obvious that the utility should increase with the contract value.
-The $ 1/|s| $ term in eq.
+Furthermore, the utility increases with the overall production (sum of the contracted values) as stated in the objective of eq.\ref{problem_1}.
+The stability objective described in eq.\ref{problem_2} is also present in this utility function through the $ \phi $ parameter.
+Finally, the $ 1/|s| $ term in eq.
 \ref{utility} indicates that we favor small coalitions, mainly because they are easier to maintain in terms of communications.
@@
 However, these clusters of strongly correlated time-series are the exact opposite of what we are seeking.
-We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves on figure \ref{Fig4}), as expected, terrible values (far worse than a random split of the agents in the same number of coalitions).
+We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves with squares on figure \ref{Fig4}), as expected, terrible values.
+We will indeed consider in section \ref{sec:results} this case as a worst case scenario, far worse than a trivial random split of the agents in the same number of coalitions.
+
 We thus opt for reversing the metric ($ \rho_{ij}^{2} $) such that uncorrelated time-series are close to each other in the graph and correlated time-series are distant.
@@
 Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore.
-Cliques for small values of $ \epsilon $ appear thus as good seeds for stable coalitions.
+Cliques for small values of $ \epsilon $ appear thus as good seeds for forming stable coalitions rather than directly acceptable coalitions.
 We define $ N_{COAL} $ as the desired number of coalitions fixed by the user.
@@
 The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (cf.
-eq.\eqref{utility}) that both maximizes the production and minimizes its volatility.
+eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem_1} and eq.\ref{problem_2}).
 As explained in section \ref{sec:related}, clique percolation leads generally to overlapping communities.
@@
 We thus implemented a simple heuristic that considers nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss.
-More formally, for a coalition \textit{s} and a node $ i \in s $, we define\footnote{\label{note1}Not to be mingled with the Shapley value} :
+More formally, for a coalition \textit{s} and a node $ i \in s $, we define\footnote{\label{note1}Not to be mingled with the Shapley value concept} :
 

### 2015-02-02 Article.tex
modifs

@@
 Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
+
 Formation of coalitions inside a smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn}.
@@
 If the coalition's production is sufficiently high and stable, the coalition can negociate on a day by day basis with the grid by announcing each day an expected production for the next day.
+
 We focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable range of production.
@@
 Meeting contract values for the prosumers is especially relevant in power trade conditions, where energy is traded based on day-ahead predictions.
+
 In this context, participants should try to minimize their prediction errors in order to maintain a stable state for the grid.
@@
 We will show later in this paper that using K-means as a clustering technique for grouping prosumers together leads to less good results.
+
 In \cite{Mantegna1999} the author introduced an approach where stocks time-series of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other.
@@
 Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities, and several topology oriented techniques for finding communities are present in the litterature (\cite{Newman2013} \cite{Girvan2002} \cite{Newman2013_2}).
+
 For our purpose where decorrelation is the closeness notion, such algorithms tend to have some difficulties because only a few inclusions of very high correlations can strongly affect the stability of a coalition.
 Furthermore, these graphs tend to exhibit far less clustering than their correlation-based counterparts, which can increase the difficulty of finding communities.
+
 However, we will see in the following, that a clique (a complete subgraph) in our filtered decoralation-based graph exhibit, by construction, low correlation.
@@
 There exists greedy heuristics based on cliques in the litterature such as clique percolation \cite{Lancichinetti} which realizes local optimizations of a fitness function and results in overlaping community structures.
+
 Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities.
@@
 To account for the production of a prosumer, we rely on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).
+
 As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering three kinds of traces (wind speed, cloudiness, and temperature) that we organize in a structure we will refer to as "climate vector" in the following.
@@
 In the following, we do not make any gaussian assumption as in the previous toy example.
-We investigate then the developement of a heuristic that tries to find a compromise between maximizing the production sold on the market (benefits) on one hand and minimizing the probability that the coalitions fail to sustain their contract values (stability) on the other hand (Table \ref{table1} summarizes the main notations) :
+We investigate then the developement of a heuristic that tries to find a compromise between maximizing the production sold on the market (benefits) on one hand (see eq.
+\ref{problem_1}) and minimizing the probability that the coalitions fail to sustain their contract values (stability) on the other hand (see eq.
+\ref{problem_2}).
+Table \ref{table1} summarizes the main notations.
 
@@
   $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
+  $ \mathbf{1}_{X} $ & Indicator function over set X \\ 
   \\
@@
 We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves with squares on figure \ref{Fig4}), as expected, terrible values.
-We will indeed consider in section \ref{sec:results} this case as a worst case scenario, far worse than a trivial random split of the agents in the same number of coalitions.
+We will indeed consider in section \ref{sec:results} this case as a worst case scenario, far worse than a trivial random split of the agents in the same number of coalitions (see the red curves with triangles of the random case on figure \ref{Fig4}).
 
@@
 Such  structures contain indeed a link between every two nodes, meaning that the overall correlation is quite small.
+
 Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques : if it is too low, $ G_{\epsilon} $ does not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly.
 For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally intensive.
+
 Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ does not seem elegant.
@@
 Cliques for small values of $ \epsilon $ appear thus as good seeds for forming stable coalitions rather than directly acceptable coalitions.
+
 We define $ N_{COAL} $ as the desired number of coalitions fixed by the user.
@@
 eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem_1} and eq.\ref{problem_2}).
+
 As explained in section \ref{sec:related}, clique percolation leads generally to overlapping communities.
@@
 Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utilities).
-Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$.
+Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions could be formed for a larger spectrum of $ P^{MIN}$.
 Obviously, the highest utility values are found for high $ \phi $, because the formed coalitions are able to announce higher contract values, yielding higher utilities.
@@
 
-In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve with squares), clique percolation (blue curve with circles), and correlated coalitions (green curves with triangles) that stands for a worst case scenario.
+In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve with triangles), clique percolation (blue curve with circles), and correlated coalitions (green curves with squares) that stands for a worst case scenario.
+
 As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
@@
 When the grid requirements are constant (figure \ref{Fig4}b), and the number of desired coalitions is low, clique percolation generally performs better than random case.
-Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed.
+Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumbles down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed.
 When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
 But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions is going down very quickly.
+
 After a few increases in $ P^{MIN} $, the percentage of the random procedure starts dropping while it stays almost constant for our algorithm.
 For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm.
-Finally, when the grid requirements becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).
+Finally, when the grid requirements become too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).
 
@@
 We then built a framework that enables the grid to specify stability and minimum production requirements for filtering the coalitions.
+
 On this basis, we proposed a simple algorithm that seeks for uncorrelated prosumer patterns as potential seeds and expand them in coalitions able to rise above the grid requirements.
@@
 The worst case scenario is represented by coalitions of correlated prosumers.
+
 Interesting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analyzed through game theory.

### 2015-02-09 Article.tex
typo
minipage update
big updates

@@
 \usepackage{array}
+
+\usepackage{float}
+\usepackage{multicol}
 
@@
 This could enable complex improvements in the energy management and leads progressively to a greener energy system \cite{Ramchurn} \cite{WuHamedHuangBook2011}.
-Distributed energy resources (DER) such as wind turbines or photovoltaic panels are not supposed to emerge only in remote farms, but also in residential areas.
+Distributed energy resources (DER) such as wind turbines or photovoltaic panels will probably not only appear in remote farms, but also in residential areas.
 Together with electric vehicles, and demand side management programs \cite{Samadi2014}, they will constitute the building blocks which will help to turn the today pure energy consumers into true actors of the grid operation \cite{Ramchurn}.
@@
 
-In this paper, we consider the case where prosumers wanted to be clustered together into coalitions in order to be able to produce enough power which may be sold to the grid.
+In this paper, we consider the case where prosumers coalesce into virtual groups (the so-called coalitions) in order to be able to produce enough power which may be sold to the grid.
 If the coalition's production is sufficiently high and stable, the coalition can negociate on a day by day basis with the grid by announcing each day an expected production for the next day.
@@
 Nevertheless, despite their popularity, they do not appear particularly well suited for meaningfully translating complex correlation relationships between time series while clustering them, especially when these series exhibit complicated underlying patterns.
-We will show later in this paper that using K-means as a clustering technique for grouping prosumers together leads to less good results.
+We will show later in this paper that using K-means as a clustering technique for grouping prosumers together leads to less good results (see fig.
+\ref{Fig4}).
 
@@
 \begin{figure}
-\includegraphics[scale=0.45]{figure2/Fig2}
+\includegraphics[scale=0.51]{figure2/Fig2}
 \caption{Process diagram}
@@
 
-In the following, we do not make any gaussian assumption as in the previous toy example.
+In the following, we do not make any gaussian assumption as in the previous toy example (see for instance the densities at the coalition level on figure \ref{Fig5}).
 We investigate then the developement of a heuristic that tries to find a compromise between maximizing the production sold on the market (benefits) on one hand (see eq.
@@
   \hline 
-  \\
   $ d_{ij} $ & Distance metric between timeseries i and j \\
@@
   $ \mathbf{1}_{X} $ & Indicator function over set X \\ 
-  \\
   \hline
@@
 Finally, the $ 1/|s| $ term in eq.
-\ref{utility} indicates that we favor small coalitions, mainly because they are easier to maintain in terms of communications.
+\ref{utility} indicates that we favor small coalitions, mainly because they are easier to maintain.
 
@@
 
-Obviously, maximising this utility function amounts to maximizing the coalition contract value with the minimum possible number of agents.
+Obviously, maximizing this utility function amounts to maximizing the coalition contract value with the minimum possible number of agents.
 In such settings, and through the clique percolation procedure (section \ref{sec:forming}), we will show that coalitions with high utilities can be computed.
@@
 However, these clusters of strongly correlated time-series are the exact opposite of what we are seeking.
-We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves with squares on figure \ref{Fig4}), as expected, terrible values.
+We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves with diamonds on figure \ref{Fig4}), as expected, terrible values.
 We will indeed consider in section \ref{sec:results} this case as a worst case scenario, far worse than a trivial random split of the agents in the same number of coalitions (see the red curves with triangles of the random case on figure \ref{Fig4}).
@@
 Such  structures contain indeed a link between every two nodes, meaning that the overall correlation is quite small.
-
 Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques : if it is too low, $ G_{\epsilon} $ does not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly.
@@
 Cliques for small values of $ \epsilon $ appear thus as good seeds for forming stable coalitions rather than directly acceptable coalitions.
-
 We define $ N_{COAL} $ as the desired number of coalitions fixed by the user.
@@
 The main idea behind this algorithm is to make the seeds grow by local optimization of a fitness function, usually based on internal and external degrees of the seeds \cite{Lancichinetti}.
-The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (cf.
-eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem_1} and eq.\ref{problem_2}).
+The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (see eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem_1} and eq.\ref{problem_2}).
 
@@
 
-In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve with triangles), clique percolation (blue curve with circles), and correlated coalitions (green curves with squares) that stands for a worst case scenario.
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.53]{./figure11/densities.pdf}
+  \caption{Production probability densities for four coalitions.
+One obtained with our algorithm based on clique percolation (blue curve with circles), one with a random formation (red curve with triangles), another one with a K-means approach (yellow curve with squares), and finally one with correlated agents (green curves with diamonds).}
+  \label{Fig5}
+\end{figure}
 
+In the following, we consider and compare four different methods : our method based on clique percolation (blue curve with circles), the previous random aggregation of agents (red curve with triangles), a K-means clustering approach (yellow curve with squares), and correlated coalitions (green curves with diamonds) obtained with a correlation graph and that stands for a "\textit{worst case scenario}".
+
+As visible on figure \ref{Fig5}, the production densities at the coalition level exhibit very different shapes.
+While our algorithm results in a nice density with a low (compared to the other curves) standard deviation, the other methods seem to form coalitions that will have a very large spectrum for their production.
+In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for the same four methods.
 As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
 For consistency, we average the results of both plots over 100 realizations (colored zones around the curves shows the standard deviations).
-When the grid requirements are constant (figure \ref{Fig4}b), and the number of desired coalitions is low, clique percolation generally performs better than random case.
+When the grid requirements are constant (fig \ref{Fig4}b), and the number of desired coalitions is low, clique percolation performs better than random case.
 Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumbles down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed.
+
+\begin{figure}
+\begin{minipage}[c]{\textwidth}
+	\begin{center}
+  		\includegraphics[scale=0.45]{figure10/perc2.pdf}
+  			\caption{Evolution of the social welfare when the number of coalitions is varying \textbf{a)}, and the percentage of coalitions able to enter the market for different values of $ P_{MIN} $ \textbf{b)}.
+Blue curves with circles and red curves with triangles represent respectively the results for clique percolation and the random process.
+For comparison, the green curves with diamonds and the yellow ones with squares show respectively the results obtained with coalitions of correlated agents and with a K-means clustering.}
+ 		\label{Fig4}
+	\end{center}
+\end{minipage}
+
+\end{figure}
+
+\begin{minipage}{\textwidth}
+\begin{multicols}{2}
 When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $.
-But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions is going down very quickly.
+But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions going down very quickly.
 
@@
 Finally, when the grid requirements become too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).
-
-\begin{figure}
- \centering
-  \includegraphics[scale=0.36]{figure8/fig8_bis.pdf}
-  \caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $.
-Blue curves with circles represent clique percolation, red curves with squares, the random process, and the green ones with triangles, correlated coalitions.}
- \label{Fig4}
-\end{figure}
 
@@
 \bibliography{Article}
+
+\end{multicols}
+
+\end{minipage}
+
 \end{document}

### 2015-02-09 figure10/script.py

@@ -1,0 +1,161 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import random
+import json
+
+#matplotlib config
+def matplotlib_setup(figsize_x=15, figsize_y=5):
+    import matplotlib as mpl
+    #cmap = cm.jet
+    mpl.rcParams['axes.labelsize'] = 13
+    mpl.rcParams['xtick.labelsize'] = 16
+    mpl.rcParams['ytick.labelsize'] = 16
+    mpl.rcParams['legend.fontsize'] = 14
+    #rcParams['font.family'] = 'serif'
+    #rcParams['font.serif'] = ['Computer Modern Roman']
+    mpl.rcParams['text.usetex'] = True
+    mpl.rcParams['axes.labelweight'] = 'bold'
+    mpl.rcParams['font.weight'] = 'bold'
+    mpl.rcParams['axes.linewidth'] = 0.75
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['axes.linewidth'] = 0.75
+    mpl.rcParams['lines.linewidth'] = 2
+    mpl.rcParams['lines.markersize'] = 8
+    mpl.rcParams['legend.numpoints'] = 1
+    # figure size in inch
+    mpl.rcParams['figure.figsize'] = figsize_x, figsize_y
+    # figure dots per inch
+    mpl.rcParams['figure.dpi'] = 300
+    #mpl.rcParams['figure.autolayout'] = True
+
+matplotlib_setup()
+#Load curve data subplot a
+perc_mean_utilities_percolation = []; perc_mean_utilities_random = []; perc_mean_utilities_corre = [] 
+perc_std_utilities_percolation = []; perc_mean_utilities_Kmeans = []; perc_std_utilities_Kmeans = []
+perc_std_utilities_random = []; perc_std_utilities_corre = []
+
+with open('./perc_percolation_mean.json', "r") as fichier:
+    perc_mean_utilities_percolation = json.load(fichier)
+with open('./perc_percolation_std.json', "r") as fichier:
+    perc_std_utilities_percolation = json.load(fichier)
+with open('./perc_random_mean.json', "r") as fichier:
+    perc_mean_utilities_random = json.load(fichier)
+with open('./perc_random_std.json', "r") as fichier:
+    perc_std_utilities_random = json.load(fichier)
[…]

### 2015-02-09 Article.tex
update

@@
 After a few increases in $ P^{MIN} $, the percentage of the random procedure starts dropping while it stays almost constant for our algorithm.
-For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm.
+For $ P^{MIN} = 8 $, we see that $ 100 \% $ of the coalitions are able to enter with our algorithm while approximately $ 75 \%$ of them enters for the random case (only $40 \%$ and $15 \%$ for the other cases).
 Finally, when the grid requirements become too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).

### 2015-02-09 Article.tex
a few modifs

@@
 In the following, we do not make any gaussian assumption as in the previous toy example (see for instance the densities at the coalition level on figure \ref{Fig5}).
-We investigate then the developement of a heuristic that tries to find a compromise between maximizing the production sold on the market (benefits) on one hand (see eq.
-\ref{problem_1}) and minimizing the probability that the coalitions fail to sustain their contract values (stability) on the other hand (see eq.
-\ref{problem_2}).
+We investigate then the developement of a heuristic that seeks to maximize the production sold on the market (sum of the contract values for all coalitions) under the \textit{reliability} and \textit{minimum contract value} constraints for every coalition (see eq.
+\ref{problem}).
 Table \ref{table1} summarizes the main notations.
@@
 \begin{equation}
-  \argmax _{\substack{ S \subset CS \\ 
-                      |S| = N_{COAL} \\ 
-                      \forall s \in S,\ |s| \neq 0 
-                      \\ P_{s}^{CRCT} \geq P^{MIN}}}
-   \sum_{s \in S}  P_{s}^{CRCT}
-  \label{problem_1}
+\begin{aligned}
+& \underset{S \subset CS}{ \text{ Maximize }}
+& & \sum_{s \in S}  P_{s}^{CRCT} \\
+& \text{subject to} 
+& &	|S| = N_{COAL} \\
+&&& \forall s \in S,\ |s| \neq 0  \\
+&&& \forall s \in S,\  P_{s}^{CRCT} \geq P^{MIN} \\
+&&& \forall s \in S,\  Pr[P_{s} \leq P_{\phi}(s)] \leq \phi
+\end{aligned}
+\label{problem}
 \end{equation}
 
- \begin{equation}
-  \argmin _{\substack{ S \subset CS \\ 
-                      |S| = N_{COAL} \\ 
-                      \forall s \in S,\ |s| \neq 0 
-                      \\ P_{s}^{CRCT} \geq P^{MIN}}}
-   \sum_{s \in S}  \Pr \bigl[ P_{s} \leq P_{s}^{CRCT} \bigr]
-  \label{problem_2}
-\end{equation}
+As explained in the example above, if a coalition \textit{s} wishes to join the market, it has to choose a contract value within a constrained set : $ P_{s}^{CRCT} \in \{ P^{MIN}, P_{\phi}(s) \}$.
+Only coalitions whose set is not empty ( $ P_{\phi}(s) \geq P^{MIN} $ ) are able to enter the market.
+More formally, let us define a \textit{valid coalition} as a coalition satisfying the following rules :
 
-This last minimization objective is explored by organizing prosumers such that the synergy terms of the standard deviations ($ \rho_{ij}\sigma_{i}\sigma_{j} $ in the example above) are minimized.
-As explained in the example above, if a coalition \textit{s} wishes to join the market, it has to choose a contract value within a constrained set.
-Only coalitions whose set is not empty are able to enter the market (valid coalitions).
-For simplification, we consider that valid coalitions will always apply the same economically consistent strategy of announcing the highest possible contract value of this set ( $ P_{s}^{CRCT} = P_{\phi}(s) $).
-Let us define a valid coalition more formally as a coalition satisfying the following rules :
 \begin{equation}
@@
 \end{equation}
+
+Normally, a valid coalition could select any value in the $ \{ P^{MIN}, P_{\phi}(s) \}$ set.
+However, for simplification, we consider that they will always apply the same economically consistent strategy of announcing the highest possible contract value of this set.
+Therefore, in the following, for a valid coalition we will use the simplification : $ P_{s}^{CRCT} = P_{\phi}(s) $.
 
@@
 \subsection{Utility function}\label{subsec:UtilityFunc}
-Adressing the problem of eq.\ref{problem_1} and \ref{problem_2}, we choose a utility function (eq.
-\ref{utility}) that derives directly from the above remarks.
-If a coalition cannot provide a valid contract value, it receives naturally a utility of zero.
-Furthermore, the utility increases with the overall production (sum of the contracted values) as stated in the objective of eq.\ref{problem_1}.
-The stability objective described in eq.\ref{problem_2} is also present in this utility function through the $ \phi $ parameter.
+As solving the optimization problem (see eq.
+\ref{problem}) is considered NP-hard, we address this problem with a heuristic based on a utility function (eq.
+\ref{utility}) that derives directly from the problem constraints.
+That is, if a coalition cannot provide a valid contract value, it receives a utility of zero.
+Furthermore, the utility increases with the overall production (sum of the contracted values) as stated in the objective of eq.\ref{problem}.
+The stability of the coalitions is also present in this utility function through the $ \phi $ parameter.
 Finally, the $ 1/|s| $ term in eq.
@@
 The main idea behind this algorithm is to make the seeds grow by local optimization of a fitness function, usually based on internal and external degrees of the seeds \cite{Lancichinetti}.
-The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (see eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem_1} and eq.\ref{problem_2}).
+The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (see eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem}).
 

### 2015-02-09 Article.tex
alpha modif
typos

@@
 \begin{abstract}
-In a smart grid environment, we study coalition formation of prosumers that aim at entering the energy market.
-It is paramount for the grid operation that the energy producers are able to sustain the grid demand in terms of stability and minimum production requirement.
-We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minimum energy level for the coalitions and a steady production level which leads to finding uncorrelated sources of energy  to form a coalition.
+In a smart grid environment, we study the coalition formation of prosumers that aim at entering the energy market.
+It is paramount for the grid operation that the energy producers are able to sustain the grid demand in terms of stability and a minimum production requirement.
+We design an algorithm that seeks to form coalitions that will meet both of these requirements:  a minimum energy level for the coalitions and a steady production level, which leads to finding uncorrelated sources of energy  to form a coalition.
 We propose an algorithm that uses graph tools such as correlation graphs or clique percolation to form coalitions that meet such complex constraints.
-We validate the algorithm against a random procedure and show that, it not only performs better in term of social welfare for the power grid, but also that it is more robust against unforeseen production variations due to changing weather conditions for instance.
+We validate the algorithm against a random procedure and show that, it not only performs better in terms of social welfare for the power grid, but also that it is more robust against unforeseen production variations due to changing weather conditions for instance.
 \end{abstract}
@@
 There seems to be a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}.
-Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
+Nevertheless, as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.
 
-Formation of coalitions inside a smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn}.
-These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.
+The formation of coalitions inside a smart grid environment could be applied to various types of agents such as self-sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn}.
+These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the meantime reduce the communication burden.
 
-In this paper, we consider the case where prosumers coalesce into virtual groups (the so-called coalitions) in order to be able to produce enough power which may be sold to the grid.
+In this paper, we consider the case where prosumers coalesce into virtual groups (so-called coalitions) in order to be able to produce enough power which may be sold to the grid.
 If the coalition's production is sufficiently high and stable, the coalition can negociate on a day by day basis with the grid by announcing each day an expected production for the next day.
 
-We focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable range of production.
+We focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable ranges of production.
 By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly.
@@
 
-In order to select groups of prosumers that lower the volatility of the coalition's productions, we define an algorithm based on a popular approach for stocks market clustering \cite{Mantegna1999}.
+In order to select groups of prosumers that lower the volatility of the coalition's productions, we define an algorithm based on a popular approach for stock market clustering \cite{Mantegna1999}.
 We use a modified clique percolation algorithm (see section \ref{sec:forming}) that will enable us to expand the cliques as needed in order to form proper coalitions that fulfill both the grid requirements and lower the production volatility.
@@
 
-Traditionally, forming coalitions in a pool of agents can be done either in a centralized way where a single central unit is responsible for all the computations or in a distributed way where agents have only local knowledges and take actions accordingly.
+Traditionally, forming coalitions in a pool of agents can be done either in a centralized way where a single central unit is responsible for all the computations or in a distributed way where agents have only local knowledge and take actions accordingly.
 It is of common use to represent the situation and assess the stability of the solution by using game theoretic tools.
-Some papers \cite{Saad2009} \cite{Luan2014} focus then on finding an optimal coalition structure giving a pool of autonomous self interested agents using distributed merge/split algorithms.
+Some papers \cite{Saad2009} \cite{Luan2014} focus then on finding an optimal coalition structure giving a pool of autonomous self-interested agents using distributed merge/split algorithms.
 
@@
 In \cite{Garas2008, Onnela2004}, the authors studied the topological properties (average clustering, connectivity, relative number of cliques) of the correlation graph against those of growing random graphs, depending on the threshold $ \epsilon $.
-However, there is no well defined method to select a right tradeoff $ \epsilon $ as a function of the network topology.
+However, there is no well-defined method to select a right tradeoff $ \epsilon $ as a function of the network topology.
 
@@
 
-However, we will see in the following, that a clique (a complete subgraph) in our filtered decoralation-based graph exhibit, by construction, low correlation.
-We indeed removed in the filtering (see $\epsilon$-graphs) all links that stands for high correlation.
-There exists greedy heuristics based on cliques in the litterature such as clique percolation \cite{Lancichinetti} which realizes local optimizations of a fitness function and results in overlaping community structures.
+However, we will see in the following, that a clique (a complete subgraph) in our filtered decorelation-based graph exhibit, by construction, low correlation.
+We indeed removed in the filtering (see $\epsilon$-graphs) all links that stand for high correlation.
+There exists greedy heuristics based on cliques in the litterature such as clique percolation, \cite{Lancichinetti} which realizes local optimizations of a fitness function and results in overlapping community structures.
 
 Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities.
-Section \ref{sec:forming} will explain how we used such techniques on the decorelation graph in order to form coalitions of prosumers.
+Section \ref{sec:forming} will explain how we used such techniques on the decorrelation graph in order to form coalitions of prosumers.
 
@@
 
-Our major concern is to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them.
+Our main concern is to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them.
 Only wind turbines and PV (photovoltaic panels) will be used as generators throughout this paper, and we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of prosumers.
@@
 
-As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering three kinds of traces (wind speed, cloudiness, and temperature) that we organize in a structure we will refer to as "climate vector" in the following.
+As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well-chosen weather stations and gathering three kinds of traces (wind speed, cloudiness, and temperature) that we organize in a structure we will refer to as "climate vector" in the following.
 We consider that climate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector.
@@
 \begin{itemize}
-\item \textbf{Daily cycles} : Consumption is low during night, and higher during the day with two picks in the morning and evening.
+\item \textbf{Daily cycles} : Consumption is low during the night, and higher during the day with two picks in the morning and evening.
 Some noise is added so that prosumers have similar but not identical cycles.
@@
 During the simulation (from $t_{0} $ to $ t_{K} $), all agents record their values with an hour time interval.
-As expected, the prosumer's time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both on the energy mix and the habits.
+As expected, the prosumer's time-series exhibit high seasonal patterns, but completely different from agent to agent because they depend both on the energy mix and their habits.
 Nevertheless, these macro variations can hide the interesting information in the correlation coefficients and twist the rest of our process.
-Thereby, we remove these seasonal effects for all agents (see second block of figure \ref{Fig1}).
+Thereby, we remove these seasonal effects for all agents (see block 2 of figure \ref{Fig1}).
 We denote by $ \mathcal{T}_{i} = \{ P_{i}(t_{0}),...,P_{i}(t_{K}) \} $ the resulting time-series for  agent \textit{i}, and by $ \mathcal{F}(P_{i}) $ the corresponding probability distribution.
@@
 
-Let us recall that our goal is to group prosumers into coalitions so that the global power production resulting of the superposition of individual prosumer's productions be sufficiently predictible and above a minimum power level.
+Let us recall that our goal is to group prosumers into coalitions so that the global power production resulting of the superposition of individual prosumer's productions be sufficiently predictable and above a minimum power level.
 We thus extend the agent notations to any coalition $ s \subset \mathcal{A} $ :
@@
   $ d_{ij} $ & Distance metric between timeseries i and j \\
-  $ \rho_{ij} $ & Pearson's correlation coefficient between timeseries i and j \\
+  $ \rho_{ij} $ & Pearson's correlation coefficient between time series i and j \\
   $ \mathcal{A} $ & Agent set \\
@@
   $ G_{\epsilon}(\mathcal{A}, E) $ & Filtered correlation graph \\
-  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
+  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of coalitions \\
   $ CS $ & Set of all coalition structures \\
-  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
+  $ \Theta(G) $ & Set of non overlapping cliques in graph G \\
   $ \mathbf{1}_{X} $ & Indicator function over set X \\ 
@@
 The stability of the coalitions is also present in this utility function through the $ \phi $ parameter.
-Finally, the $ 1/|s| $ term in eq.
-\ref{utility} indicates that we favor small coalitions, mainly because they are easier to maintain.
+Finally, the $ 1/|s|^{\alpha} $ term in eq.
+\ref{utility} controls the sizes of the coalitions (for simplicity, in the rest of the paper $ \alpha $ is set to one).
 
 \begin{equation}
- \mathcal{U}_{\phi,\ P^{MIN}}(s) = \mathbf{1}_{\textit{s\ valid}} \dfrac{P_{\phi}(s)}{|s|} 
+ \mathcal{U}_{\phi,\ P^{MIN}}(s) = \mathbf{1}_{\textit{s\ valid}} \dfrac{P_{\phi}(s)}{|s|^{\alpha}} 
 \label{utility}
@@
 Nevertheless, as the number of agents grows, the time-series tend to exhibit similar patterns.
-This is apparent when creating a correlation graph with similar metric such as defined in \cite{Garas2008} or \cite{Onnela2004} ($ 1 - \rho_{ij}^{2} $), where well defined clusters appear in the $ \epsilon $-graph for any values of $ \epsilon $.
+This is apparent when creating a correlation graph with similar metric such as defined in \cite{Garas2008} or \cite{Onnela2004} ($ 1 - \rho_{ij}^{2} $), where well-defined clusters appear in the $ \epsilon $-graph for any values of $ \epsilon $.
 
@@
 Such  structures contain indeed a link between every two nodes, meaning that the overall correlation is quite small.
-Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques : if it is too low, $ G_{\epsilon} $ does not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly.
+Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques : if it is too low, $ G_{\epsilon} $ does not provide enough cliques, conversely, if $\epsilon $ is too high, we lose important information as the graph becomes very dense and cliques tend to overlap strongly.
 For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally intensive.
@@
 Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ does not seem elegant.
-Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore.
+Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents, even if they don't form a clique anymore.
 Cliques for small values of $ \epsilon $ appear thus as good seeds for forming stable coalitions rather than directly acceptable coalitions.
 We define $ N_{COAL} $ as the desired number of coalitions fixed by the user.
-Furthermore, let $ \Theta(G_{\epsilon}) $ be the set of non overlaping cliques in $ G_{\epsilon} $, we choose the $ \epsilon $ parameter as the smallest value such that $ G_{\epsilon} $ contains at least $ N_{COAL} $ non overlaping cliques :
+Furthermore, let $ \Theta(G_{\epsilon}) $ be the set of non overlapping cliques in $ G_{\epsilon} $, we choose the $ \epsilon $ parameter as the smallest value such that $ G_{\epsilon} $ contains at least $ N_{COAL} $ non overlapping cliques :
 
@@
 We thus implemented a simple heuristic that considers nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss.
-More formally, for a coalition \textit{s} and a node $ i \in s $, we define\footnote{\label{note1}Not to be mingled with the Shapley value concept} :
+More formally, for a coalition \textit{s} and a node $ i \in s $, we define :
 
@@
 Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions could be formed for a larger spectrum of $ P^{MIN}$.
-Obviously, the highest utility values are found for high $ \phi $, because the formed coalitions are able to announce higher contract values, yielding higher utilities.
+Obviously, the highest utility values are found for high $ \phi $, because formed coalitions are able to announce higher contract values, yielding higher utilities.
 
@@
 Looking at fig.
-\ref{Fig3}, we can see first that our algorithm performs better than the random formation algorithm at finding high valued coalitions.
+\ref{Fig3}, we can see first that our algorithm performs better than the random formation algorithm at finding high-valued coalitions.
 The blue dots allowed to enter the market (non hatched zone) are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in fig.
@@
 As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $.
-For consistency, we average the results of both plots over 100 realizations (colored zones around the curves shows the standard deviations).
+For consistency, we average the results of both plots over 100 realizations (colored zones around the curves show the standard deviations).
 When the grid requirements are constant (fig \ref{Fig4}b), and the number of desired coalitions is low, clique percolation performs better than random case.

### 2015-02-17 Article.tex
last

@@
 
-However, we will see in the following, that a clique (a complete subgraph) in our filtered decorelation-based graph exhibit, by construction, low correlation.
+However, we will see in the following, that a clique (a complete subgraph) in our filtered decorelation-based graph exhibits, by construction, low correlation.
 We indeed removed in the filtering (see $\epsilon$-graphs) all links that stand for high correlation.
@@
 $ \rho_{ij} $ is the Pearson correlation coefficient between $ P_{i} $ and $ P_{j} $ (see table \ref{table1} for notations).
-Because all distributions are gaussians, we can easily write the probability $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] $ as :
+Because all distributions are gaussians in this example, we can easily write the probability $ Pr[P_{ij} \leq P_{ij}^{CRCT} ] $ as :
 
@@
 \subsection{Utility function}\label{subsec:UtilityFunc}
-As solving the optimization problem (see eq.
+As solving directly the optimization problem (see eq.
 \ref{problem}) is considered NP-hard, we address this problem with a heuristic based on a utility function (eq.
@@
 We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves with diamonds on figure \ref{Fig4}), as expected, terrible values.
-We will indeed consider in section \ref{sec:results} this case as a worst case scenario, far worse than a trivial random split of the agents in the same number of coalitions (see the red curves with triangles of the random case on figure \ref{Fig4}).
+We will indeed consider in section \ref{sec:results} this case as a \textit{"worst case scenario"}, far worse than a trivial random search (see the red curves with triangles of the random case on figure \ref{Fig4}).
 
@@
 After that value the coalitions newly formed won't be able to pass the grid requirements.
-Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs, meaning that the algorithm performance will also decrease.
+Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs and a decrease of the algorithm's performances.
 

### 2015-02-17 figure10/script.py

@@ -34,26 +34,25 @@ perc_mean_utilities_percolation = []; perc_mean_utilities_random = []; perc_mean
 perc_std_utilities_percolation = []; perc_mean_utilities_Kmeans = []; perc_std_utilities_Kmeans = []
 perc_std_utilities_random = []; perc_std_utilities_corre = []
 
-with open('./perc_percolation_mean.json', "r") as fichier:
+with open('./perc_percolation_mean_2.json', "r") as fichier:
     perc_mean_utilities_percolation = json.load(fichier)
-with open('./perc_percolation_std.json', "r") as fichier:
+with open('./perc_percolation_std_2.json', "r") as fichier:
     perc_std_utilities_percolation = json.load(fichier)
-with open('./perc_random_mean.json', "r") as fichier:
+with open('./perc_random_mean_2.json', "r") as fichier:
     perc_mean_utilities_random = json.load(fichier)
-with open('./perc_random_std.json', "r") as fichier:
+with open('./perc_random_std_2.json', "r") as fichier:
     perc_std_utilities_random = json.load(fichier)
-with open('./perc_corre_mean.json', "r") as fichier:
+with open('./perc_corre_mean_2.json', "r") as fichier:
     perc_mean_utilities_corre = json.load(fichier)
-with open('./perc_corre_std.json', "r") as fichier:
+with open('./perc_corre_std_2.json', "r") as fichier:
     perc_std_utilities_corre = json.load(fichier)
-with open('./perc_Kmeans_mean.json', "r") as fichier:
+with open('./perc_Kmeans_mean_2.json', "r") as fichier:
     perc_mean_utilities_Kmeans = json.load(fichier)
-with open('./perc_Kmeans_std.json', "r") as fichier:
+with open('./perc_Kmeans_std_2.json', "r") as fichier:
     perc_std_utilities_Kmeans = json.load(fichier)
 
-Pmax = 16
+Pmax = 21
 alph = 0.3	
-	
 
 #Load curve data subplot a
 mean_utilities_percolation = []; mean_utilities_random = []; mean_utilities_corre = [] 
@@ -64,19 +63,26 @@ with open('./percolation_mean.json', "r") as fichier:
[…]
